### 2022-08-03 how_to.qmd
testing add apps

---
title: "How to"
editor: visual
---

To create a folder of shinylive apps deployed to netlify within your repo...

1.  Get netlify CLI up an running - see here: [https://docs.netlify.com/cli/get-started](https://docs.netlify.com/cli/get-started/)
2.  Run `netlify init` in the command line (from your local repo).
    1.  Follow the options.

    2.  Provide a domain name

    3.  For the build command, use `hugo build`

    4.  For the directory to deploy, use the 'site' folder associated with the folder of apps. For me this is `./site_apps`

    5.  Accept default functions folder

    6.  Upon prompting ask to create new `netlify.toml` file