### 2022-08-03 test_pages/test_with_shinylive.qmd
Add publish and new app shell scripts

---
title: "A test notebook with Shinylive"
---

``` {r}
library(tidyverse)
library(here)
```

``` {r}
a<- c(1:10)
b<- a * 2 + 1

df<- data.frame(a, b)

fp<- paste0(here(),"/shiny_apps/app1/data.csv")

write.csv(df, fp)

```

```{=html}
<iframe src="https://shiny-apps.netlify.app/app1/" height="700" width="100%" title="Shinylive test"></iframe>
```