### 2022-10-07 PhD_pages/paper_pages/paper1.qmd

---
title: "Paper 1: Triangulating trends in inequality around the world from secondary databases"
---

```{r, echo=FALSE, message = FALSE, warning=FALSE}

library(tidyverse)
library(trackdown)

```

What has happened to within-country inequality globally?

Is it that there has been a secular rise - suggesting global forces.

Or is it mixed - suggesting much scope for national policy to influence?

Different headline - WID suggests general rise (find a quote). PIP much more mixed... even a fall.

Untangle which methods, definitions.

# Methods and data

## Data sources

### PIP

### WID

### LIS

### Other sources

Should we use Solt? As another check on results?

What about GCIP? For trying to flip between income and consumption...

## Coverage

## Lining up methods

Method 1: Reference years

Method 2: Fully extrapolating/interpolating

The equivalent of the reference year approach where there is no max threshold gap.

Method 3: For the time trend, use a regression with country fixed effects to allow for entry and exit from the sample.

## Comparing distributions

### Comparing welfare concepts

Getting a *global* view of inequalities within countries faces a Catch-22 situation: If you want broad coverage you get very heterogeneous data; if you want comparable data, or to be able to consider different welfare concepts consistently, the coverage you get is much smaller.

Here I'll take two avenues.

Firstly, what we're going for with the 'triangulation' approach is dominance - if different welfare metrics agree

Second we will cross reference to attempts to 'standardise' across these differences: using Solt and GCIP. How valid that is I don't know. But taking the two approaches together we get as close as we can to robust observations.

### A range of inequality metrics

### Lorenz curves

Dominance implies any (well-behaved) summary measure of inequality will agree on the ranking.

Operationalize this by taking 99 points on the Lorenze curve: the percentiles.

This leaves out the *very* top. But this would be quite a warped conception of inequality. I.e. if you have two distributions, A and B, and the share of income received by every whole percentile threshold in B is higher than A, but the Lorenze curves crossed within the top percentile - to claim that inequality is higher in B you'd need a very extreme conception of inequality.

# Results

### Outliers

Russia

# Conclusion