### 2022-10-07 how_to.qmd

---
title: "How I built this site"
editor: visual
---

This site is built on three main ingredients:

-   Quarto notebooks, gathered together into a Quarto website

-   These are published to the web via Netlify

-   Text editing is done using Google Docs, pulled in to the Quarto notebooks using the R **trackdown** package

<!-- To create a folder of shinylive apps deployed to netlify within your repo... -->

<!-- 1.  Get netlify CLI up an running - see here: [https://docs.netlify.com/cli/get-started](https://docs.netlify.com/cli/get-started/) -->

<!-- 2.  Run `netlify init` in the command line (from your local repo). -->

<!--     1.  Follow the options. -->

<!--     2.  Provide a domain name -->

<!--     3.  For the build command, use `hugo build` -->

<!--     4.  For the directory to deploy, use the 'site' folder associated with the folder of apps. For me this is `./site_apps` -->

<!--     5.  Accept default functions folder -->

<!--     6.  Upon prompting ask to create new `netlify.toml` file -->

<!-- To add a new zsh shell script... -->

<!-- 1.  dsf -->

<!-- 2.  Sort out permissions by running: `chmod u+x <FILNAME>.sh` -->