### 2022-10-11 PhD_pages/scripts/PIP/prepare_pip_percentiles.qmd

```{r}
library(tidyverse)

source('PhD_pages/scripts/functions.r')

```

```{r}
#These variables uniquely id a set of survey results
pip_id_vars<- c('Entity', 'Year', 'reporting_level', 'welfare_type')

```

```{r}
pip_perc_orig<- read.csv("PhD_pages/data/original/pip_percentiles.csv")

```

```{r}
pip_main_vars_orig<- read.csv("PhD_pages/data/original/pip_main_vars.csv")

```

#### Prep 'percentiles' data file

```{r}
pip_perc<- pip_perc_orig
```

```{r}
# Drop duplicate observations of 'headcount' - The PIP API returns the . Given the discrete nature of the survey data, it sometimes returns a headcount ratio that are fairly far from the target percentile queried and the same headcount ratio can be returned for multiple percentile queries. Here we keep only one.

pip_perc<- drop_dupes(pip_perc,
                       id_vars = pip_id_vars,
                        duplicate_data_var = 'headcount')

```

```{r}

# Browse for nonmonotonic rows
# browse_nonmonotonic<- monotonicity_check(
#   df = pip_perc, 
#   id_vars = pip_id_vars,
#   order_var = 'headcount',
#   data_var = 'percentile_value', 
#   action = 'browse')

#There seems to be an issue with China in 1993,  driven by rural China.
# br<- pip_perc %>%
#   filter(Entity == "China",
#          reporting_level == 'national',
#          Year == 1993)

# Let's drop these rows
pip_perc<- monotonicity_check(
  df = pip_perc, 
  id_vars = pip_id_vars,
  order_var = 'headcount',
  data_var = 'percentile_value', 
  action = 'drop')

```

```{r}
#Drop world regions from percentile data

# These are entities with no reporting level - as veried below
# pip_perc %>%
#   filter(reporting_level=="") %>% 
#   select(Entity) %>%
#   unique() %>%
#   print()

pip_perc<- pip_perc %>%
    filter(reporting_level!="")

```

#### Prep 'main vars' data file

```{r}

pip_main_vars<- pip_main_vars_orig %>%
  rename(Entity = country_name,
         Year = reporting_year) %>%
  select(Entity, Year, reporting_level, welfare_type, mean, reporting_pop, reporting_gdp, gini, mld)
```

#### Merge the percentile and main data files

```{r}
pip_perc<- left_join(pip_perc, pip_main_vars)

```

#### Additional cleaning

```{r}
#Keep only 'National' estimates, except for country-years that don't have national estimates.

# Count how many reporting levels there are for each estimate
count_reporting_levels<- pip_main_vars %>%
  select(Entity, Year, reporting_level) %>%
  unique() %>%
  count(Entity, Year) %>%
  rename(number_of_reporting_levels = n)

# We see there only 1 and 3 reporting levels - i.e. either national, rural or uban alone, or else all three are present.
count_reporting_levels %>% select(number_of_reporting_levels) %>% unique()

# Merge the counts back into dataframe
pip_main_vars<- left_join(pip_main_vars, count_reporting_levels)

# Filter to see if there are any cases where only non-national estimates are available
pip_main_vars %>%
filter(number_of_reporting_levels<3 & reporting_level!='national') %>%
select(Entity) %>%
unique()

# Keep only national estimates except for these cases

pip_main_vars<- pip_main_vars %>%
filter(reporting_level=='national' | number_of_reporting_levels<3)

# Check again if there is more than one reporting level
pip_main_vars %>%
  select(Entity, Year, reporting_level) %>%
  unique() %>%
  count(Entity, Year) %>%
  rename(number_of_reporting_levels = n) %>%
    filter(number_of_reporting_levels>1)

```

### Prepare reference year data

```{r}
# I'll prepare the observations from the main_vars file, and then merge in the percentiles

target_years<- c(1990, 2000, 2010, 2018)

df_pip_ref_years<- keep_closest(
  df= pip_main_vars,
  id_vars = c("Entity", "reporting_level"),
  distance_varname = "Year",
  targets = target_years,
  tie_break_dist_up_or_down = "up",
  tie_break_varname = "welfare_type",
  tie_break_value = "income")

# We see that if we impose that a max distance, country coverage falls.
df_count<- df_pip_ref_years %>%
  count(Entity)

df_pip_ref_years_filter<- df_pip_ref_years %>%
  filter(abs_distance<3)

df_count<- df_pip_ref_years_filter %>%
  count(Entity)

```