### 2022-10-31 PhD_pages/data_appendices/3_A_prepare_KLEMS.qmd
Adding data appendix for paper 3

---
title: "Appendix 3.A: Preparing EU KLEMS industry data"
format: html
---

```{r}
library(tidyverse)
```

## Introduction

## Raw data files

Make a list of filenames (both to reference the correct file in the url and as the filename to save locally).
```{r}

files<- c("32_countries",
          "Canada",
          "US_alt")

# The urls where the files can be retrieved
download_urls<- c("http://www.euklems.net/data/09i/all_countries_09I.txt",
               "https://www.worldklems.net/data/basic/CAN_WK_07_2012.xlsx",
               "https://www.worldklems.net/data/basic/usa_wk_mar_2017.xlsx"
)

names(download_urls)<- files

# uses the last part of the download urls as the filenames
filenames<- sub(".*\\/", "", download_urls)

names(filenames)<- files

# Add directory to specify filepaths in which to save a local copy of each file
filepaths<- paste0("PhD_pages/data/original/KLEMS/", filenames)

names(filepaths)<- files

```

Save files locally.
```{r}

for (file in files){
  print(file)
download.file(download_urls[file], filepaths[file])
}

```

## Preparing dataframes

Importing the main data file - with data for 32 countries.
```{r}

df <- read.csv(filepaths["32_countries"], quote="")

# convert to long format and rename variables
df<- df %>%
  pivot_longer(cols = starts_with("X_"), names_to = "year",
    names_prefix = "X_", values_to ="value",
    names_transform = list(year=as.integer),
    values_transform = list(value=as.character)) %>%
  rename(country = X.country,
         transaction = var)

# Remove quotations (These come from not properly specifying the file structure in the read.csv command above. But I was unable to find the right format spec.)
df <- df %>% 
  mutate(country = str_replace(country, '"', ''),
         value = as.numeric(str_replace(value, '"', '')))

```

## Calculatiing derived variables