### 2022-10-31 PhD_pages/data_appendices/3_A_prepare_KLEMS.qmd
prep OECD data

---
title: "Appendix 3.A: Preparing EU KLEMS industry data"
format: html
---

```{r}
library(tidyverse)
library(readxl) # For handling xls source data files

```

## Introduction

## Raw data files

Make a list of filenames (both to reference the correct file in the url and as the filename to save locally).
```{r}

files<- c("32_countries",
          "Canada",
          "US_alt",
          "Japan")

# The urls where the files can be retrieved
download_urls<- c("http://www.euklems.net/data/09i/all_countries_09I.txt",
               "https://www.worldklems.net/data/basic/CAN_WK_07_2012.xlsx",
               "https://www.worldklems.net/data/basic/usa_wk_apr_2013.xlsx",
               "https://www.worldklems.net/data/basic/JPN_wk_may_2013.xlsx"
)

names(download_urls)<- files

# uses the last part of the download urls as the filenames
filenames<- sub(".*\\/", "", download_urls)

names(filenames)<- files

# Add directory to specify filepaths in which to save a local copy of each file
filepaths<- paste0("PhD_pages/data/original/KLEMS/", filenames)

names(filepaths)<- files

```

Save files locally.
```{r}

for (file in files){
  print(file)
download.file(download_urls[file], filepaths[file])
}

```

## Preparing dataframes

```{r}
df_list<- list()
```

Importing the main data file - with data for 32 countries.
```{r}

df_list[["32_countries"]]<- read.csv(filepaths["32_countries"], quote="")

# convert to long format and rename variables
df_list[["32_countries"]]<- df_list[["32_countries"]] %>%
  pivot_longer(cols = starts_with("X_"), names_to = "year",
    names_prefix = "X_", values_to ="value",
    names_transform = list(year=as.integer),
    values_transform = list(value=as.character)) %>%
  rename(country = X.country,
         transaction = var)

# Remove quotations (These come from not properly specifying the file structure in the read.csv command above. But I was unable to find the right format spec.)
df_list[["32_countries"]]<- df_list[["32_countries"]] %>%
  mutate(country = str_replace(country, '"', ''),
         value = as.numeric(str_replace(value, '"', '')))

```

Data for Canada, Japan and an alternative (longer) series for the US.
```{r}
countries<- c("Canada", "US_alt", "Japan")

for (country in countries){
df_list[[country]]<- as.data.frame(read_excel(filepaths[country],
                                                         sheet = 'DATA'))

# convert to long format and rename variables
df_list[[country]]<- df_list[[country]] %>%
  pivot_longer(cols = starts_with("_"), names_to = "year",
    names_prefix = "_", values_to ="value",
    names_transform = list(year=as.integer)) %>%
  rename(transaction = Variable) %>%
  mutate(country = country)

}

```

Tidying Canada data
```{r}
# Drop some NA values (- data for 2009 and 2010 is given as missing)
df_list[["Canada"]]<- df_list[["Canada"]] %>%
  filter(!is.na(value))

# For Canada we need to calculate some industry aggregates to be comparable with other countries
agg_industries<- list()

agg_industries[['D']]<- c(
  "15t16",
  "17t19",
  "20",
  "21t22",
  "23",
  "24",
  "25",
  "26",
  "27t28",
  "29",
  "30t33",
  "34t35",
  "36t37")

agg_industries[['G']]<- c(
  "50",
  "52")

agg_industries[['I']]<- c(
  "60t63",
  "64")

agg_industries[['K']]<- c(
  "70",
  "71t74")

for (agg_ind in names(agg_industries)){
  print(agg_ind)

  df_totals<- df_list[["Canada"]] %>%
  filter(code %in% agg_industries[[agg_ind]]) %>%
    group_by(transaction, year) %>%
    summarise(value = sum(value)) %>%
    ungroup() %>%
    mutate(code = agg_ind) %>%
    mutate(country = "Canada")

  df_list[["Canada"]]<- bind_rows(df_list[["Canada"]], df_totals)

}

```
Tidying Japan data
```{r}
# Drop NA values
df_list[["Japan"]]<- df_list[["Japan"]] %>% filter(is.na(value))

# Rename the capital compensation variable, to agree with other files
df_list[["Japan"]]<- df_list[["Japan"]] %>%
  mutate(transaction = replace(transaction,
                               transaction == "CAP_GFCF",
                               "CAP"))

```

Join the different dataframes together into a single dataframe.
```{r}
df<- bind_rows(
  df_list
)

```

Make a dictionary of industry codes and there descriptions from the combined data sources.
```{r}

df_industries<- df %>%
  select(code, desc) %>%
  drop_na() %>%
  unique()

# There are a couple of different spellings etc. from the different datasets. Here we just take select one (arbitrarily)
df_industries<- df_industries %>%
  distinct(code, .keep_all = TRUE)

# Add in the uniform industry code descriptions
df<- df %>%
  select(-desc)

df<- left_join(df, df_industries)

```

Some more tidying
```{r}
df<- df %>%
  filter(transaction %in% c("VA", "CAP", "LAB", "COMP"))

df<- df %>%
  pivot_wider(names_from = "transaction",
              values_from = "value")

# Any industry where VA = CAP but LAB is missing, set LAB = 0
  # This is generally only the case for sector Q, where VA=CAP=0.
df<- df %>%
  mutate(LAB = replace(LAB,
                       VA==CAP & is.na(LAB),
                       0))

# Any industry where VA = LAB but CAP is missing, set CAP = 0
  # This is only for sector P for US (NAICS data) and Korea (and Japan where P VA = 0)
df<- df %>%
  mutate(CAP = replace(CAP,
                       VA==LAB & is.na(CAP),
                       0))

# Drop any country-year-industry obs that do not have data for all 4 transactions
  # To-do check if COMP is needed, and calculate for US_alt and Japan if so)

df<- df %>% filter(!(is.na(VA) & is.na(CAP) & is.na(LAB)) )

```

Drop USA-NAICS data (in favour of of SIC data)
```{r}
df<- df %>%
  filter(country!="USA-NAICS")
```

Standardize country names
```{r}

# Mapping file
fp<- "PhD_pages/data/original/KLEMS/KLEMS_country_name_mapping.csv"

KLEMS_country_names<- read.csv(fp)

KLEMS_country_names<- KLEMS_country_names %>%
  rename(country = Country)

# aggregate shares
df<- merge(df, KLEMS_country_names)

df <- df %>%
  select(-country) %>%
  rename(country = Our.World.In.Data.Name)

```

## Complete dataset for aggregate industries

```{r}

# This list gives the codes of the main industries
main_industry_codes<- c("AtB",
                        "C",
                        "D",
                        "E",
                        "F",
                        "G",
                        "H",
                        "I",
                        "J",
                        "K",
                        "L",
                        "M",
                        "N",
                        "O",
                        "P",
                        "Q")

# filter for these codes sub-industries
industry_level_values<- df %>%
  filter(code %in% main_industry_codes)

# Store reported totals separately

reported_total_values<- df %>%
  filter(code == "TOT") %>%
  select(-code) %>%
  gather(transaction, reported_total, -c(country, year)) %>%
  arrange(country, year, transaction)

```