### 2022-11-30 PhD_pages/data_appendices/3_A_prepare_KLEMS.qmd
Updating papers and data appendices

---
title: "Appendix 3.A: Preparing EU KLEMS industry data"
format: html
---

## Processing steps

The titles in each of the boxes below summarise the steps taken to prepare the KLEMS industry data. In each case you can expand the box to see the code used.

::: {.callout-note collapse="true" appearance="minimal"}
### Load libraries and functions

```{r}
library(tidyverse)
library(readxl) # For handling xls source data files
library(knitr)

source("PhD_pages/scripts/general_functions.R")
```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Specify the location of four original data files

We obtain data from four KLEMS datasets:

-   EU KLEMS version ...
-   Canada...
-   An additional series for the US (different industry classification, longer series..)
-   Japan

Here we specify corresponding lists of URLs where the original files are located, short names for the files, and local filepaths where a copy of the original data files will be saved.

```{r}

files<- c("32_countries",
          "CAN",
          "USA")

# The urls where the files can be retrieved
download_urls<- c("http://www.euklems.net/data/09i/all_countries_09I.txt",
               "https://www.worldklems.net/data/basic/CAN_WK_07_2012.xlsx",
               "https://www.worldklems.net/data/basic/usa_wk_apr_2013.xlsx"
)

# JAPAN OLD: "https://www.worldklems.net/data/basic/JPN_wk_may_2013.xlsx"

names(download_urls)<- files

# uses the last part of the download urls as the filenames
filenames<- sub(".*\\/", "", download_urls)

names(filenames)<- files

# The directory where the copies will be saved
fp<- "PhD_pages/data/original/KLEMS/"

# Append this directory to the file names to specify file paths
filepaths<- paste0(fp, filenames)

names(filepaths)<- files

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Save a local copy of the original data files

If a new copy of the original data is required, the `grab_original` variable can be set to `TRUE`. Otherwise this step is skipped.

```{r}
grab_original<- FALSE

if (grab_original){
  for (file in files){
    print(paste0("Downloading ", file))
    download.file(download_urls[file], filepaths[file])
  }
}

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Load locally saved copies of original data as data frames

```{r}
df_list<- list()
```

Importing the main data file - with data for 32 countries.

```{r}

df_list[["32_countries"]]<- read.csv(filepaths["32_countries"], quote="")

# convert to long format and rename variables
df_list[["32_countries"]]<- df_list[["32_countries"]] %>%
  pivot_longer(cols = starts_with("X_"), names_to = "year",
    names_prefix = "X_", values_to ="value",
    names_transform = list(year=as.integer),
    values_transform = list(value=as.character)) %>%
  rename(country = X.country,
         transaction = var)

# Remove quotations (Ideally this would be handled by specifying the right format in the read.csv command above. But I was unable to find the right format specification)
df_list[["32_countries"]]<- df_list[["32_countries"]] %>%
  mutate(country = str_replace(country, '"', ''),
         value = as.numeric(str_replace(value, '"', '')))

# Recode the Entity name for US: 'USA-NAICS' -> 'USA'
df_list[["32_countries"]]<- df_list[["32_countries"]] %>%
  mutate(country = if_else(country == "USA-NAICS",
                           "USA",
                           country))

```

Data for Canada and the alternative series for the US.

```{r}
countries<- c("CAN", "USA")

for (country in countries){
df_list[[country]]<- as.data.frame(read_excel(filepaths[country],
                                                         sheet = 'DATA'))

# convert to long format and rename variables
df_list[[country]]<- df_list[[country]] %>%
  pivot_longer(cols = starts_with("_"), names_to = "year",
    names_prefix = "_", values_to ="value",
    names_transform = list(year=as.integer)) %>%
  rename(transaction = Variable) %>%
  mutate(country = country)

}

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Tidying Canada data

```{r}
# Drop some NA values (- data for 2009 and 2010 is given as missing)
df_list[["CAN"]]<- df_list[["CAN"]] %>%
  filter(!is.na(value))

# For Canada we need to calculate some industry aggregates to be comparable with other countries
agg_industries<- list()

agg_industries[['D']]<- c(
  "15t16",
  "17t19",
  "20",
  "21t22",
  "23",
  "24",
  "25",
  "26",
  "27t28",
  "29",
  "30t33",
  "34t35",
  "36t37")

agg_industries[['G']]<- c(
  "50",
  "52")

agg_industries[['I']]<- c(
  "60t63",
  "64")

agg_industries[['K']]<- c(
  "70",
  "71t74")

for (agg_ind in names(agg_industries)){
  print(paste0("Adding aggregate data for industry: ", agg_ind))

  df_totals<- df_list[["CAN"]] %>%
  filter(code %in% agg_industries[[agg_ind]]) %>%
    group_by(transaction, year) %>%
    summarise(value = sum(value)) %>%
    ungroup() %>%
    mutate(code = agg_ind) %>%
    mutate(country = "CAN")

  df_list[["CAN"]]<- bind_rows(df_list[["CAN"]], df_totals)

}

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Tidying Japan data - OLD, TO DELETE

```{r}
# Drop NA values
# df_list[["JPN"]]<- df_list[["JPN"]] %>% filter(is.na(value))
# 
# # Rename the capital compensation variable, to agree with other files
# df_list[["JPN"]]<- df_list[["JPN"]] %>%
#   mutate(transaction = replace(transaction,
#                                transaction == "CAP_GFCF",
#                                "CAP"))

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Join the different dataframes together into a single dataframe

```{r}
df<- bind_rows(
  df_list, .id = "source"
)

# Make a copy of the original merged dataframe, for use later on.
df_orig<- df
```

:::

::: {.callout-note collapse="true" appearance="minimal"}
### Drop duplicate series for the US

Data for the US is included in the main data file using the NAICS industry classification.

A separate source provides a much longer series using SIC classification.

Here we show the data for aggregate Value Added from the two datasets.
```{r}

df %>%
  filter(country %in% c("USA"),
         code == "TOT",
 transaction == "VA") %>%
  ggplot(aes(x = year, y = value, shape = source, colour = source, alpha=0.5)) +
  geom_point()

```

We drop the shorter NAICS data in favour of the SIC data.
```{r}
df<- df %>%
  filter(!(country == "USA" & source == "32_countries"))

```

:::

::: {.callout-note collapse="true" appearance="minimal"}
### Standardize country names

Here is the mapping of country names we use to standardize across data sources.
```{r}

KLEMS_country_mapping<- read.csv(paste0(fp, "KLEMS_country_name_mapping", ".csv"))

kable(KLEMS_country_mapping)

```

Here we map the joined KLEMS data sets to these standardized names.
```{r}
df<- map_names(
  df = df, 
  df_mapping = KLEMS_country_mapping, 
  old_country_varname_df = "country", 
  old_country_varname_df_mapping = "Country_KLEMS",
  new_country_varname_df_mapping = "Our_World_In_Data_country_name",
  new_country_varname_final = "country"
)

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Select relevant transactions and convert to wide format

We keep only the transactions we need: Value Added, Capital income, Labour income and Compensation of employees. And we spread the data to a wider format.

```{r}
df<- df %>%
  filter(transaction %in% c("VA", "CAP", "LAB", "COMP"))

df<- df %>%
  pivot_wider(id_cols = c(country, year, code),
              names_from = "transaction",
              values_from = "value")

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Replace some NA values with 0

There are many (`r df %>% filter(VA == CAP & is.na(LAB)) %>% count() %>% as.numeric()`) country-year-industry observations where where CAP is 100% of VA, but LAB is missing.

Most of these (`r df %>% filter(VA == CAP & is.na(LAB) & VA == 0) %>% count() %>% as.numeric()`) are where both capital income and value added equal zero. Such observations appear only for the following industries:

```{r}
kable(
  df %>% filter(VA == CAP & is.na(LAB) & VA == 0) %>% count(code) 
)
```

There are also a small number of such observations where VA>0, listed here:

```{r}
kable(
  df %>% filter(VA == CAP & is.na(LAB) & VA>0)
)

```

For both cases (i.e. where LAB is NA and either VA = CAP = 0 or VA = CAP > 0) we set LAB = 0.

```{r}
df<- df %>%
  mutate(LAB = replace(LAB,
                       VA==CAP & is.na(LAB),
                       0))
```

There is a similar set of country-year-industry observations (n = `r df %>% filter(VA == LAB & is.na(CAP)) %>% count() %>% as.numeric()`) where VA = LAB but CAP is NA. These we list here.

```{r}
kable(
  df %>% filter(VA == LAB & is.na(CAP) & VA>0)
)

```

For these cases (i.e. where CAP is NA and VA = LAB) we set CAP = 0.

```{r}
df<- df %>%
  mutate(CAP = replace(CAP,
                       VA==LAB & is.na(CAP),
                       0))
```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Filter and store data for main industries and the aggregate economy separately

```{r}

# This list gives the codes of the main industries
main_industry_codes<- c("AtB",
                        "C",
                        "D",
                        "E",
                        "F",
                        "G",
                        "H",
                        "I",
                        "J",
                        "K",
                        "L",
                        "M",
                        "N",
                        "O",
                        "P",
                        "Q")

# filter for these codes
main_industry_accounts<- df %>%
  filter(code %in% main_industry_codes)

# Store reported totals separately

reported_totals<- df %>%
  filter(code == "TOT") %>%
  select(-code)

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Drop country-year-industry observations that do not have all variables needed

There are some observations (n=`r main_industry_accounts %>% filter(is.na(VA) & is.na(CAP) & is.na(LAB)) %>% count() %>% as.numeric()`) without .... . We group these by country and industry.

```{r}
kable(
  main_industry_accounts %>% filter(is.na(VA) | is.na(CAP) | is.na(LAB)) %>% count(country, code)
)

```

We filter these out.

```{r}

  # TODO check if COMP is needed, and add in check here and/or calculate it where missing? if so)

main_industry_accounts<- main_industry_accounts %>% filter(!(is.na(VA) & is.na(CAP) & is.na(LAB)) )

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Add industry code descriptions

Here are the industry descriptions provided in the data.
```{r}

industry_descriptions<- df_orig %>%
  filter(code %in% main_industry_codes) %>%
  select(code, desc) %>%
  drop_na() %>%
  unique() %>%
  arrange(code)

kable(industry_descriptions)

```

We join these back into the clean dataframe.
```{r}

main_industry_accounts<- left_join(main_industry_accounts, industry_descriptions)

```
:::

## Sense checking

### How does the sum across all industries compare to the reported economy-wide totals?

::: {.callout-note collapse="true" appearance="minimal"}
### Calculate aggregate across industries and merge with reported economy-wide totals

```{r}
agg_industries<- main_industry_accounts %>%
    group_by(country, year) %>%
    summarise(VA = sum(VA, na.rm = TRUE),
              CAP = sum(CAP, na.rm = TRUE),
              COMP = sum(COMP, na.rm = TRUE),
              LAB = sum(LAB, na.rm = TRUE)
              )

reported_totals <- reported_totals %>%
  rename(VA_reported = VA,
         CAP_reported = CAP,
         COMP_reported = COMP,
         LAB_reported = LAB)

check_totals<- left_join(agg_industries, reported_totals) %>%
  mutate(VA_ratio = VA/VA_reported,
         CAP_ratio = CAP/CAP_reported,
         COMP_ratio = COMP/COMP_reported,
         LAB_ratio = LAB/LAB_reported)

tolerance<- 0.001 # tenth of a percent tolerance

```
:::

In general the sum across industries is very close to to the reported economy-wide aggregate. In the table below we show all observations where the ratio deviates from 100% by more than `r tolerance*100`%.

Note that in Canada - which is one of the countries we will go on to use this data for - the sum across available industries falls short of the reported aggregate (for value added, Capital income, Labour income and Compensation of employees) by around 5%, with no trend over time.

::: {.panel-tabset}
## Value added
```{r}

kable(
  check_totals %>%
  filter(VA_ratio<(1-tolerance) | VA_ratio>(1+tolerance)) %>%
    select(country, year, VA_ratio, VA, VA_reported)
)
```
## Capital income
```{r}
kable(
  check_totals %>%
  filter(CAP_ratio<(1-tolerance) | CAP_ratio>(1+tolerance)) %>%
    select(country, year, CAP_ratio, CAP, CAP_reported)
)

```

## Labour income

```{r}
kable(
  check_totals %>%
  filter(LAB_ratio<(1-tolerance) | LAB_ratio>(1+tolerance)) %>%
    select(country, year, LAB_ratio, LAB, LAB_reported)
)

```

## Comp. of employees
```{r}
kable(
  check_totals %>%
  filter(COMP_ratio<(1-tolerance) | COMP_ratio>(1+tolerance)) %>%
    select(country, year, COMP_ratio, COMP, COMP_reported)
)

```

:::

## Save cleaned data

::: {.callout-note collapse="true" appearance="minimal"}
### Write cleaned data to csv
```{r}
fp<- "PhD_pages/data/clean/"

write.csv(main_industry_accounts, paste0(fp, "KLEMS_gen_income_by_industry", ".csv"))

```

:::