### 2022-11-30 PhD_pages/data_appendices/3_B_prepare_OECD.qmd

---
title: "Appendix 3.B: Preparing OECD industry data"
format: html
---

## Processing steps

The titles in each of the boxes below summarise the steps taken to prepare the OECD industry data. In each case you can expand the box to see the code used.

::: {.callout-note collapse="true" appearance="minimal"}
### Load libraries and functions

```{r}
library(tidyverse)
library(rsdmx)
library(knitr)

source("PhD_pages/scripts/general_functions.R")
```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Load industry generation of income and employment data

```{r}
fp<- "PhD_pages/data/original/OECD/"

load(paste0(fp, "industry_accounts", ".Rda"))
load(paste0(fp, "employment_by_industry", ".Rda"))

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Standardize country names

Here is the mapping of country names we use to standardize across data sources.
```{r}

OECD_country_mapping<- read.csv(paste0(fp, "OECD_country_name_mapping", ".csv"))

kable(OECD_country_mapping)

```

Here we map the two data sets to these standardized names.
```{r}

industry_accounts<- map_names(
  df = industry_accounts, 
  df_mapping = OECD_country_mapping, 
  old_country_varname_df = "LOCATION", 
  old_country_varname_df_mapping = "OECD_country_name",
  new_country_varname_df_mapping = "Our_World_In_Data_country_name",
  new_country_varname_final = "country"
)

employment_by_industry<- map_names(
  df = employment_by_industry, 
  df_mapping = OECD_country_mapping, 
  old_country_varname_df = "LOCATION", 
  old_country_varname_df_mapping = "OECD_country_name",
  new_country_varname_df_mapping = "Our_World_In_Data_country_name",
  new_country_varname_final = "country"
)

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Clean year variable in both files

```{r}
clean_year<- function(df){
  df<- df %>%
    mutate(year = as.numeric(obsTime))
}

industry_accounts<- clean_year(industry_accounts)
employment_by_industry<- clean_year(employment_by_industry)

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Further tidying of industry generation of income data

```{r}

# Filter for only current price data and the needed transactions
industry_accounts<- industry_accounts %>%
  filter(MEASURE == "C" & 
           TRANSACT %in% c("B1GA", "D1A","B2G_B3GA", "B2N_B3NA", "D29_D39A"))

# Recode transactions from national accounts codes to something more legible
recoding_transactions <- c(
  B1GA = "VA", # (Gross) Value added
  D1A = "COMP", # Compensation of employees
  B2G_B3GA = "GOS", # Gross operating surplus
  B2N_B3NA = "NOS", # Net operating surplus (net of depreciation)
  D29_D39A = "TOP") # Taxes on production

industry_accounts<- industry_accounts %>%
  mutate(TRANSACT = recode(TRANSACT, !!!recoding_transactions))

# flip to wide format
industry_accounts<- industry_accounts %>%
  pivot_wider(names_from = TRANSACT, values_from = obsValue)

# Select and rename needed vars
industry_accounts<- industry_accounts %>%
  rename(ind_code = ACTIVITY) %>%
  select(country, year, ind_code, VA, COMP, GOS, NOS, TOP)

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Filter and store generation of income data for main industries and the aggregate economy separately

```{r}

# Main aggregate industries

main_industry_codes<- c("VA0", "VB", "VC", "VD", "VE", "VF", "VG", "VH", "VI",
                        "VJ", "VK", "VL", "VM", "VN", "VO", "VP", "VQ", "VR",
                        "VS", "VT", "VU")

# Filter industry data for main industry codes only
main_industry_accounts<- industry_accounts %>%
  filter(ind_code %in% main_industry_codes)

# Store economy-wide aggregates separately
reported_totals <- industry_accounts %>%
  filter(ind_code == 'VTOT')

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Further tidying of industry employment data.

```{r}
# Select self-employed and employed hours and flip to wide format
employment_by_industry<- employment_by_industry %>%
  filter(TRANSACT %in% c("ESEA", "EEMA")) %>%  # codes for self-employed and employed
  pivot_wider(id_cols = c(country, year, ACTIVITY), names_from = TRANSACT, values_from = obsValue)

# Select and rename needed vars
employment_by_industry<- employment_by_industry %>%
  rename(self_employed_hrs = ESEA, 
         employed_hrs = EEMA,
         ind_code = ACTIVITY)

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Join generation of income and employment data

```{r}
main_industry_accounts<- left_join(main_industry_accounts, employment_by_industry)
```

:::

::: {.callout-note collapse="true" appearance="minimal"}
### Filter out observations where the implied wage is infinite

In the joined data, there are a small number of rows where employed hours = 0 but compensation of employees \> 0 --- implying an infinite wage.

Here we show these illogical rows, grouped by country and industry.

```{r}
kable(
  main_industry_accounts %>%
  filter((COMP>0 & employed_hrs == 0))
)
```

We drop these.

```{r}
main_industry_accounts<- main_industry_accounts %>%
  filter(!(COMP>0 & employed_hrs == 0))

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Filter out country-year-industry observations without all the variables we need

There are some rows where we do not have all the variables we will need for the calculations: VA, COMP, GOS, and employed and self-employed hours.

```{r}
kable(
  main_industry_accounts %>%
    filter(is.na(VA) | is.na(COMP) | is.na(GOS) | is.na(employed_hrs) | is.na(self_employed_hrs)) %>%
    count(country, ind_code)
)
```

We drop these.

```{r}
main_industry_accounts<-  main_industry_accounts %>%
    filter(!(is.na(VA) | is.na(COMP) | is.na(GOS) | is.na(employed_hrs) | is.na(self_employed_hrs)))
```
:::

::: {.callout-note collapse="true" appearance="minimal"}
#### Filter out country-year observations with partial industry coverage

The objective is to have complete coverage of the economy. There are some country years where coverage across industries looks decidedly incomplete.

For instance, out of `r length(main_industry_codes)` aggregate industry codes, these country years have data for fewer than 18:

```{r}
kable( 
 to_drop<- main_industry_accounts %>%
   count(country, year) %>%
   filter(n<18)
)

```

We drop these country-years with apparently incomplete data.

```{r}
main_industry_accounts<- left_join(main_industry_accounts, to_drop) %>%
  filter(is.na(n)) %>%
  select(-n)

```

But because of differences in national accounting methods, missing data for a given industry does not imply the accounts are incomplete.

In particular, the dataset contains only `r main_industry_accounts %>% filter(ind_code == "VU")%>%count()` country-year observations for `VU` (as compared to `r main_industry_accounts %>% filter(ind_code == "VA0")%>%count()` for `VA0`) and these are in general only NA or 0 values.

```{r}
summary(main_industry_accounts %>% filter(ind_code == "VU"))
```

In the table below we show all remaining country years in which at least one industry --- other than `VU` --- is missing.

We assume that these apparent gaps do not indicate incompleteness of the data, rather just differences in national accounting practices.

```{r}

kable( main_industry_accounts %>%
  select(year, country, ind_code, VA) %>%
  pivot_wider(names_from = ind_code, values_from = VA) %>%
  filter(if_any(VA0:VT, ~ is.na(.)))
)

```
:::

## Sense checking

### How does the sum across all industries compare to the reported economy-wide totals?

::: {.callout-note collapse="true" appearance="minimal"}
### Calculate aggregate across industries and merge with reported economy-wide totals

```{r}
agg_industries<- main_industry_accounts %>%
    group_by(country, year) %>%
    summarise(VA = sum(VA, na.rm = TRUE),
              COMP = sum(COMP, na.rm = TRUE),
              GOS = sum(GOS, na.rm = TRUE),
              NOS = sum(NOS, na.rm = TRUE)
              )

reported_totals <- reported_totals %>%
  rename(VA_reported = VA,
         COMP_reported = COMP,
         GOS_reported = GOS,
         NOS_reported = NOS) %>%
  select(-ind_code)

check_totals<- left_join(agg_industries, reported_totals) %>%
  mutate(VA_ratio = VA/VA_reported,
         COMP_ratio = COMP/COMP_reported,
         GOS_ratio = GOS/GOS_reported,
         NOS_ratio = NOS/NOS_reported)

tolerance<- 0.005 # half a percent tolerance

```
:::

In general the sum across industries is very close to to the reported economy-wide aggregate. In the table below we show all observations where the ratio deviates from 100% by more than `r tolerance*100`%.

Note that in Finland - which is one of the countries we will go on to use this data for - the sum across available industries falls short of the reported aggregate (for value added, compensation of employees and gross operating surplus) by around 5%, with no trend over time.

::: {.panel-tabset}
## Value added
```{r}

kable(
  check_totals %>%
  filter(VA_ratio<(1-tolerance) | VA_ratio>(1+tolerance)) %>%
    select(country, year, VA_ratio, VA, VA_reported)
)
```
## Comp. of employees
```{r}
kable(
  check_totals %>%
  filter(COMP_ratio<(1-tolerance) | COMP_ratio>(1+tolerance)) %>%
    select(country, year, COMP_ratio, COMP, COMP_reported)
)

```
## Gross operating surplus
```{r}
kable(
  check_totals %>%
  filter(GOS_ratio<(1-tolerance) | GOS_ratio>(1+tolerance)) %>%
    select(country, year, GOS_ratio, GOS, GOS_reported)
)

```

:::

## Save cleaned data

::: {.callout-note collapse="true" appearance="minimal"}
### Write cleaned data to csv
```{r}
fp<- "PhD_pages/data/clean/"

write.csv(main_industry_accounts, paste0(fp, "OECD_gen_income_and_employment_by_industry", ".csv"))

```
:::