### 2023-01-08 PhD_pages/paper_pages/paper1.qmd
add test shinylive app that pulls in csv

---
title: "Paper 1: Triangulating trends in inequality around the world from secondary databases"
---

::: {.callout-note collapse="true" appearance="minimal"}
### Load packages

```{r}
library(tidyverse)
library(knitr)
library(data.table)

```
:::

What has happened to within-country inequality globally?

Different headlines - WID suggests general rise (find a quote). PIP much more mixed... even a fall.

Which world we live in matters for how we think we should tackle inequality.

A secular, general rise - suggesting global forces, perhaps something intrinsic to capitalism (Piketty, 2014).

A mixed picture of inequality around the world suggests reflect specific circumstances, with accordingly much more scope for national policy to influence.

Both are reasonable methods and have strengths and weaknesses.

I want to:

a\) Triangulate the key stylized facts. Which observations are robust to across available datasets with their different methods, and which are contingent?

b\) Untangle which methods, definitions affect the different conclusions arrived at.

# Methods and data

## Data sources

### PIP

::: {.callout-note collapse="true" appearance="minimal"}
### Load PIP data

```{r}
# df_pip<- read.csv("https://raw.githubusercontent.com/owid/poverty-data/main/datasets/pip_dataset.csv")

df_pip<- read.csv("PhD_pages/data/pip_dataset (2).csv")

df_pip<- df_pip %>%
  filter(ppp_version == 2017) %>%
  rename("Gini" = "gini",
         "Top_10_share" = "decile10_share",
         "Entity" = "country",
         "Year" = "year")

# drop subnational data where national data is available
df_pip<- df_pip %>%
  filter((!Entity %in% c("China", "India", "Indonesia")) | reporting_level == "national")

# drop income if both income and consumption are available
df_pip<- df_pip %>%
  group_by(Entity, reporting_level, Year) %>%
  mutate(n = n()) %>%
  filter(n<2 | welfare_type == "consumption") %>%
  ungroup()

 #drop regional aggregates
drop_aggs<- c(
  "East Asia and Pacific",
  "Europe and Central Asia",
  "High income countries",
  "Latin America and the Caribbean",
  "Middle East and North Africa",
  "Sub-Saharan Africa",
  "World"
)

df_pip<- df_pip %>%
  filter(!Entity %in% drop_aggs)

```
:::

### WID

::: {.callout-note collapse="true" appearance="minimal"}
### Load WID data

```{r}
# df_wid<- read.csv("https://raw.githubusercontent.com/owid/notebooks/main/BetterDataDocs/PabloArriagada/WID/data/final/wid_pretax.csv")

df_wid<- read.csv("PhD_pages/data/wid.csv")

df_wid<- df_wid %>%
  rename("Gini" = "Gini.coefficient",
         "Top_10_share" = "P90.P100...share.of.the.top.10.")

drop_entities<- c(
"China - rural",
"China - urban",
"East Germany",
"Other Russia and Central Asia",
"Other East Asia",
"Other Western Europe",
"Other Latin America",
"Other MENA",
"Other South & South-East Asia",
"Other Sub-Saharan Africa",
"Africa",
"Asia",
"Europe",
"Oceania",
"Central Asia",
"East Africa",
"East Asia",
"Eastern Europe",
"Middle Africa",
"North Africa",
"North America",
"South-East Asia",
"South Africa region",
"West Africa",
"West Asia",
"Western Europe",
"European Union",
"World",
"Asia (excluding Middle East)",
"North America and Oceania",
"Sub-Saharan Africa",
"Latin America",
"Middle East",
"MENA",
"Russia and Central Asia",
"South & South-East Asia"
)

df_wid<- df_wid %>%
  filter(!Entity %in% drop_entities)
```

:::

### LIS

### Other sources

Should we use Solt? As another check on results?

What about GCIP? For trying to flip between income and consumption...

### Coverage

Give an overview of coverage in each dataset.

::: {.callout-note collapse="true" appearance="minimal"}
### Standardize coutry names across datasets
```{r}

pip_countries<- df_pip %>%
  select(Entity) %>%
  unique() %>%
  mutate(Entity_pip = Entity)

wid_countries<- df_wid %>%
  select(Entity) %>%
  unique() %>%
  mutate(Entity_wid = Entity)

countries<- full_join(pip_countries, wid_countries, keep = TRUE) %>%
  select(Entity_pip, Entity_wid)

```

Countries that are in the PIP data but not the WID data:
```{r}
kable(
  countries %>%
       filter(is.na(Entity_wid))
)
```
Countries that are in the WID data but not the PIP data:
(TO-DO: drop regions and sub-national data from the WID data at an eariler step.)
```{r}
kable(
  countries %>%
       filter(is.na(Entity_pip))
)
```
:::

## Comparing databases

#### Method 1: Reference years, with a certain tolerance

::: {.callout-note collapse="true" appearance="minimal"} 
## A function for selecting data around a set of reference years, within a certain tolerance

```{r}
## This has been superseded by the function below
# select_ref_years<- function(df, target_varname, group_by_varname, ref_varname, ref_values,tolerance,tie_break_round_up){
#   
#   ## Example arguments for testing
#   # df_in<- df_wid
#   # target_varname<- "Gini.coefficient"
#   # group_by_varname<- "Entity"
#   # ref_varname<- "Year"
#   # ref_values<- c(1990, 2020)
#   # tolerance<- 5
#   # tie_break_round_up<- TRUE
#   
#     # rename vars toß make nonstandard evaluation easier
#    df_in<- df %>%
#     rename(target_var = !!target_varname,
#            group_by_var = !!group_by_varname,
#            ref_var = !!ref_varname)
#    
#    out_list<- list()
#    
#    for(ref in ref_values){
#      
#      df_ref<- df_in %>%
#        select(target_var, group_by_var, ref_var) %>%
#        drop_na() %>%
#        mutate(distance = ref_var - ref) %>%
#        mutate(abs_distance = abs(distance)) %>%
#        group_by(group_by_var) %>%
#        mutate(min_distance = min(abs_distance)) %>%
#        filter(abs_distance == min_distance) %>%
#        group_by(group_by_var, min_distance) %>%
#        mutate(n = n()) %>%
#        ungroup()
#      
#     # address tie-breaks. If 'round_up' is true, the higher value wins out - i.e. the distance to the reference value must positive. And visa versa.
#      if(tie_break_round_up){
#        df_ref<- df_ref %>%
#          filter(n==1 | distance>0)
#      } else {
#        df_ref<- df_ref %>%
#          filter(n==1 | distance<0)
#      }
#      
#      out_list[[as.character(ref)]]<- df_ref
#      
#      
#    }
#  
#   
#   df_out<- bind_rows(out_list, .id = paste0("reference_", ref_varname))
#   
#   # If a tolerance has been set, filter out matches beyond tolerance
#   if(!is.na(tolerance)){
#       df_out<- df_out %>%
#          filter(abs_distance <= tolerance)
#      }
#   
#   # names back to original and drop not needed vars
#    df_out<- df_out %>%
#     rename(!!target_varname := target_var,
#            !!group_by_varname := group_by_var,
#            !!ref_varname := ref_var) %>%
#      select(-c(distance, min_distance, n))
#    
# }

```

```{r}
# Next, we define a function that takes a dataframe, a pair of years (year1 < year2), a distance, a list of variable names, and a list of ID columns as arguments
# # calculate_change <- function(data, year1, year2, distance, vars, id_cols) {

# #   calc_distances<- list()

# #   for(ref in c(year1, year2)){

# #   # We add a column stating the reference year and group the data by the ID columns
# #   data_subset <- data %>%
# #     mutate(ref_year = ref) %>%
# #     group_by(across(all_of(id_cols)))

# #     # We calculate the difference and absolute difference from the reference year
# #   data_subset$diff <- data_subset$Year - ref
# #   data_subset$abs_diff <- abs(data_subset$diff)

# #   # Store in list
# #   calc_distances[[ref]]<- data_subset %>% ungroup()
# #   }

# #   data_long<- bind_rows(calc_distances)

# #   # A list where we will store the matched observations and changes for each variable
# #   return_matches<- list()

# #   # na_treatment<- "obs_for_all_vars"
# #   vars_plus<- c(vars, "obs_for_all_vars_case")

# #     #For each var...
# #     for(var in vars_plus){

# #       # Filter out observations with missing values in var of interest and calculate the min absolute distance from the two reference years
# #     if(var == "obs_for_all_vars_case"){
# #       matches <- data_long %>%
# #         filter_at(vars,all_vars(!is.na(.)))
# #     } else {
# #       matches <- data_long %>%
# #       filter(!is.na(!!sym(var)))
# #     }

# #     matches <- matches %>%
# #       group_by(across(all_of(c(id_cols, "ref_year")))) %>% 
# #       mutate(min_diff = min(abs_diff)) %>%
# #       filter(abs_diff == min_diff)

# #      # If there is a tie-break (with two matches, one above and one below the reference year), settle the tie-break so as to maximise the gap between the two observations. That means for the early year, below matches win; for the later year, above matches win.

# #     matches <- matches %>%
# #       group_by(across(all_of(c(id_cols, "ref_year", "min_diff")))) %>%
# #       mutate(n = n()) %>% # count number of matches (n>1 if there is a tie-break) %>%
# #     ungroup()

# #     matches <- matches %>% 
# #       group_by(across(all_of(c(id_cols, "ref_year")))) %>% 
# #       filter(n==1 | (ref_year == year1 & diff<0) | (ref_year == year2 & diff>0))

# #     # If the same year is matched for the two reference years, drop the observations
# #      matches <- matches %>% 
# #       group_by(across(all_of(c(id_cols, "Year")))) %>% 
# #       mutate(n = n()) %>% # count number of times a year is matched
# #       filter(!n>1) # Drop if same year is matched

# #     if(var == "obs_for_all_vars_case"){
# #       keep_vars<- c(id_cols, "Year", vars, "ref_year" )
# #     } else {
# #       keep_vars<- c(id_cols, "Year", var, "ref_year" )
# #     }

# #      matches <- matches %>%
# #        select_at(keep_vars)

# #      # Make a list to store the matched observations, and also the calculated distances
# #     return_matches[[var]]<- matches

# #     }

# #   # Finally, we return the ...
# #   return(return_matches)
# # }

# # # To test the function, we can create a sample dataframe
# # data <- data.frame(Year = c(1999, 2000, 2001, 2002, 2003, 2004, 2003),
# #                    Entity = c("A","A", "A", "A", "B", "B", "C"),
# #                    Var1 = c(40, NA, 20, 30, 40, 50, 60),
# #                    Var2 = c(30, 100, 200, 300, 400, 500, 400))

# # year1<- 2000
# # year2<- 2005
# # distance<- 2
# # vars<- c("Var1","Var2")
# # id_cols<- c("Entity")

# # # Then, we can call the function with a pair of years, a distance, a list of variable names, and a list of ID columns
# # calculate_change(data, 2000, 2004, 2, c("Var1", "Var2"), c("Entity"))

# ```

# ```{r}
# compare_ref_years<- function(df_list, target_varname, group_by_varname, ref_varname, ref_values,tolerance,tie_break_round_up, rise_fall_threshold) {

#   # Example arguments for testing
#   df_list<- list("pip" = df_pip, "wid" = df_wid)
#   target_varname<- "Gini"
#   group_by_varname<- "Entity"
#   ref_varname<- "Year"
#   ref_values<- c(1990, 2020)
#   tolerance<- 5
#   tie_break_round_up<- TRUE
#   rise_fall_threshold<- 0.01

#   # Store the names of the two dataframes as a vector
#   df_names<- names(df_list)

#   # Initiate a list to store the processed data for each dataframe
#   df_out_list<- list()

#   # loop through both data frame names
#   for(i in 1:2){
#     df_name<- df_names[i]

#     # rename vars to avoid nonstandard evaluation later on
#     df_in<- df_list[[df_name]] %>%
#       rename(target_var = !!target_varname,
#            group_by_var = !!group_by_varname,
#            ref_var = !!ref_varname)

#     # initiate a list to store the filtered data for each reference (e.g. reference year)
#     ref_list_list<- list()

#     # Loop through the references
#     for(ref in ref_values){

#       # filter for the closest observation, by group (e.g. by country)
#       df_ref<- df_in %>%
#         select(target_var, group_by_var, ref_var) %>%
#         drop_na() %>%
#         mutate(distance = ref_var - ref) %>%
#         mutate(abs_distance = abs(distance)) %>%
#         group_by(group_by_var) %>%
#         mutate(min_distance = min(abs_distance)) %>%
#         filter(abs_distance == min_distance) %>%
#         group_by(group_by_var, min_distance) %>%
#         mutate(n = n()) %>% # count number of matches (n>1 if there is a tie-break)
#         ungroup()

#       # Settle tie-breaks. If 'round_up' is true, the higher value wins out - i.e. the distance to the reference value must positive. And visa versa.
#       if(tie_break_round_up){
#         df_ref<- df_ref %>%
#          filter(n==1 | distance>0)
#       } else {
#         df_ref<- df_ref %>%
#          filter(n==1 | distance<0)
#       }

#       ref_list_list[[as.character(ref)]]<- df_ref

#     }

#     # Collate the matches for each reference observation
#     df_out<- bind_rows(ref_list_list, .id = "reference")

#     # If a tolerance has been set, filter out matches beyond tolerance
#     if(!is.na(tolerance)){
#         df_out<- df_out %>%
#          filter(abs_distance <= tolerance)
#      }

#     # drop not needed vars
#     df_out<- df_out %>%
#       select(-c(distance, min_distance, n))

#     # code rise, fall, stable
#     df_out<- df_out %>%
#       drop_na() %>%
#       pivot_longer(cols = c(target_var, ref_var, abs_distance), names_to="metric", values_to = "value" ) %>%
#       pivot_wider(id_cols = group_by_var, names_from = c(metric,reference), values_from = value) %>%
#   drop_na() %>%
#   mutate(change = !!sym(paste0("target_var_", max(ref_values))) - !!sym(paste0("target_var_", min(ref_values)))) %>%
#   mutate(rise_fall = "stable") %>%
#   mutate(rise_fall = if_else(change< -(rise_fall_threshold),
#                              "fall",
#                              rise_fall)) %>%
#   mutate(rise_fall = if_else(change > rise_fall_threshold,
#                              "rise",
#                              rise_fall))

#     # Store the results in a list
#     df_out_list[[df_name]]<- df_out

#   }

#   # Join the results gathered from each input dataframe
#   df_out_list[["joint"]]<- full_join(
#     df_out_list[[df_names[1]]], df_out_list[[df_names[2]]],
#     by = c("group_by_var"),
#     suffix = c(paste0("_",df_names[1]), paste0("_",df_names[2])),
#     keep = TRUE
#     )

#   # Make a list of fall/rise summaries

#     # Initiate the list
#     rise_fall_list<- list()

#     #Loop through the two sources
#     for(df_name in df_names){

#        # For the individual source observations
#         rise_fall_list[[paste0(df_name, "_allObs")]]<- df_out_list[[df_name]] %>%
#           count(rise_fall)

#         # for the observations that are in both sources
#         rise_fall_list[[paste0(df_name, "_sharedObs")]]<- df_out_list[["joint"]] %>% 
#       drop_na() %>%
#       count(across(paste0("rise_fall_", df_name)))
#     }

#   # Make a list of the averages in the two reference periods

#     # Initiate the list
#     average_list<- list()

#     #Loop through the two sources
#     for(df_name in df_names){

#        # For the individual source observations
#         average_list[[paste0(df_name, "_allObs")]]<- df_out_list[[df_name]] %>%
#           summarise(!!sym(paste0("avg_", ref_values[1])) := mean(!!sym(paste0("target_var_", ref_values[1]))),
#                     !!sym(paste0("avg_", ref_values[2])) := mean(!!sym(paste0("target_var_", ref_values[2]))))

#         # for the observations that are in both sources
#         average_list[[paste0(df_name, "_sharedObs")]]<- df_out_list[["joint"]] %>% 
#       drop_na() %>%
#       summarise(!!sym(paste0("avg_", ref_values[1])) := mean(!!sym(paste("target_var", ref_values[1], df_name, sep= "_"))),
#                     !!sym(paste0("avg_", ref_values[2])) := mean(!!sym(paste("target_var", ref_values[1], df_name, sep= "_"))))

#     }

#   # Undo the renaming (used early to avoid nonstandard evaluation problems)
#   for(df_name in names(df_out_list)){

#   names(df_out_list[[df_name]]) <- gsub(x = names(df_out_list[[df_name]]), pattern = "target_var", replacement = target_varname)

#   names(df_out_list[[df_name]]) <- gsub(x = names(df_out_list[[df_name]]), pattern = "group_by_var", replacement = group_by_varname)

#   names(df_out_list[[df_name]]) <- gsub(x = names(df_out_list[[df_name]]), pattern = "ref_var", replacement = ref_varname)
# }

#  outputs<- list("dataframes" = df_out_list,
#                 "summaries" = summary_list)

# return(outputs)
# }

```

```{r}

# compare_pip_wid<- list()

# for(var in c("Gini", "Top_10_share")){

#   if(var == "Gini"){rise_fall_threshold<- 0.1}
#   if(var == "Top_10_share"){rise_fall_threshold<- 1}

# compare_for_var<- compare_ref_years(
#     df_list = list("pip" = df_pip, "wid" = df_wid),
#     target_varname = var,
#     group_by_varname = "Entity",
#     ref_varname = "Year",
#     ref_values = c(1990, 2020),
#     tolerance = 5,
#     tie_break_round_up = FALSE,
#     rise_fall_threshold = rise_fall_threshold
#     )

# compare_pip_wid[[var]]<- compare_for_var 
# }

```

A scatter chart to compare
```{r}

```

A table to compare rises and falls
```{r}

# tab_compare_rise_and_falls<- function(summaries_list){

#   summaries_list<- compare_pip_wid[["Gini"]][["summaries"]]
#   tab<- data.frame(rise_fall = c("rise", "stable", "fall"))

#   source_names<- names(summaries_list)

#   for(source_name in source_names) {
#   tab<- left_join(tab, summaries_list[[source_name]]) %>%
#     rename(!!(paste0(source_name, "_all")) := n)
#   }

#   # Adding data for shared observations

# }

```

Averages
```{r}

```

:::

#### Method 2: Use a regression to obtain the time trend

### Results for a range of inequality metrics

::: {.callout-note collapse="true" appearance="minimal"} 
## Calculations for comparisons

Calculations for Gini
```{r}

# group_by_varname<- "Entity"
# ref_varname<- "Year"
# ref_values<- c(1990, 2020)
# tolerance<- 5
# tie_break_round_up<- FALSE

# target_varname<- "Gini"

# compare_gini<- compare_pip_and_wid_ref_years(target_varname = "Gini")

# # Spread wide and calculate change over time

# # PIP only
# compare_gini_wide_pip<- compare_gini

# names(compare_gini_wide_pip) <- gsub(x = names(compare_gini_wide_pip), pattern = "_pip", replacement = "")

# compare_gini_wide_pip<- compare_gini_wide_pip %>%
#   select(Entity, reference_Year, Gini) %>%
#   drop_na() %>%
#   pivot_wider(id_cols = Entity, names_from = reference_Year, values_from = Gini) %>%
#   drop_na() %>%
#   mutate(change = `2020` - `1990`) %>%
#   mutate(rise_fall = "stable") %>%
#   mutate(rise_fall = if_else(change< -0.01,
#                              "fall",
#                              rise_fall)) %>%
#   mutate(rise_fall = if_else(change > 0.01,
#                              "rise",
#                              rise_fall))

# compare_gini_pip_summary<- compare_gini_wide_pip %>%
#   count(rise_fall)

# # WID only
# compare_gini_wide_wid<- compare_gini

# names(compare_gini_wide_wid) <- gsub(x = names(compare_gini_wide_wid), pattern = "_wid", replacement = "")

# compare_gini_wide_wid<- compare_gini_wide_wid %>%
#   select(Entity, reference_Year, Gini) %>%
#   drop_na() %>%
#   pivot_wider(id_cols = Entity, names_from = reference_Year, values_from = Gini) %>%
#   drop_na() %>%
#   mutate(change = `2020` - `1990`) %>%
#   mutate(rise_fall = "stable") %>%
#   mutate(rise_fall = if_else(change< -0.01,
#                              "fall",
#                              rise_fall)) %>%
#   mutate(rise_fall = if_else(change > 0.01,
#                              "rise",
#                              rise_fall))

# compare_gini_wid_summary<- compare_gini_wide_wid %>%
#   count(rise_fall)

# # In both PIP and WID

# compare_gini_pip_and_wid<- full_join(compare_gini_wide_pip, compare_gini_wide_wid, by= c("Entity"), keep = TRUE) %>%
#   drop_na() %>%
#   rename(rise_fall_pip = rise_fall.x,
#          rise_fall_wid = rise_fall.y)

# compare_gini_pip_wid_summary_pip<- compare_gini_pip_and_wid %>%
#   count(rise_fall_pip)

# compare_gini_pip_wid_summary_wid<- compare_gini_pip_and_wid %>%
#   count(rise_fall_wid)

# # Make a table

# df_gini_table<- data.frame(
#   rise_fall = c("rise", "fall", "stable")
# )

# df_gini_table<- left_join(df_gini_table, compare_gini_pip_summary) %>%
#   rename(pip_all = n)

# df_gini_table<- left_join(df_gini_table, compare_gini_pip_wid_summary_pip, by = c("rise_fall" = "rise_fall_pip")) %>%
#   rename(pip_shared = n)

# df_gini_table<- left_join(df_gini_table, compare_gini_wid_summary) %>%
#   rename(wid_all = n)

# df_gini_table<- left_join(df_gini_table, compare_gini_pip_wid_summary_wid, by = c("rise_fall" = "rise_fall_wid")) %>%
#   rename(wid_shared = n)

# # country by country table for rise and fall
# df_gini_country_rise_fall_table<- full_join(compare_gini_wide_pip, compare_gini_wide_wid, by= c("Entity"), keep = TRUE) %>%
#   filter(rise_fall.x != rise_fall.y) %>%
#   mutate(Entity = Entity.x) %>%
#   mutate(Entity = if_else(!is.na(Entity.y),
#                           Entity.y,
#                           Entity.x)) %>%
#   select(Entity, rise_fall.x, rise_fall.y) %>%
#   rename(pip = rise_fall.x,
#          wid = rise_fall.y)

```

Calculations for Top 10% share
```{r}

# compare_top_10<- compare_pip_and_wid_ref_years(target_varname = "Top_10_share")

# # Spread wide and calculate change over time

# # PIP only
# compare_top_10_wide_pip<- compare_top_10

# names(compare_top_10_wide_pip) <- gsub(x = names(compare_top_10_wide_pip), pattern = "_pip", replacement = "")

# compare_top_10_wide_pip<- compare_top_10_wide_pip %>%
#   select(Entity, reference_Year, Top_10_share) %>%
#   drop_na() %>%
#   pivot_wider(id_cols = Entity, names_from = reference_Year, values_from = Top_10_share) %>%
#   drop_na() %>%
#   mutate(change = `2020` - `1990`) %>%
#   mutate(rise_fall = "stable") %>%
#   mutate(rise_fall = if_else(change< -1,
#                              "fall",
#                              rise_fall)) %>%
#   mutate(rise_fall = if_else(change > 1,
#                              "rise",
#                              rise_fall))

# compare_top_10_pip_summary<- compare_top_10_wide_pip %>%
#   count(rise_fall)

# # WID only
# compare_top_10_wide_wid<- compare_top_10

# names(compare_top_10_wide_wid) <- gsub(x = names(compare_top_10_wide_wid), pattern = "_wid", replacement = "")

# compare_top_10_wide_wid<- compare_top_10_wide_wid %>%
#   select(Entity, reference_Year, Top_10_share) %>%
#   drop_na() %>%
#   pivot_wider(id_cols = Entity, names_from = reference_Year, values_from = Top_10_share) %>%
#   drop_na() %>%
#   mutate(change = `2020` - `1990`) %>%
#   mutate(rise_fall = "stable") %>%
#   mutate(rise_fall = if_else(change< -1,
#                              "fall",
#                              rise_fall)) %>%
#   mutate(rise_fall = if_else(change > 1,
#                              "rise",
#                              rise_fall))

# compare_top_10_wid_summary<- compare_top_10_wide_wid %>%
#   count(rise_fall)

# # In both PIP and WID

# compare_top_10_pip_and_wid<- full_join(compare_top_10_wide_pip, compare_top_10_wide_wid, by= c("Entity"), keep = TRUE) %>%
#   drop_na() %>%
#   rename(rise_fall_pip = rise_fall.x,
#          rise_fall_wid = rise_fall.y)

# compare_top_10_pip_wid_summary_pip<- compare_top_10_pip_and_wid %>%
#   count(rise_fall_pip)

# compare_top_10_pip_wid_summary_wid<- compare_top_10_pip_and_wid %>%
#   count(rise_fall_wid)

# # Make a table

# df_top_10_table<- data.frame(
#   rise_fall = c("rise", "fall", "stable")
# )

# df_top_10_table<- left_join(df_top_10_table, compare_top_10_pip_summary) %>%
#   rename(pip_all = n)

# df_top_10_table<- left_join(df_top_10_table, compare_top_10_pip_wid_summary_pip, by = c("rise_fall" = "rise_fall_pip")) %>%
#   rename(pip_shared = n)

# df_top_10_table<- left_join(df_top_10_table, compare_top_10_wid_summary) %>%
#   rename(wid_all = n)

# df_top_10_table<- left_join(df_top_10_table, compare_top_10_pip_wid_summary_wid, by = c("rise_fall" = "rise_fall_wid")) %>%
#   rename(wid_shared = n)

# # country by country table for rise and fall
# df_top_10_country_rise_fall_table<- full_join(compare_top_10_wide_pip, compare_top_10_wide_wid, by= c("Entity"), keep = TRUE) %>%
#   filter(rise_fall.x != rise_fall.y) %>%
#   mutate(Entity = Entity.x) %>%
#   mutate(Entity = if_else(!is.na(Entity.y),
#                           Entity.y,
#                           Entity.x)) %>%
#   select(Entity, rise_fall.x, rise_fall.y) %>%
#   rename(pip = rise_fall.x,
#          wid = rise_fall.y)

```

:::

#### Comparison between 1990 and 2020 (+/- 5 years))
::: {.panel-tabset}
### Gini
```{r}
# kable(df_gini_table)

# kable(df_gini_country_rise_fall_table)

```

### Top 10% share
```{r}
# kable(df_top_10_table)

# kable(df_top_10_country_rise_fall_table)

```

### MLD
:::

## Explaining differences between PIP and WID data

### For a subset of countries we can bridge between the PIP and WID data

Besides the underlying data sources, there are many differences between what is being measured in the PIP and WID data - notably the welfare concept and unit of analysis/equivalization.

In order to get a better understanding of where differences between the datasets arise, for a subset of countries we can bridge between the two somewhat. For example:

-   WID has post-tax income series for some countries
-   We can use LIS to look at how changing the welfare concept, unit of analysis, equivalisation affects things. (For equivalisation, we could also possibly use the newly available PIP harmonized microdata).

### More in-depth discussion of particular outliers

-   Russia is a very notable case where the two datasets tell very different stories.

# Conclusion