### 2023-01-08 test_pages/test_with_shinylive.qmd

---
title: Shinylive in Quarto example
format: html
filters:
  - shinylive
---

This is a Shinylive application embedded in a Quarto doc.

These are Shiny for Python applications which run completely in the browser, using Pyodide (Python compiled to WebAssembly).

See the [Shilylive quarto extension](https://github.com/quarto-ext/shinylive) for further documentation.

And the documentation on [Shiny for Python](https://shiny.rstudio.com/py/docs/putting-it-together.html).

```{shinylive-python}
#| standalone: true

from shiny import *

app_ui = ui.page_fluid(
    ui.input_slider("n", "N", 0, 100, 40),
    ui.output_text_verbatim("txt"),
)

def server(input, output, session):
    @output
    @render.text
    def txt():
        return f"The value of n*2 is {input.n() * 2}"

app = App(app_ui, server)

```

Here is a second app I am testing:

```{shinylive-python}
#| standalone: true
#| viewerHeight: 420

from shiny import *
from plotnine import *
from pyodide.http import open_url
import pandas as pd

app_ui = ui.page_fluid(
    ui.output_plot("example_plot"),
)

def server(input, output, session):
    @output
    @render.plot
    def example_plot():

        url = 'https://raw.githubusercontent.com/owid/notebooks/main/BetterDataDocs/JoeHasell/PIP/data/ppp_2017/final/PIP_data_public_download/full_dataset/cons_only/poverty_cons_only.csv'

        df = pd.read_csv(open_url(url))

        plot = (ggplot(df, aes('Year', 'headcount_ratio_365', color = 'Entity'))
        + geom_line())

        # d = {'col1': [1, 2], 'col2': [3, 4]}
        # df = pd.DataFrame(data=d)
        # plot = (ggplot(df, aes('col1', 'col2'))
        # + geom_point())

        return plot

app = App(app_ui, server)

```