### 2023-01-08 PhD_pages/data_appendices/1_B_within_country_trends.qmd

---
title: "Appendix 1.B: Analysing global trends in within-country inequality"
format: html
warning: false
filters:
  - shinylive
---

A key difficulty of comparing trends is that the data is incomplete coverage. The paper uses two methods to calculate and compare aggregate within-country trends across incomplete data: using reference years and using year fixed effects in a regression.

## Method 1: Reference years

I need to allow for the income/consumption issue (currently the data includes both – i.e. multi observations per country year).

## Method 2: regression analysis

## Explore the date

I plan to build a Shiny app to help compare trends across datasets (using Shinylive – built on Shiny for Python - ).

Here is a test app just so I can test the wiring of how such an app works.

```{shinylive-python}
#| standalone: true
#| viewerHeight: 420

from shiny import *
from plotnine import *
from pyodide.http import open_url
import pandas as pd

app_ui = ui.page_fluid(
    ui.output_plot("example_plot"),
)

def server(input, output, session):
    @output
    @render.plot
    def example_plot():

        url = 'https://raw.githubusercontent.com/owid/notebooks/main/BetterDataDocs/JoeHasell/PIP/data/ppp_2017/final/PIP_data_public_download/full_dataset/cons_only/poverty_cons_only.csv'

        df = pd.read_csv(open_url(url))

        plot = (ggplot(df, aes('Year', 'headcount_ratio_365', color = 'Entity'))
        + geom_line())

        # d = {'col1': [1, 2], 'col2': [3, 4]}
        # df = pd.DataFrame(data=d)
        # plot = (ggplot(df, aes('col1', 'col2'))
        # + geom_point())

        return plot

app = App(app_ui, server)

```