### 2023-01-20 presentations/DSPI_CSP_masters_Jan2023.qmd

---
title: "From $1.90 to $2.15: the updated International Poverty Line"
author: "Joe Hasell"
format: revealjs
---

## The International Poverty Line {.smaller}
:::: {.columns}

::: {.column width="30%"}
Left column

- some point
- another point
:::

::: {.column width="70%"}
```{=html}
<iframe src="https://ourworldindata.org/grapher/share-of-population-in-extreme-poverty?country=BGD~BOL~MDG~IND~CHN~ETH" loading="lazy" style="width: 100%; height: 600px; border: 0px none;"></iframe>
```

:::

::::

## The IPL has been updated from $1.90 to $2.15

## Basic story {.smaller}

:::: {.columns}

::: {.column width="30%"}

- An adjustment for inflation
- $1 in 2017 prices < $1 in 2011 prices
- The value of the IPL has, broadly speaking, been maintined
- Global extreme poverty very similar measured in either terms.
:::

::: {.column width="70%"}

```{=html}
<iframe src="https://ourworldindata.org/explorers/poverty-explorer-2011-vs-2017-ppp?time=2013..latest&hideControls=true&Metric=Share+in+poverty&International-%24=Compare+2017+and+2011+prices&Poverty+line=International+Poverty+Line&Household+survey+data+type=Show+data+from+both+income+and+expenditure+surveys&country=~OWID_WRL" loading="lazy" style="width: 100%; height: 600px; border: 0px none;"></iframe>
```

:::

::::