### 2023-01-29 PhD_pages/data_appendices/1_C_prepare_LIS.qmd

---
title: "Appendix 1.C: Preparing LIS inequality data"
format: html
warning: false

---

```{python}
# For some reason I can't get the owid package working when rendering... only when running in the interactive window in vs code
# from owid import catalog
import pandas as pd

pd.set_option('display.max_rows', None)
```

We load the data from a file prepared from the WID Stata package by Our World in Data.

```{r}

from owid import catalog

# results = catalog.find('population')
# owid_pop = results[results['path'] == 'garden/owid/latest/key_indicators/population']

# df_pop = owid_pop.iloc[0].load().reset_index().rename(columns={'year':'Year', 'country': 'Entity'})

# df_pop.head()

# df_wid<- read.csv("https://raw.githubusercontent.com/owid/notebooks/main/BetterDataDocs/PabloArriagada/WID/data/final/wid_pretax.csv")

# A local version, in case internet isn't available 
# df_wid<- read.csv("data/original/wid.csv")

```