### 2023-01-29 PhD_pages/data_appendices/1_D_prepare_population.qmd

---
title: "Appendix 1.D: Preparing population data"
format: html
warning: false

---

::: {.callout-note collapse="true" appearance="minimal"}
### Packages and settings
```{python}
from owid import catalog
import pandas as pd

pd.set_option('display.max_rows', None)
```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Load population data from Our World in Data

I grab the main population series prepared by Our World in Data.
```{python}

results = catalog.find('population')
owid_pop = results[results['path'] == 'garden/owid/latest/key_indicators/population']

df_pop = owid_pop.iloc[0].load().reset_index().rename(columns={'year':'Year', 'country': 'Entity'})

df_pop.head()

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Check there is population data for each country in WID, PIP, and LIS.

All countries in the PIP data are matched:
```{python}

df_pip = pd.read_csv("data/clean/pip.csv")

check_pip = pd.merge(df_pip, df_pop, how = 'left')
check_pip[check_pip['population'].isna()]

```

There is no population data for Zanzibar, present in the WID data.

```{python}
df_wid = pd.read_csv("data/clean/wid.csv")

check_wid = pd.merge(df_wid, df_pop, how = 'left')
check_wid[check_wid['population'].isna()]

```

:::

::: {.callout-note collapse="true" appearance="minimal"}
### Add in Zanzibar population

WID release their own population data, but this currently insn't gathered in our data pipeline. For now I make up a population series for Zanzibar based on two data points from Wikipedia. At some point I will come back and replaces this with the WID population data.

```{python}
zanz_pop_1978 = 476111
zanz_pop_2022 = 1889773

# Assume a constant growth rate
zanz_g = pow(zanz_pop_2022/zanz_pop_1978, 1/(2022-1978))-1

year = 2022
pop = zanz_pop_1978 * pow(1+zanz_g, year - 1978)

df_zanz_pop = df_wid[df_wid['Entity'] == "Zanzibar"].copy()
df_zanz_pop['population'] = zanz_pop_1978 * pow(1+zanz_g, df_zanz_pop['Year'] - 1978)

df_zanz_pop = df_zanz_pop[['Entity', 'Year', 'population']]

df_pop = pd.concat([df_pop, df_zanz_pop])

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### See the prepared data

The first 100 rows of the dataframe looks like this:
```{python}
df_pop.head(100)
```

:::

::: {.callout-note collapse="true" appearance="minimal"}
### Save
```{python}
df_pop.to_csv('data/clean/population.csv')

```

:::