### 2023-02-01 PhD_pages/data_appendices/1_F_within_country_trends.qmd
Tidying names and layout

---
title: "Appendix 1.F: Analysing global trends in within-country inequality"
format: html
warning: false
filters:
  - shinylive

---
```{python}

from plotnine import *
import pandas as pd
import numpy as np

```

This first block of code.....

```{python}

#  ---------------------------------
#  ---------- SECTION 1: Data prep functions -------------
#  ----------------------------------

#  Function get matching for ref years
def closest_to_reference(df, reference_val, max_dist_from_ref, reference_col, group_by_col, value_col, tie_break):

  df = df.loc[:, [reference_col, group_by_col, value_col]]

  # Drop NAs
  df = df.dropna()

  # Calculate absolute distance from reference value
  df['ref_diff'] = abs(df[reference_col] - reference_val)

  # Drop any rows with a distance beyond threshold
  if not pd.isna(max_dist_from_ref):
    df = df.loc[df['ref_diff'] <= max_dist_from_ref]

  # Keep closest observation to reference value – including tie-breaks (where there is a match above and below the ref value)
  df = df[df.groupby(group_by_col)['ref_diff'].transform('min') == df['ref_diff']].reset_index(drop=True)

  # Settle tie-breaks
  if tie_break == 'below':
    df = df[df.groupby(group_by_col)[reference_col].transform('min') == df[reference_col]].reset_index(drop=True)

  elif tie_break == 'above':
    df = df[df.groupby(group_by_col)[reference_col].transform('max') == df[reference_col]].reset_index(drop=True)

  df = df.drop('ref_diff', axis=1)

  df = df\
    .rename(columns={value_col: "value"})

  return df

# Merge matches for different reference points
def merge_two_ref_matches(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):

# Make sure the pair of reference values are in ascending order
  reference_vals.sort()

# Maximise distance between two refs by settling tie-breaks below the lowest ref and above the highest ref

# Find matches for lower reference value
  lower_ref_matches = closest_to_reference(df, reference_vals[0], max_dist_from_refs, reference_col, group_by_col, value_col, 'below')

# Find matches for higher reference value
  higher_ref_matches = closest_to_reference(df, reference_vals[1], max_dist_from_refs, reference_col, group_by_col, value_col, 'above')

# Merge the two sets of matches
  merged_df = pd.merge(lower_ref_matches, higher_ref_matches, on=group_by_col, suffixes=(reference_vals[0], reference_vals[1]))

# Drop obs that do not have data for both ref values
  merged_df = merged_df.dropna()

# Drop obs where the matched data does not meet the min distance requirement
  if not pd.isna(min_dist_between):

  # Store the names of the reference column returned from the two matches
    ref_var_high = f'{reference_col}{reference_vals[1]}'
    ref_var_low = f'{reference_col}{reference_vals[0]}'

  # Keep only rows >= to the min distance
    merged_df = merged_df.loc[(merged_df[ref_var_high] - merged_df[ref_var_low]) >= min_dist_between, :]

  return merged_df

  # For PIP run this three times – first filtering data for just consumpion only, then with income only, then with a dataset that prefers income over consumption
def pip_welfare_routine(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):

    # Specify the name of the column in which the income/consumption welfare definition is stored
    welfare_colname = 'welfare_type'

    # Creat dataframes for thee scenarios:
    # Scenario 1: only allow income data
    df_inc_filter = df.loc[df[welfare_colname] == "income", :]
    df_inc_filter.name = "Income"

    # Scenario 2: only allow consumption data
    df_cons_filter = df.loc[df[welfare_colname] == "consumption", :]
    df_cons_filter.name = "Consumption"
    # Scenario 3: allow a mix – dropping consumption data where income data is available in the same year
    df_mixed = df.copy()

    df_mixed['welfare_count'] = df_mixed.groupby([reference_col, group_by_col])[welfare_colname].transform('count')

    df_mixed = df_mixed.loc[(df_mixed['welfare_count'] == 1) | (df_mixed[welfare_colname] == "income")]

    df_mixed.name = "Mixed"
    #  Store the scneario dataframes in a list
    df_scenarios = [df_inc_filter, df_cons_filter, df_mixed]

    # Run the matching function on each scenario
    scenario_matches = [merge_two_ref_matches(
      df_scenario, 
      reference_vals, 
      max_dist_from_refs, 
      min_dist_between, 
      reference_col, 
      group_by_col, 
      value_col) for df_scenario in df_scenarios]

    # Combine the first two scenarios.
    df_combined_matches = pd.concat([scenario_matches[0], scenario_matches[1]], keys=[df_scenarios[0].name, df_scenarios[1].name])

    # Tidy up indexes
    df_combined_matches = df_combined_matches.reset_index()

    df_combined_matches = df_combined_matches.drop('level_1', axis=1)

    df_combined_matches = df_combined_matches\
      .rename(columns={"level_0": "pip_welfare"})

    # Add in third scenario.
    df_combined_matches = pd.concat([df_combined_matches, scenario_matches[2]])

    # add scenario name to te pip_welfare column
    df_combined_matches['pip_welfare'] = df_combined_matches['pip_welfare'].fillna(df_scenarios[2].name)

    # Keep only one match per group (e.g. per Country) - in the priority laid out in the df_scenarios list above (income only -> consumption only -> mixed)
      # First count the matches
    df_combined_matches['match_count'] = df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
      # Then drop any matches from the lowest priority where there are multiple matches
    df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[2].name)]
      #  Repeat at the next level of priority
    df_combined_matches['match_count'] = df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
    df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[1].name)]

    # Drop the match count column
    df_combined_matches = df_combined_matches.drop('match_count', axis=1)

    return df_combined_matches

```

This second block of code we read in the data (this needs to be done in a different way in the shiny app...)

```{python}

#  ---------------------------------
#  ---------- SECTION 2: Read in data -------------
#  ----------------------------------

url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/pip.csv'

df_pip = pd.read_csv(url)

url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/wid.csv'

df_wid = pd.read_csv(url)

url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/lis.csv'

df_lis = pd.read_csv(url)

url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/region_mapping.csv'

df_regions = pd.read_csv(url)

url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/population.csv'

df_pop = pd.read_csv(url)

```

In a third block, we set specifications for the data matching. This is different in the app because some parameters are controlled by the app.

```{python}
#  ---------------------------------
#  ---------- SECTION 3: Specifications -------------
#  ----------------------------------

# These parameters are not controlled in the app.
reference_col = 'Year'
group_by_col = 'Entity'

tolerance = 1
outlier_cut_off_upper = None

source_count_requirement = "all"
# source_count_requirement = "any"

reference_vals = [1990,2018]
max_dist_from_refs = 4
min_dist_between = 5

data = {"PIP": df_pip,
        "WID": df_wid,
        "LIS": df_lis}

data_to_match = {
  "sources": ["PIP", "WID", "LIS"],
  "var_names": ["Top_10_share", "Top_10_share_pretax", "Top_10_share_equiv_disposable"]
  }

```

In this code block we run the code and prepare final plots and tables.

```{python}

#  ---------------------------------
#  ---------- SECTION 4: Run data functions and produce plot -------------
#  ----------------------------------

# Combine the source and varname strings element-wise to produce keys that will be used for the concat later
source_var_keys = [i + j for i, j in zip(data_to_match["sources"], data_to_match["var_names"])]

def prep_and_merge_multiple_vars(data_to_match, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col):

  matches = []

  for i in range(len(data_to_match['sources'])):

    source = data_to_match["sources"][i]
    var = data_to_match["var_names"][i]
    print(i)
    print(source)
    print(var)
    df = data[source]

    if source == "PIP":

      matched = pip_welfare_routine(
          df = df,
          reference_vals = reference_vals,
          max_dist_from_refs = max_dist_from_refs,
          min_dist_between = min_dist_between,
          reference_col = reference_col,
          group_by_col = group_by_col,
          value_col = var
        )

    else:

      matched = merge_two_ref_matches(
          df = df,
          reference_vals = reference_vals,
          max_dist_from_refs = max_dist_from_refs,
          min_dist_between = min_dist_between,
          reference_col = reference_col,
          group_by_col = group_by_col,
          value_col = var
        )

    matches.append(matched)

  ref_matches = pd.concat(matches, keys=source_var_keys)

  # Tidy up indexes
  ref_matches = ref_matches.reset_index()

  ref_matches = ref_matches.drop('level_1', axis=1)

  ref_matches = ref_matches\
    .rename(columns={"level_0": "source_var"})

  return ref_matches

def prep_plot_and_tables(
  data_to_match,
  reference_vals, 
  max_dist_from_refs,
  min_dist_between,
  reference_col,
  group_by_col,
  tolerance,
  outlier_cut_off_upper,
  source_count_requirement):

  # Other spec:
  region_col = 'region_alt'

  # ---- Prep data ---------

  plot_data = prep_and_merge_multiple_vars(
    data_to_match = data_to_match,
    reference_vals = reference_vals, 
    max_dist_from_refs = max_dist_from_refs,
    min_dist_between = min_dist_between,
    reference_col = reference_col,
    group_by_col = group_by_col
  )

  # Store the names of the columns to be used onthe X and Y axis
  x_axis = f'value{reference_vals[0]}'
  y_axis = f'value{reference_vals[1]}'

  x_ref_val = f'{reference_col}{reference_vals[0]}'
  y_ref_val = f'{reference_col}{reference_vals[1]}'

  # Apply outlier cut off, if specified
  if not pd.isna(outlier_cut_off_upper):
    plot_data = plot_data.loc[plot_data[x_axis] <= outlier_cut_off_upper]
    plot_data = plot_data.loc[plot_data[y_axis] <= outlier_cut_off_upper]

  # Add a count by country – counting whether data is available from each source or not
  plot_data['source_count'] = plot_data.groupby(group_by_col)['source_var'].transform('count')

  # Apply source requirement (whether to include only observations with data from all sources)
  if source_count_requirement == "all":
    plot_data = plot_data.loc[plot_data['source_count'] == plot_data['source_count'].max()]

  # Drop source_count column
  plot_data = plot_data.drop('source_count', axis=1)

  # Add in region classification
  plot_data = pd.merge(plot_data, df_regions, how = 'left')

  # Add in population data
    # For first ref
  df_pop_ref = df_pop.loc[df_pop['Year'] == reference_vals[0], ['Entity', 'population'] ]

  plot_data = pd.merge(plot_data, df_pop_ref, how = 'left')

  plot_data = plot_data.rename(columns={'population':'population_ref1'})

    # For second ref
  df_pop_ref = df_pop.loc[df_pop['Year'] == reference_vals[1], ['Entity', 'population'] ]

  plot_data = pd.merge(plot_data, df_pop_ref, how = 'left')

  plot_data = plot_data.rename(columns={'population':'population_ref2'})

    # Calculate average population two reference periods
  plot_data['avg_pop'] = (plot_data['population_ref1'] + plot_data['population_ref2'])/2

  # ----- Prep plot ------

  # I grab what I am guessing to be the max and min tick marks on the x axist, in order to define the coordinates of a 'tolerance' ribbon.
  x_min = plot_data[x_axis].min()
  x_min_floor = np.floor(x_min * 10) / 10

  x_max = plot_data[x_axis].max()
  x_max_ceiling = np.ceil(x_max * 10) / 10

  # Set the coordinates of the shaded ribbon
  shaded_coord = pd.DataFrame({'x': [x_min_floor, x_max_ceiling, x_max_ceiling, x_min_floor], 
    'y': [x_min_floor-tolerance, x_max_ceiling-tolerance, x_max_ceiling+tolerance , x_min_floor+tolerance]})

  plot = (ggplot(plot_data
  , aes(x_axis, y_axis, alpha=0.5))
  + geom_point(aes(color=region_col, size='avg_pop', alpha = 0.6))
  + facet_wrap('~ source_var')
  + geom_polygon(aes(x='x', y='y', alpha = 0.2), data=shaded_coord)
  + theme_light()
  + theme(legend_position='none')
  )

  # ---- Prep summary table ----

  # Calculate the change between the two ref periods
  plot_data['change'] = (plot_data[y_axis] - plot_data[x_axis])

  group_yes_vars = ['source_var', region_col]
  group_no_vars = ['source_var']

  group_scenarios = [group_no_vars, group_yes_vars]

  aggs = []

  for group_vars in group_scenarios:

  # Prepare pop-weighted average change
    agg_level = plot_data.copy()

    # Calulate pop weights by source and region
    agg_level['regional_pop_weights'] = agg_level.groupby(group_vars)['avg_pop'].transform(lambda x: x/x.sum())

    agg_level['global_pop_weights'] = agg_level.groupby('source_var')['avg_pop'].transform(lambda x: x/x.sum())

    # Multiply the change by the pop weights 
    agg_level['region_weighted_change'] = agg_level['change'] * agg_level['regional_pop_weights']

  # Prepare fall/stable/rise categories
    agg_level['fall'] = agg_level['change'] < -tolerance

    agg_level['stable'] = (agg_level['change'] <= tolerance) & (agg_level['change'] >= -tolerance)

    agg_level['rise'] = agg_level['change'] > tolerance

    agg_level[['fall', 'stable', 'rise']] = agg_level[['fall', 'stable', 'rise']].astype(int)

    agg_level = agg_level.groupby(group_vars)[['change', 'region_weighted_change', 'global_pop_weights', 'fall', 'stable', 'rise']]\
      .agg(['sum', 'mean', 'count'])

  # Aggregate by source and region
    if group_vars == group_yes_vars:

      agg_level = agg_level\
        .unstack().T.reset_index()

    elif group_vars == group_no_vars:

      agg_level = agg_level\
        .T.reset_index()

      agg_level[region_col] = "World"

    aggs.append(agg_level)

  df_summary = pd.concat(aggs)

    # filter for the aggregations we need and label in the 'summary' column
  df_summary.loc[(df_summary['level_0'] == 'change') & (df_summary['level_1'] == 'mean'), 'summary'] = 'avg change'

  df_summary.loc[(df_summary['level_0'] == 'region_weighted_change') & (df_summary['level_1'] == 'sum'), 'summary'] = 'Regional pop-weighted avg change'

  df_summary.loc[(df_summary['level_0'] == 'global_pop_weights') & (df_summary['level_1'] == 'sum'), 'summary'] = 'Global pop weights'

  df_summary.loc[(df_summary['level_0'] == 'change') & (df_summary['level_1'] == 'count'), 'summary'] = 'n'

  df_summary.loc[(df_summary['level_0'] == 'fall') & (df_summary['level_1'] == 'sum'), 'summary'] = 'fall'

  df_summary.loc[(df_summary['level_0'] == 'stable') & (df_summary['level_1'] == 'sum'), 'summary'] = 'stable'

  df_summary.loc[(df_summary['level_0'] == 'rise') & (df_summary['level_1'] == 'sum'), 'summary'] = 'rise'

  df_summary = df_summary[df_summary['summary'].notnull()]

  df_summary = df_summary.drop(['level_0', 'level_1'], axis = 1)

  df_summary['summary'] = pd.Categorical(df_summary['summary'], categories=['fall', 'stable', 'rise','n', 'avg change', 'Regional pop-weighted avg change', 'Global pop weights'], ordered=True)

  df_summary = df_summary.sort_values([region_col,'summary']).set_index([region_col,'summary'])

  # format to show appropriate d.p. I run a loop over each row for want of figuring out a better way!
  repeat_length = 7
  for i in range(0, len(df_summary.index.levels[0].unique())):
    df_summary.iloc[i*repeat_length,:] = df_summary.iloc[i*repeat_length,:].map('{:.0f}'.format)
    df_summary.iloc[1 + i*repeat_length,:] = df_summary.iloc[1 + i*repeat_length,:].map('{:.0f}'.format)
    df_summary.iloc[2 + i*repeat_length,:] = df_summary.iloc[2 + i*repeat_length,:].map('{:.0f}'.format)
    df_summary.iloc[3 + i*repeat_length,:] = df_summary.iloc[3 + i*repeat_length,:].map('{:.0f}'.format)
    df_summary.iloc[4 + i*repeat_length,:] = df_summary.iloc[4 + i*repeat_length,:].map('{:.2f}'.format)
    df_summary.iloc[5 + i*repeat_length,:] = df_summary.iloc[5 + i*repeat_length,:].map('{:.2f}'.format)
    df_summary.iloc[6 + i*repeat_length,:] = df_summary.iloc[6 + i*repeat_length,:].map('{:.2f}'.format)

  # ----- Return outputs ------
  plot_and_tables = {
    "table_data": plot_data,
    "plot": plot,
    "summary": df_summary
  }

  return plot_and_tables

```

```{python}
plot_and_tables = prep_plot_and_tables(
    data_to_match = data_to_match,
    reference_vals = reference_vals, 
    max_dist_from_refs = max_dist_from_refs,
    min_dist_between = min_dist_between,
    reference_col = reference_col,
    group_by_col = group_by_col,
    tolerance = tolerance,
    outlier_cut_off_upper = outlier_cut_off_upper,
    source_count_requirement = source_count_requirement
    )

plot = plot_and_tables['plot']
summary = plot_and_tables['summary']

```

```{python}
plot

```

```{python}
summary

```

## Explore the data

I plan to build a Shiny app to help compare trends across datasets (using Shinylive – built on Shiny for Python - ).

Here is a test app just so I can test the wiring of how such an app works.

```{shinylive-python}
#| standalone: true
#| viewerHeight: 1000

from shiny import *
from plotnine import *
from pyodide.http import open_url
import pandas as pd
import numpy as np

#  ---------------------------------
#  ---------- SECTION 1: Data prep functions -------------
#  ----------------------------------

#  Function get matching for ref years
def closest_to_reference(df, reference_val, max_dist_from_ref, reference_col, group_by_col, value_col, tie_break):

  df = df.loc[:, [reference_col, group_by_col, value_col]]

  # Drop NAs
  df = df.dropna()

  # Calculate absolute distance from reference value
  df['ref_diff'] = abs(df[reference_col] - reference_val)

  # Drop any rows with a distance beyond threshold
  if not pd.isna(max_dist_from_ref):
    df = df.loc[df['ref_diff'] <= max_dist_from_ref]

  # Keep closest observation to reference value – including tie-breaks (where there is a match above and below the ref value)
  df = df[df.groupby(group_by_col)['ref_diff'].transform('min') == df['ref_diff']].reset_index(drop=True)

  # Settle tie-breaks
  if tie_break == 'below':
    df = df[df.groupby(group_by_col)[reference_col].transform('min') == df[reference_col]].reset_index(drop=True)

  elif tie_break == 'above':
    df = df[df.groupby(group_by_col)[reference_col].transform('max') == df[reference_col]].reset_index(drop=True)

  df = df.drop('ref_diff', axis=1)

  df = df\
    .rename(columns={value_col: "value"})

  return df

# Merge matches for different reference points
def merge_two_ref_matches(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):

# Make sure the pair of reference values are in ascending order
  reference_vals.sort()

# Maximise distance between two refs by settling tie-breaks below the lowest ref and above the highest ref

# Find matches for lower reference value
  lower_ref_matches = closest_to_reference(df, reference_vals[0], max_dist_from_refs, reference_col, group_by_col, value_col, 'below')

# Find matches for higher reference value
  higher_ref_matches = closest_to_reference(df, reference_vals[1], max_dist_from_refs, reference_col, group_by_col, value_col, 'above')

# Merge the two sets of matches
  merged_df = pd.merge(lower_ref_matches, higher_ref_matches, on=group_by_col, suffixes=(reference_vals[0], reference_vals[1]))

# Drop obs that do not have data for both ref values
  merged_df = merged_df.dropna()

# Drop obs where the matched data does not meet the min distance requirement
  if not pd.isna(min_dist_between):

  # Store the names of the reference column returned from the two matches
    ref_var_high = f'{reference_col}{reference_vals[1]}'
    ref_var_low = f'{reference_col}{reference_vals[0]}'

  # Keep only rows >= to the min distance
    merged_df = merged_df.loc[(merged_df[ref_var_high] - merged_df[ref_var_low]) >= min_dist_between, :]

  return merged_df

  # For PIP run this three times – first filtering data for just consumpion only, then with income only, then with a dataset that prefers income over consumption
def pip_welfare_routine(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):

    # Specify the name of the column in which the income/consumption welfare definition is stored
    welfare_colname = 'welfare_type'

    # Creat dataframes for thee scenarios:
    # Scenario 1: only allow income data
    df_inc_filter = df.loc[df[welfare_colname] == "income", :]
    df_inc_filter.name = "Income"

    # Scenario 2: only allow consumption data
    df_cons_filter = df.loc[df[welfare_colname] == "consumption", :]
    df_cons_filter.name = "Consumption"
    # Scenario 3: allow a mix – dropping consumption data where income data is available in the same year
    df_mixed = df.copy()

    df_mixed['welfare_count'] = df_mixed.groupby([reference_col, group_by_col])[welfare_colname].transform('count')

    df_mixed = df_mixed.loc[(df_mixed['welfare_count'] == 1) | (df_mixed[welfare_colname] == "income")]

    df_mixed.name = "Mixed"
    #  Store the scneario dataframes in a list
    df_scenarios = [df_inc_filter, df_cons_filter, df_mixed]

    # Run the matching function on each scenario
    scenario_matches = [merge_two_ref_matches(
      df_scenario, 
      reference_vals, 
      max_dist_from_refs, 
      min_dist_between, 
      reference_col, 
      group_by_col, 
      value_col) for df_scenario in df_scenarios]

    # Combine the first two scenarios.
    df_combined_matches = pd.concat([scenario_matches[0], scenario_matches[1]], keys=[df_scenarios[0].name, df_scenarios[1].name])

    # Tidy up indexes
    df_combined_matches = df_combined_matches.reset_index()

    df_combined_matches = df_combined_matches.drop('level_1', axis=1)

    df_combined_matches = df_combined_matches\
      .rename(columns={"level_0": "pip_welfare"})

    # Add in third scenario.
    df_combined_matches = pd.concat([df_combined_matches, scenario_matches[2]])

    # add scenario name to te pip_welfare column
    df_combined_matches['pip_welfare'] = df_combined_matches['pip_welfare'].fillna(df_scenarios[2].name)

    # Keep only one match per group (e.g. per Country) - in the priority laid out in the df_scenarios list above (income only -> consumption only -> mixed)
      # First count the matches
    df_combined_matches['match_count'] = df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
      # Then drop any matches from the lowest priority where there are multiple matches
    df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[2].name)]
      #  Repeat at the next level of priority
    df_combined_matches['match_count'] = df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
    df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[1].name)]

    # Drop the match count column
    df_combined_matches = df_combined_matches.drop('match_count', axis=1)

    return df_combined_matches

#  ---------------------------------
#  ---------- SECTION 2: Read in data -------------
#  ----------------------------------

url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/pip.csv'

df_pip = pd.read_csv(open_url(url))

url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/wid.csv'

df_wid = pd.read_csv(open_url(url))

url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/lis.csv'

df_lis = pd.read_csv(open_url(url))

url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/region_mapping.csv'

df_regions = pd.read_csv(open_url(url))

url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/population.csv'

df_pop = pd.read_csv(open_url(url))

#  ---------------------------------
#  ---------- SECTION 3: Specifications -------------
#  ----------------------------------

# These parameters are not controlled in the app.
reference_col = 'Year'
group_by_col = 'Entity'

tolerance = 1
outlier_cut_off_upper = None

data = {"PIP": df_pip,
        "WID": df_wid,
        "LIS": df_lis}

# make an empty list to add all the varnames
all_vars = list()

# A string to use to separate the source name from the var name
sep = ": "

# for each df in 'data'
for k,v in data.items():

  # make a list of varnames
  vars = list(v.columns)

  # remove the id cols
  vars.remove(reference_col)
  vars.remove(group_by_col)

  # append the name of the source (the key of 'data') at the front
  vars = [f'{k}{sep}' + var for var in vars]

  # append to the running list
  all_vars = all_vars + vars

#  ---------------------------------
#  ---------- App code -------------
#  ----------------------------------

def panel_box(*args, **kwargs):
    return ui.div(
        ui.div(*args, class_="card-body"),
        **kwargs,
        class_="card mb-3",
    )

app_ui = ui.page_fluid(
    {"class": "p-4"},
    ui.row(
        ui.column(
            4,
            panel_box(
                ui.input_slider("reference_vals", "Reference years", 1975, 2020, value=[1990,2018]),
                ui.input_numeric(
                            "max_dist_from_refs",
                            "=/- (years)",
                            4,
                        ),
                ui.input_numeric(
                            "min_dist_between",
                            "Min distance between observations (years)",
                            23,
                        ),
                ui.input_selectize(
                  "metrics", "Metrics", all_vars, multiple=True
                  ),
                ui.input_checkbox("all_metrics_requirement", "Show only countries with data for all metrics", True),
                ui.input_action_button(
                    "download", "Download data and image", class_="btn-primary w-100"
                ),
            ),
        ),
        ui.column(
            8,
            ui.output_plot("example_plot")
            ),
    ),
)

def server(input, output, session):
    @output
    @render.plot
    def example_plot():

      #  ---------------------------------
      #  ---------- SECTION 4: Run data functions and produce plot -------------
      #  ----------------------------------

      def prep_and_merge_multiple_vars(data_to_match, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col):

        # Combine the source and varname strings element-wise to produce keys that will be used for the concat later
        source_var_keys = [i + j for i, j in zip(data_to_match["sources"], data_to_match["var_names"])]

        matches = []

        for i in range(len(data_to_match['sources'])):

          source = data_to_match["sources"][i]
          var = data_to_match["var_names"][i]

          df = data[source]

          if source == "PIP":

            matched = pip_welfare_routine(
                df = df,
                reference_vals = reference_vals,
                max_dist_from_refs = max_dist_from_refs,
                min_dist_between = min_dist_between,
                reference_col = reference_col,
                group_by_col = group_by_col,
                value_col = var
              )

          else:

            matched = merge_two_ref_matches(
                df = df,
                reference_vals = reference_vals,
                max_dist_from_refs = max_dist_from_refs,
                min_dist_between = min_dist_between,
                reference_col = reference_col,
                group_by_col = group_by_col,
                value_col = var
              )

          matches.append(matched)

        ref_matches = pd.concat(matches, keys=source_var_keys)

        # Tidy up indexes
        ref_matches = ref_matches.reset_index()

        ref_matches = ref_matches.drop('level_1', axis=1)

        ref_matches = ref_matches\
          .rename(columns={"level_0": "source_var"})

        return ref_matches

      def prep_plot_and_tables(
        data_to_match,
        reference_vals, 
        max_dist_from_refs,
        min_dist_between,
        reference_col,
        group_by_col,
        tolerance,
        outlier_cut_off_upper,
        all_metrics_requirement):

        # Other spec:
        region_col = 'region_alt'

        # ---- Prep data ---------

        plot_data = prep_and_merge_multiple_vars(
          data_to_match = data_to_match,
          reference_vals = reference_vals, 
          max_dist_from_refs = max_dist_from_refs,
          min_dist_between = min_dist_between,
          reference_col = reference_col,
          group_by_col = group_by_col
        )

        # Store the names of the columns to be used onthe X and Y axis
        x_axis = f'value{reference_vals[0]}'
        y_axis = f'value{reference_vals[1]}'

        x_ref_val = f'{reference_col}{reference_vals[0]}'
        y_ref_val = f'{reference_col}{reference_vals[1]}'

        # Apply outlier cut off, if specified
        if not pd.isna(outlier_cut_off_upper):
          plot_data = plot_data.loc[plot_data[x_axis] <= outlier_cut_off_upper]
          plot_data = plot_data.loc[plot_data[y_axis] <= outlier_cut_off_upper]

        # Add a count by country – counting whether data is available from each source or not
        plot_data['source_count'] = plot_data.groupby(group_by_col)['source_var'].transform('count')

        # Apply metrics requirement (whether to include only observations with data from all metrics)
        if all_metrics_requirement:
          plot_data = plot_data.loc[plot_data['source_count'] == plot_data['source_count'].max()]

        # Drop source_count column
        plot_data = plot_data.drop('source_count', axis=1)

        # Add in region classification
        plot_data = pd.merge(plot_data, df_regions, how = 'left')

        # Add in population data
          # For first ref
        df_pop_ref = df_pop.loc[df_pop['Year'] == reference_vals[0], ['Entity', 'population'] ]

        plot_data = pd.merge(plot_data, df_pop_ref, how = 'left')

        plot_data = plot_data.rename(columns={'population':'population_ref1'})

          # For second ref
        df_pop_ref = df_pop.loc[df_pop['Year'] == reference_vals[1], ['Entity', 'population'] ]

        plot_data = pd.merge(plot_data, df_pop_ref, how = 'left')

        plot_data = plot_data.rename(columns={'population':'population_ref2'})

          # Calculate average population two reference periods
        plot_data['avg_pop'] = (plot_data['population_ref1'] + plot_data['population_ref2'])/2

        # ----- Prep plot ------

        # I grab what I am guessing to be the max and min tick marks on the x axist, in order to define the coordinates of a 'tolerance' ribbon.
        x_min = plot_data[x_axis].min()
        x_min_floor = np.floor(x_min * 10) / 10

        x_max = plot_data[x_axis].max()
        x_max_ceiling = np.ceil(x_max * 10) / 10

        # Set the coordinates of the shaded ribbon
        shaded_coord = pd.DataFrame({'x': [x_min_floor, x_max_ceiling, x_max_ceiling, x_min_floor], 
          'y': [x_min_floor-tolerance, x_max_ceiling-tolerance, x_max_ceiling+tolerance , x_min_floor+tolerance]})

        plot = (ggplot(plot_data
        , aes(x_axis, y_axis, alpha=0.5))
        + geom_point(aes(color=region_col, size='avg_pop', alpha = 0.6))
        + facet_wrap('~ source_var')
        + geom_polygon(aes(x='x', y='y', alpha = 0.2), data=shaded_coord)
        + theme_light()
        + theme(legend_position='none')
        )

        # ---- Prep summary table ----

        # Calculate the change between the two ref periods
        plot_data['change'] = (plot_data[y_axis] - plot_data[x_axis])

        group_yes_vars = ['source_var', region_col]
        group_no_vars = ['source_var']

        group_scenarios = [group_no_vars, group_yes_vars]

        aggs = []

        for group_vars in group_scenarios:

        # Prepare pop-weighted average change
          agg_level = plot_data.copy()

          # Calulate pop weights by source and region
          agg_level['regional_pop_weights'] = agg_level.groupby(group_vars)['avg_pop'].transform(lambda x: x/x.sum())

          agg_level['global_pop_weights'] = agg_level.groupby('source_var')['avg_pop'].transform(lambda x: x/x.sum())

          # Multiply the change by the pop weights 
          agg_level['region_weighted_change'] = agg_level['change'] * agg_level['regional_pop_weights']

        # Prepare fall/stable/rise categories
          agg_level['fall'] = agg_level['change'] < -tolerance

          agg_level['stable'] = (agg_level['change'] <= tolerance) & (agg_level['change'] >= -tolerance)

          agg_level['rise'] = agg_level['change'] > tolerance

          agg_level[['fall', 'stable', 'rise']] = agg_level[['fall', 'stable', 'rise']].astype(int)

          agg_level = agg_level.groupby(group_vars)[['change', 'region_weighted_change', 'global_pop_weights', 'fall', 'stable', 'rise']]\
            .agg(['sum', 'mean', 'count'])

        # Aggregate by source and region
          if group_vars == group_yes_vars:

            agg_level = agg_level\
              .unstack().T.reset_index()

          elif group_vars == group_no_vars:

            agg_level = agg_level\
              .T.reset_index()

            agg_level[region_col] = "World"

          aggs.append(agg_level)

        df_summary = pd.concat(aggs)

          # filter for the aggregations we need and label in the 'summary' column
        df_summary.loc[(df_summary['level_0'] == 'change') & (df_summary['level_1'] == 'mean'), 'summary'] = 'avg change'

        df_summary.loc[(df_summary['level_0'] == 'region_weighted_change') & (df_summary['level_1'] == 'sum'), 'summary'] = 'Regional pop-weighted avg change'

        df_summary.loc[(df_summary['level_0'] == 'global_pop_weights') & (df_summary['level_1'] == 'sum'), 'summary'] = 'Global pop weights'

        df_summary.loc[(df_summary['level_0'] == 'change') & (df_summary['level_1'] == 'count'), 'summary'] = 'n'

        df_summary.loc[(df_summary['level_0'] == 'fall') & (df_summary['level_1'] == 'sum'), 'summary'] = 'fall'

        df_summary.loc[(df_summary['level_0'] == 'stable') & (df_summary['level_1'] == 'sum'), 'summary'] = 'stable'

        df_summary.loc[(df_summary['level_0'] == 'rise') & (df_summary['level_1'] == 'sum'), 'summary'] = 'rise'

        df_summary = df_summary[df_summary['summary'].notnull()]

        df_summary = df_summary.drop(['level_0', 'level_1'], axis = 1)

        df_summary['summary'] = pd.Categorical(df_summary['summary'], categories=['fall', 'stable', 'rise','n', 'avg change', 'Regional pop-weighted avg change', 'Global pop weights'], ordered=True)

        df_summary = df_summary.sort_values([region_col,'summary']).set_index([region_col,'summary'])

        # format to show appropriate d.p. I run a loop over each row for want of figuring out a better way!
        repeat_length = 7
        for i in range(0, len(df_summary.index.levels[0].unique())):
          df_summary.iloc[i*repeat_length,:] = df_summary.iloc[i*repeat_length,:].map('{:.0f}'.format)
          df_summary.iloc[1 + i*repeat_length,:] = df_summary.iloc[1 + i*repeat_length,:].map('{:.0f}'.format)
          df_summary.iloc[2 + i*repeat_length,:] = df_summary.iloc[2 + i*repeat_length,:].map('{:.0f}'.format)
          df_summary.iloc[3 + i*repeat_length,:] = df_summary.iloc[3 + i*repeat_length,:].map('{:.0f}'.format)
          df_summary.iloc[4 + i*repeat_length,:] = df_summary.iloc[4 + i*repeat_length,:].map('{:.2f}'.format)
          df_summary.iloc[5 + i*repeat_length,:] = df_summary.iloc[5 + i*repeat_length,:].map('{:.2f}'.format)
          df_summary.iloc[6 + i*repeat_length,:] = df_summary.iloc[6 + i*repeat_length,:].map('{:.2f}'.format)

        # ----- Return outputs ------
        plot_and_tables = {
          "table_data": plot_data,
          "plot": plot,
          "summary": df_summary
        }

        return plot_and_tables

      def prep_selected_vars(selected_vars):

        # selected_vars = [
        #   'LIS: Top_10_share_equiv_market',
        #   'WID: Bottom_50_share_posttax_dis',
        #   'PIP: Top_10_share']

        selected_sources = [var.split(sep, 1)[0] for var in selected_vars]
        selected_var_names = [var.split(sep, 1)[1] for var in selected_vars]

        match_data = {
          "sources": selected_sources,
          "var_names": selected_var_names
          }

        return match_data

      data_to_match = prep_selected_vars(input.metrics())

      plot_and_tables = prep_plot_and_tables(
        data_to_match = data_to_match,
        reference_vals = list(input.reference_vals()), 
        max_dist_from_refs = input.max_dist_from_refs(),
        min_dist_between = input.min_dist_between(),
        reference_col = reference_col,
        group_by_col = group_by_col,
        tolerance = tolerance,
        outlier_cut_off_upper = outlier_cut_off_upper,
        all_metrics_requirement = input.all_metrics_requirement()
        )

      plot = plot_and_tables['plot']

      return plot

app = App(app_ui, server)

```