### 2023-02-09 PhD_pages/data_appendices/1_F_within_country_trends.qmd
Paper 1: reorganise app and initial drafting of paper

---
title: "Appendix 1.F: Analysing global trends in within-country inequality"
format: html
warning: false
filters:
  - shinylive
sidebar: false

---

:::{.column-page}

```{shinylive-python}
#| standalone: true
#| viewerHeight: 1000

## file: app.py

import jinja2
from shiny import *
import pandas as pd
from pyodide.http import open_url
import numpy as np
from plotnine import *

#  ---------------------------------
#  ---------- SECTION 1: Data prep functions -------------
#  ----------------------------------

#  Function get matching for ref years
def closest_to_reference(df, reference_val, max_dist_from_ref, reference_col, group_by_col, value_col, tie_break):

  df = df.loc[:, [reference_col, group_by_col, value_col]]

  # Drop NAs
  df = df.dropna()

  # Calculate absolute distance from reference value
  df['ref_diff'] = abs(df[reference_col] - reference_val)

  # Drop any rows with a distance beyond threshold
  if not pd.isna(max_dist_from_ref):
    df = df.loc[df['ref_diff'] <= max_dist_from_ref]

  # Keep closest observation to reference value – including tie-breaks (where there is a match above and below the ref value)
  df = df[df.groupby(group_by_col)['ref_diff'].transform('min') == df['ref_diff']].reset_index(drop=True)

  # Settle tie-breaks
  if tie_break == 'below':
    df = df[df.groupby(group_by_col)[reference_col].transform('min') == df[reference_col]].reset_index(drop=True)

  elif tie_break == 'above':
    df = df[df.groupby(group_by_col)[reference_col].transform('max') == df[reference_col]].reset_index(drop=True)

  df = df.drop('ref_diff', axis=1)

  df = df\
    .rename(columns={value_col: "value"})

  return df

# Merge matches for different reference points
def merge_two_ref_matches(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):

# Make sure the pair of reference values are in ascending order
  reference_vals.sort()

# Maximise distance between two refs by settling tie-breaks below the lowest ref and above the highest ref

# Find matches for lower reference value
  lower_ref_matches = closest_to_reference(df, reference_vals[0], max_dist_from_refs, reference_col, group_by_col, value_col, 'below')

# Find matches for higher reference value
  higher_ref_matches = closest_to_reference(df, reference_vals[1], max_dist_from_refs, reference_col, group_by_col, value_col, 'above')

# Merge the two sets of matches
  merged_df = pd.merge(lower_ref_matches, higher_ref_matches, on=group_by_col, suffixes=(reference_vals[0], reference_vals[1]))

# Drop obs that do not have data for both ref values
  merged_df = merged_df.dropna()

# Drop obs where the matched data does not meet the min distance requirement
  if not pd.isna(min_dist_between):

  # Store the names of the reference column returned from the two matches
    ref_var_high = f'{reference_col}{reference_vals[1]}'
    ref_var_low = f'{reference_col}{reference_vals[0]}'

  # Keep only rows >= to the min distance
    merged_df = merged_df.loc[(merged_df[ref_var_high] - merged_df[ref_var_low]) >= min_dist_between, :]

  return merged_df

  # For PIP run this three times – first filtering data for just consumpion only, then with income only, then with a dataset that prefers income over consumption
def pip_welfare_routine(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):

  # Specify the name of the column in which the income/consumption welfare definition is stored
  welfare_colname = 'welfare_type'

  # Creat dataframes for thee scenarios:
  # Scenario 1: only allow income data
  df_inc_filter = df.loc[df[welfare_colname] == "income", :]
  df_inc_filter.name = "Income"

  # Scenario 2: only allow consumption data
  df_cons_filter = df.loc[df[welfare_colname] == "consumption", :]
  df_cons_filter.name = "Consumption"
  # Scenario 3: allow a mix – dropping consumption data where income data is available in the same year
  df_mixed = df.copy()

  df_mixed['welfare_count'] = df_mixed.groupby([reference_col, group_by_col])[welfare_colname].transform('count')

  df_mixed = df_mixed.loc[(df_mixed['welfare_count'] == 1) | (df_mixed[welfare_colname] == "income")]

  df_mixed.name = "Mixed"
  #  Store the scneario dataframes in a list
  df_scenarios = [df_inc_filter, df_cons_filter, df_mixed]

  # Run the matching function on each scenario
  scenario_matches = [merge_two_ref_matches(
    df_scenario, 
    reference_vals, 
    max_dist_from_refs, 
    min_dist_between, 
    reference_col, 
    group_by_col, 
    value_col) for df_scenario in df_scenarios]

  # Combine the first two scenarios.
  df_combined_matches = pd.concat([scenario_matches[0], scenario_matches[1]], keys=[df_scenarios[0].name, df_scenarios[1].name])

  # Tidy up indexes
  df_combined_matches = df_combined_matches.reset_index()

  df_combined_matches = df_combined_matches.drop('level_1', axis=1)

  df_combined_matches = df_combined_matches\
    .rename(columns={"level_0": "pip_welfare"})

  # Add in third scenario.
  df_combined_matches = pd.concat([df_combined_matches, scenario_matches[2]])

  # add scenario name to te pip_welfare column
  df_combined_matches['pip_welfare'] = df_combined_matches['pip_welfare'].fillna(df_scenarios[2].name)

  # Keep only one match per group (e.g. per Country) - in the priority laid out in the df_scenarios list above (income only -> consumption only -> mixed)
    # First count the matches
  df_combined_matches['match_count'] = df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
    # Then drop any matches from the lowest priority where there are multiple matches
  df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[2].name)]
    #  Repeat at the next level of priority
  df_combined_matches['match_count'] = df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
  df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[1].name)]

  # Drop the match count column
  df_combined_matches = df_combined_matches.drop('match_count', axis=1)

  return df_combined_matches

def prep_selected_vars(selected_vars):

  # selected_vars = [
  #   'LIS: Top_10_share_equiv_market',
  #   'WID: Bottom_50_share_posttax_dis',
  #   'PIP: Top_10_share']

  selected_sources = [var.split(sep, 1)[0] for var in selected_vars]
  selected_var_names = [var.split(sep, 1)[1] for var in selected_vars]

  match_data = {
    "sources": selected_sources,
    "var_names": selected_var_names
          }

  return match_data

def prep_data(
      data_to_match,
      reference_vals,
      max_dist_from_refs,
      min_dist_between,
      all_metrics_requirement,
      reference_col,
      group_by_col,
      tolerance,
      outlier_cut_off_upper):

  # Combine the source and varname strings element-wise to produce keys that will be used for the concat later
  source_var_keys = [i + j for i, j in zip(data_to_match["sources"], data_to_match["var_names"])]

  matches = []

  for i in range(len(data_to_match['sources'])):

    source = data_to_match["sources"][i]
    var = data_to_match["var_names"][i]

    df = data[source]

    if source == "PIP":

      matched = pip_welfare_routine(
          df = df,
          reference_vals = reference_vals,
          max_dist_from_refs = max_dist_from_refs,
          min_dist_between = min_dist_between,
          reference_col = reference_col,
          group_by_col = group_by_col,
          value_col = var
        )

    else:

      matched = merge_two_ref_matches(
          df = df,
          reference_vals = reference_vals,
          max_dist_from_refs = max_dist_from_refs,
          min_dist_between = min_dist_between,
          reference_col = reference_col,
          group_by_col = group_by_col,
          value_col = var
        )

    matches.append(matched)

  plot_data = pd.concat(matches, keys=source_var_keys)

  # Tidy up indexes
  plot_data = plot_data.reset_index()

  plot_data = plot_data.drop('level_1', axis=1)

  plot_data = plot_data\
    .rename(columns={"level_0": "source_var"})

  # Store the names of the columns to be used onthe X and Y axis
  x_axis = f'value{reference_vals[0]}'
  y_axis = f'value{reference_vals[1]}'

  x_ref_val = f'{reference_col}{reference_vals[0]}'
  y_ref_val = f'{reference_col}{reference_vals[1]}'

  # Apply outlier cut off, if specified
  if not pd.isna(outlier_cut_off_upper):
    plot_data = plot_data.loc[plot_data[x_axis] <= outlier_cut_off_upper]
    plot_data = plot_data.loc[plot_data[y_axis] <= outlier_cut_off_upper]

  # Add a count by country – counting whether data is available from each source or not
  plot_data['source_count'] = plot_data.groupby(group_by_col)['source_var'].transform('count')

  # Apply metrics requirement (whether to include only observations with data from all metrics)
  if all_metrics_requirement:
    plot_data = plot_data.loc[plot_data['source_count'] == plot_data['source_count'].max()]

  # Drop source_count column
  plot_data = plot_data.drop('source_count', axis=1)

  # Add in region classification
  plot_data = pd.merge(plot_data, df_regions, how = 'left')

  # Add in population data
    # For first ref
  df_pop_ref = df_pop.loc[df_pop['Year'] == reference_vals[0], ['Entity', 'population'] ]

  plot_data = pd.merge(plot_data, df_pop_ref, how = 'left')

  plot_data = plot_data.rename(columns={'population':'population_ref1'})

    # For second ref
  df_pop_ref = df_pop.loc[df_pop['Year'] == reference_vals[1], ['Entity', 'population'] ]

  plot_data = pd.merge(plot_data, df_pop_ref, how = 'left')

  plot_data = plot_data.rename(columns={'population':'population_ref2'})

    # Calculate average population two reference periods
  plot_data['avg_pop'] = (plot_data['population_ref1'] + plot_data['population_ref2'])/2

  #  PREP SUMMARY TABLE

  # Calculate the change between the two ref periods
  plot_data['change'] = (plot_data[y_axis] - plot_data[x_axis])

  group_yes_vars = ['source_var', region_col]
  group_no_vars = ['source_var']

  group_scenarios = [group_no_vars, group_yes_vars]

  aggs = []

  for group_vars in group_scenarios:

  # Prepare pop-weighted average change
    agg_level = plot_data.copy()

    # Calulate pop weights by source and region
    agg_level['regional_pop_weights'] = agg_level.groupby(group_vars)['avg_pop'].transform(lambda x: x/x.sum())

    agg_level['global_pop_weights'] = agg_level.groupby('source_var')['avg_pop'].transform(lambda x: x/x.sum())

    # Multiply the change by the pop weights 
    agg_level['region_weighted_change'] = agg_level['change'] * agg_level['regional_pop_weights']

  # Prepare fall/stable/rise categories
    agg_level['fall'] = agg_level['change'] < -tolerance

    agg_level['stable'] = (agg_level['change'] <= tolerance) & (agg_level['change'] >= -tolerance)

    agg_level['rise'] = agg_level['change'] > tolerance

    agg_level[['fall', 'stable', 'rise']] = agg_level[['fall', 'stable', 'rise']].astype(int)

    agg_level = agg_level.groupby(group_vars)[['change', 'region_weighted_change', 'global_pop_weights', 'fall', 'stable', 'rise']]\
      .agg(['sum', 'mean', 'count'])

  # Aggregate by source and region
    if group_vars == group_yes_vars:

      agg_level = agg_level\
        .unstack().T.reset_index()

    elif group_vars == group_no_vars:

      agg_level = agg_level\
        .T.reset_index()

      agg_level[region_col] = "World"

    aggs.append(agg_level)

  df_summary = pd.concat(aggs)

    # filter for the aggregations we need and label in the 'summary' column
  df_summary.loc[(df_summary['level_0'] == 'change') & (df_summary['level_1'] == 'mean'), 'summary'] = 'avg change'

  df_summary.loc[(df_summary['level_0'] == 'region_weighted_change') & (df_summary['level_1'] == 'sum'), 'summary'] = 'Regional pop-weighted avg change'

  df_summary.loc[(df_summary['level_0'] == 'global_pop_weights') & (df_summary['level_1'] == 'sum'), 'summary'] = 'Global pop weights'

  df_summary.loc[(df_summary['level_0'] == 'change') & (df_summary['level_1'] == 'count'), 'summary'] = 'n'

  df_summary.loc[(df_summary['level_0'] == 'fall') & (df_summary['level_1'] == 'sum'), 'summary'] = 'fall'

  df_summary.loc[(df_summary['level_0'] == 'stable') & (df_summary['level_1'] == 'sum'), 'summary'] = 'stable'

  df_summary.loc[(df_summary['level_0'] == 'rise') & (df_summary['level_1'] == 'sum'), 'summary'] = 'rise'

  df_summary = df_summary[df_summary['summary'].notnull()]

  df_summary = df_summary.drop(['level_0', 'level_1'], axis = 1)

  df_summary['summary'] = pd.Categorical(df_summary['summary'], categories=['fall', 'stable', 'rise','n', 'avg change', 'Regional pop-weighted avg change', 'Global pop weights'], ordered=True)

  df_summary = df_summary.sort_values([region_col,'summary']).set_index([region_col,'summary'])

  # format to show appropriate d.p. I run a loop over each row for want of figuring out a better way!
  repeat_length = 7
  for i in range(0, len(df_summary.index.levels[0].unique())):
    df_summary.iloc[i*repeat_length,:] = df_summary.iloc[i*repeat_length,:].map('{:.0f}'.format)
    df_summary.iloc[1 + i*repeat_length,:] = df_summary.iloc[1 + i*repeat_length,:].map('{:.0f}'.format)
    df_summary.iloc[2 + i*repeat_length,:] = df_summary.iloc[2 + i*repeat_length,:].map('{:.0f}'.format)
    df_summary.iloc[3 + i*repeat_length,:] = df_summary.iloc[3 + i*repeat_length,:].map('{:.0f}'.format)
    df_summary.iloc[4 + i*repeat_length,:] = df_summary.iloc[4 + i*repeat_length,:].map('{:.2f}'.format)
    df_summary.iloc[5 + i*repeat_length,:] = df_summary.iloc[5 + i*repeat_length,:].map('{:.2f}'.format)
    df_summary.iloc[6 + i*repeat_length,:] = df_summary.iloc[6 + i*repeat_length,:].map('{:.2f}'.format)

  return plot_data, df_summary

#  ---------------------------------
#  ---------- SECTION 2: Specifications -------------
#  ----------------------------------
# These parameters are not controlled in the app.

reference_col = 'Year'
group_by_col = 'Entity'

tolerance = 1
outlier_cut_off_upper = None
region_col = 'region_alt'

#  ---------------------------------
#  ---------- SECTION 3: Read in data -------------
#  ----------------------------------
# Note when testing outside of shinylive, this needs to be adjusted to a normal pd_read, without the pyodide step.

url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/pip.csv'

df_pip = pd.read_csv(open_url(url))

url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/wid.csv'

df_wid = pd.read_csv(open_url(url))

url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/lis.csv'

df_lis = pd.read_csv(open_url(url))

url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/region_mapping.csv'

df_regions = pd.read_csv(open_url(url))

url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/population.csv'

df_pop = pd.read_csv(open_url(url))

#  ---------------------------------
#  ---------- SECTION 4: Gather data into dictionary and collete varnames -------------
#  ----------------------------------

data = {"PIP": df_pip,
        "WID": df_wid,
        "LIS": df_lis}

# make an empty list to add all the varnames
all_vars = list()

# A string to use to separate the source name from the var name
sep = ": "

# for each df in 'data'
for k,v in data.items():

  # make a list of varnames
  vars = list(v.columns)

  # remove the id cols
  vars.remove(reference_col)
  vars.remove(group_by_col)

  # append the name of the source (the key of 'data') at the front
  vars = [f'{k}{sep}' + var for var in vars]

  # append to the running list
  all_vars = all_vars + vars

#  ---------------------------------
#  ---------------------------------

#  ---------------------------------
#  ---------- App code -------------
#  ----------------------------------

def panel_box(*args, **kwargs):
    return ui.div(
        ui.div(*args, class_="card-body"),
        **kwargs,
        class_="card mb-3",
    )

app_ui = ui.page_fluid(
    {"class": "p-4"},
    ui.row(
        ui.column(
            4,
            panel_box(
                ui.input_slider("reference_vals", "Reference years", 1975, 2020, value=[1990,2018]),
                ui.input_numeric(
                            "max_dist_from_refs",
                            "=/- (years)",
                            4,
                        ),
                ui.input_numeric(
                            "min_dist_between",
                            "Min distance between observations (years)",
                            23,
                        ),
                ui.input_selectize(
                  "metrics", "Metrics", all_vars, multiple=True
                  ),
                ui.input_checkbox("all_metrics_requirement", "Show only countries with data for all metrics", True),
                ui.input_action_button(
                    "download", "Download data and image", class_="btn-primary w-100"
                ),
            ),
        ),
        ui.column(
            8,
            ui.output_plot("plot"),
            ui.output_table("table")
            ),
    ),
)

def server(input, output, session):

    @reactive.Calc
    def reactive_data():

      data_to_match = prep_selected_vars(input.metrics())

      reference_vals = list(input.reference_vals())

      max_dist_from_refs = input.max_dist_from_refs()
      min_dist_between = input.min_dist_between()
      all_metrics_requirement = input.all_metrics_requirement()

      prepped_data = prep_data(
        data_to_match = data_to_match,
        reference_vals = reference_vals,
        max_dist_from_refs = max_dist_from_refs,
        min_dist_between = min_dist_between,
        all_metrics_requirement = all_metrics_requirement,
        reference_col = reference_col,
        group_by_col = group_by_col,
        tolerance = tolerance,
        outlier_cut_off_upper = outlier_cut_off_upper)

      return prepped_data

    @output
    @render.plot
    def plot():

        reference_vals = list(input.reference_vals())

        prepped_data = reactive_data()

        plot_data = prepped_data[0]

        # Store the names of the columns to be used onthe X and Y axis
        x_axis = f'value{reference_vals[0]}'
        y_axis = f'value{reference_vals[1]}'

        # I grab what I am guessing to be the max and min tick marks on the x axist, in order to define the coordinates of a 'tolerance' ribbon.
        x_min = plot_data[x_axis].min()
        x_min_floor = np.floor(x_min * 10) / 10

        x_max = plot_data[x_axis].max()
        x_max_ceiling = np.ceil(x_max * 10) / 10

        plot = (ggplot(plot_data
        , aes(x_axis, y_axis, alpha=0.5))
        + geom_point(aes(color=region_col, size='avg_pop', alpha = 0.6))
        + geom_abline(intercept=0,slope=1) 
        + facet_wrap('~ source_var')
        + theme_light()
        + theme(legend_position='none')
        )

        return plot

    @output
    @render.table
    def table():

        prepped_data = reactive_data()

        summary = prepped_data[1].reset_index()

        return summary

app = App(app_ui, server)

## file: requirements.txt
Jinja2

```

:::

TESTING!!!
These are the same sections as above, to be used for testing.

```{python}

from plotnine import *
import plotly.express as px
import pandas as pd
import numpy as np

```

```{python}

#  ---------------------------------
#  ---------- SECTION 1: Data prep functions -------------
#  ----------------------------------

#  Function get matching for ref years
def closest_to_reference(df, reference_val, max_dist_from_ref, reference_col, group_by_col, value_col, tie_break):

  df = df.loc[:, [reference_col, group_by_col, value_col]]

  # Drop NAs
  df = df.dropna()

  # Calculate absolute distance from reference value
  df['ref_diff'] = abs(df[reference_col] - reference_val)

  # Drop any rows with a distance beyond threshold
  if not pd.isna(max_dist_from_ref):
    df = df.loc[df['ref_diff'] <= max_dist_from_ref]

  # Keep closest observation to reference value – including tie-breaks (where there is a match above and below the ref value)
  df = df[df.groupby(group_by_col)['ref_diff'].transform('min') == df['ref_diff']].reset_index(drop=True)

  # Settle tie-breaks
  if tie_break == 'below':
    df = df[df.groupby(group_by_col)[reference_col].transform('min') == df[reference_col]].reset_index(drop=True)

  elif tie_break == 'above':
    df = df[df.groupby(group_by_col)[reference_col].transform('max') == df[reference_col]].reset_index(drop=True)

  df = df.drop('ref_diff', axis=1)

  df = df\
    .rename(columns={value_col: "value"})

  return df

# Merge matches for different reference points
def merge_two_ref_matches(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):

# Make sure the pair of reference values are in ascending order
  reference_vals.sort()

# Maximise distance between two refs by settling tie-breaks below the lowest ref and above the highest ref

# Find matches for lower reference value
  lower_ref_matches = closest_to_reference(df, reference_vals[0], max_dist_from_refs, reference_col, group_by_col, value_col, 'below')

# Find matches for higher reference value
  higher_ref_matches = closest_to_reference(df, reference_vals[1], max_dist_from_refs, reference_col, group_by_col, value_col, 'above')

# Merge the two sets of matches
  merged_df = pd.merge(lower_ref_matches, higher_ref_matches, on=group_by_col, suffixes=(reference_vals[0], reference_vals[1]))

# Drop obs that do not have data for both ref values
  merged_df = merged_df.dropna()

# Drop obs where the matched data does not meet the min distance requirement
  if not pd.isna(min_dist_between):

  # Store the names of the reference column returned from the two matches
    ref_var_high = f'{reference_col}{reference_vals[1]}'
    ref_var_low = f'{reference_col}{reference_vals[0]}'

  # Keep only rows >= to the min distance
    merged_df = merged_df.loc[(merged_df[ref_var_high] - merged_df[ref_var_low]) >= min_dist_between, :]

  return merged_df

  # For PIP run this three times – first filtering data for just consumpion only, then with income only, then with a dataset that prefers income over consumption
def pip_welfare_routine(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):

  # Specify the name of the column in which the income/consumption welfare definition is stored
  welfare_colname = 'welfare_type'

  # Creat dataframes for thee scenarios:
  # Scenario 1: only allow income data
  df_inc_filter = df.loc[df[welfare_colname] == "income", :]
  df_inc_filter.name = "Income"

  # Scenario 2: only allow consumption data
  df_cons_filter = df.loc[df[welfare_colname] == "consumption", :]
  df_cons_filter.name = "Consumption"
  # Scenario 3: allow a mix – dropping consumption data where income data is available in the same year
  df_mixed = df.copy()

  df_mixed['welfare_count'] = df_mixed.groupby([reference_col, group_by_col])[welfare_colname].transform('count')

  df_mixed = df_mixed.loc[(df_mixed['welfare_count'] == 1) | (df_mixed[welfare_colname] == "income")]

  df_mixed.name = "Mixed"
  #  Store the scneario dataframes in a list
  df_scenarios = [df_inc_filter, df_cons_filter, df_mixed]

  # Run the matching function on each scenario
  scenario_matches = [merge_two_ref_matches(
    df_scenario, 
    reference_vals, 
    max_dist_from_refs, 
    min_dist_between, 
    reference_col, 
    group_by_col, 
    value_col) for df_scenario in df_scenarios]

  # Combine the first two scenarios.
  df_combined_matches = pd.concat([scenario_matches[0], scenario_matches[1]], keys=[df_scenarios[0].name, df_scenarios[1].name])

  # Tidy up indexes
  df_combined_matches = df_combined_matches.reset_index()

  df_combined_matches = df_combined_matches.drop('level_1', axis=1)

  df_combined_matches = df_combined_matches\
    .rename(columns={"level_0": "pip_welfare"})

  # Add in third scenario.
  df_combined_matches = pd.concat([df_combined_matches, scenario_matches[2]])

  # add scenario name to te pip_welfare column
  df_combined_matches['pip_welfare'] = df_combined_matches['pip_welfare'].fillna(df_scenarios[2].name)

  # Keep only one match per group (e.g. per Country) - in the priority laid out in the df_scenarios list above (income only -> consumption only -> mixed)
    # First count the matches
  df_combined_matches['match_count'] = df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
    # Then drop any matches from the lowest priority where there are multiple matches
  df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[2].name)]
    #  Repeat at the next level of priority
  df_combined_matches['match_count'] = df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
  df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[1].name)]

  # Drop the match count column
  df_combined_matches = df_combined_matches.drop('match_count', axis=1)

  return df_combined_matches

def prep_selected_vars(selected_vars):

  # selected_vars = [
  #   'LIS: Top_10_share_equiv_market',
  #   'WID: Bottom_50_share_posttax_dis',
  #   'PIP: Top_10_share']

  selected_sources = [var.split(sep, 1)[0] for var in selected_vars]
  selected_var_names = [var.split(sep, 1)[1] for var in selected_vars]

  match_data = {
    "sources": selected_sources,
    "var_names": selected_var_names
          }

  return match_data

# Specify the tolerances for each kind of metric
def return_tolerance_for_var(var):

  if 'Gini' in var:
    tolerance = 1

  elif 'Top_10_share' in var:
    tolerance = 1

  elif 'Bottom_50_share' in var:
    tolerance = 1

  else:
    tolerance = None

  return tolerance

def prep_data(
      data_to_match,
      reference_vals,
      max_dist_from_refs,
      min_dist_between,
      all_metrics_requirement,
      reference_col,
      group_by_col,
      outlier_cut_off_upper):

  # Combine the source and varname strings element-wise to produce keys that will be used for the concat later
  source_var_keys = [i + j for i, j in zip(data_to_match["sources"], data_to_match["var_names"])]

  matches = []

  for i in range(len(data_to_match['sources'])):

    source = data_to_match["sources"][i]
    var = data_to_match["var_names"][i]

    tolerance = return_tolerance_for_var(
      data_to_match["var_names"][i]
    )

    df = data[source]

    if source == "PIP":

      matched = pip_welfare_routine(
          df = df,
          reference_vals = reference_vals,
          max_dist_from_refs = max_dist_from_refs,
          min_dist_between = min_dist_between,
          reference_col = reference_col,
          group_by_col = group_by_col,
          value_col = var
        )

    else:

      matched = merge_two_ref_matches(
          df = df,
          reference_vals = reference_vals,
          max_dist_from_refs = max_dist_from_refs,
          min_dist_between = min_dist_between,
          reference_col = reference_col,
          group_by_col = group_by_col,
          value_col = var
        )

    matches.append(matched)

  plot_data = pd.concat(matches, keys=source_var_keys)

  # Tidy up indexes
  plot_data = plot_data.reset_index()

  plot_data = plot_data.drop('level_1', axis=1)

  plot_data = plot_data\
    .rename(columns={"level_0": "source_var"})

  # Store the names of the columns to be used onthe X and Y axis
  x_axis = f'value{reference_vals[0]}'
  y_axis = f'value{reference_vals[1]}'

  x_ref_val = f'{reference_col}{reference_vals[0]}'
  y_ref_val = f'{reference_col}{reference_vals[1]}'

  # Apply outlier cut off, if specified
  if not pd.isna(outlier_cut_off_upper):
    plot_data = plot_data.loc[plot_data[x_axis] <= outlier_cut_off_upper]
    plot_data = plot_data.loc[plot_data[y_axis] <= outlier_cut_off_upper]

  # Add a count by country – counting whether data is available from each source or not
  plot_data['source_count'] = plot_data.groupby(group_by_col)['source_var'].transform('count')

  # Apply metrics requirement (whether to include only observations with data from all metrics)
  if all_metrics_requirement:
    plot_data = plot_data.loc[plot_data['source_count'] == plot_data['source_count'].max()]

  # Drop source_count column
  plot_data = plot_data.drop('source_count', axis=1)

  # Add in region classification
  plot_data = pd.merge(plot_data, df_regions, how = 'left')

  # Add in population data
    # For first ref
  df_pop_ref = df_pop.loc[df_pop['Year'] == reference_vals[0], ['Entity', 'population'] ]

  plot_data = pd.merge(plot_data, df_pop_ref, how = 'left')

  plot_data = plot_data.rename(columns={'population':'population_ref1'})

    # For second ref
  df_pop_ref = df_pop.loc[df_pop['Year'] == reference_vals[1], ['Entity', 'population'] ]

  plot_data = pd.merge(plot_data, df_pop_ref, how = 'left')

  plot_data = plot_data.rename(columns={'population':'population_ref2'})

    # Calculate average population two reference periods
  plot_data['avg_pop'] = (plot_data['population_ref1'] + plot_data['population_ref2'])/2

  #  PREP SUMMARY TABLE

  # Calculate the change between the two ref periods
  plot_data['change'] = (plot_data[y_axis] - plot_data[x_axis])

  group_yes_vars = ['source_var', region_col]
  group_no_vars = ['source_var']

  group_scenarios = [group_no_vars, group_yes_vars]

  aggs = []

  for group_vars in group_scenarios:

  # Prepare pop-weighted average change
    agg_level = plot_data.copy()

    # Calulate pop weights by source and region
    agg_level['regional_pop_weights'] = agg_level.groupby(group_vars)['avg_pop'].transform(lambda x: x/x.sum())

    agg_level['global_pop_weights'] = agg_level.groupby('source_var')['avg_pop'].transform(lambda x: x/x.sum())

    # Multiply the change by the pop weights 
    agg_level['region_weighted_change'] = agg_level['change'] * agg_level['regional_pop_weights']

  # Prepare fall/stable/rise categories
    agg_level['fall'] = agg_level['change'] < -tolerance

    agg_level['stable'] = (agg_level['change'] <= tolerance) & (agg_level['change'] >= -tolerance)

    agg_level['rise'] = agg_level['change'] > tolerance

    agg_level[['fall', 'stable', 'rise']] = agg_level[['fall', 'stable', 'rise']].astype(int)

    agg_level = agg_level.groupby(group_vars)[['change', 'region_weighted_change', 'global_pop_weights', 'fall', 'stable', 'rise']]\
      .agg(['sum', 'mean', 'count'])

  # Aggregate by source and region
    if group_vars == group_yes_vars:

      agg_level = agg_level\
        .unstack().T.reset_index()

    elif group_vars == group_no_vars:

      agg_level = agg_level\
        .T.reset_index()

      agg_level[region_col] = "World"

    aggs.append(agg_level)

  df_summary = pd.concat(aggs)

    # filter for the aggregations we need and label in the 'summary' column
  df_summary.loc[(df_summary['level_0'] == 'change') & (df_summary['level_1'] == 'mean'), 'summary'] = 'avg change'

  df_summary.loc[(df_summary['level_0'] == 'region_weighted_change') & (df_summary['level_1'] == 'sum'), 'summary'] = 'Regional pop-weighted avg change'

  df_summary.loc[(df_summary['level_0'] == 'global_pop_weights') & (df_summary['level_1'] == 'sum'), 'summary'] = 'Global pop weights'

  df_summary.loc[(df_summary['level_0'] == 'change') & (df_summary['level_1'] == 'count'), 'summary'] = 'n'

  df_summary.loc[(df_summary['level_0'] == 'fall') & (df_summary['level_1'] == 'sum'), 'summary'] = 'fall'

  df_summary.loc[(df_summary['level_0'] == 'stable') & (df_summary['level_1'] == 'sum'), 'summary'] = 'stable'

  df_summary.loc[(df_summary['level_0'] == 'rise') & (df_summary['level_1'] == 'sum'), 'summary'] = 'rise'

  df_summary = df_summary[df_summary['summary'].notnull()]

  df_summary = df_summary.drop(['level_0', 'level_1'], axis = 1)

  df_summary['summary'] = pd.Categorical(df_summary['summary'], categories=['fall', 'stable', 'rise','n', 'avg change', 'Regional pop-weighted avg change', 'Global pop weights'], ordered=True)

  df_summary = df_summary.sort_values([region_col,'summary']).set_index([region_col,'summary'])

  # format to show appropriate d.p. I run a loop over each row for want of figuring out a better way!
  repeat_length = 7
  for i in range(0, len(df_summary.index.levels[0].unique())):
    df_summary.iloc[i*repeat_length,:] = df_summary.iloc[i*repeat_length,:].map('{:.0f}'.format)
    df_summary.iloc[1 + i*repeat_length,:] = df_summary.iloc[1 + i*repeat_length,:].map('{:.0f}'.format)
    df_summary.iloc[2 + i*repeat_length,:] = df_summary.iloc[2 + i*repeat_length,:].map('{:.0f}'.format)
    df_summary.iloc[3 + i*repeat_length,:] = df_summary.iloc[3 + i*repeat_length,:].map('{:.0f}'.format)
    df_summary.iloc[4 + i*repeat_length,:] = df_summary.iloc[4 + i*repeat_length,:].map('{:.2f}'.format)
    df_summary.iloc[5 + i*repeat_length,:] = df_summary.iloc[5 + i*repeat_length,:].map('{:.2f}'.format)
    df_summary.iloc[6 + i*repeat_length,:] = df_summary.iloc[6 + i*repeat_length,:].map('{:.2f}'.format)

  return plot_data, df_summary

```

```{python}

#  ---------------------------------
#  ---------- SECTION 2: Specifications -------------
#  ----------------------------------
# These parameters are not controlled in the app.

reference_col = 'Year'
group_by_col = 'Entity'

outlier_cut_off_upper = None
region_col = 'region_alt'

```

```{python}

#  ---------------------------------
#  ---------- SECTION 3: Read in data -------------
#  ----------------------------------

url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/pip.csv'

df_pip = pd.read_csv(url)

url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/wid.csv'

df_wid = pd.read_csv(url)

url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/lis.csv'

df_lis = pd.read_csv(url)

url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/region_mapping.csv'

df_regions = pd.read_csv(url)

url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/population.csv'

df_pop = pd.read_csv(url)

```

```{python}

#  ---------------------------------
#  ---------- SECTION 4: Gather data into dictionary and collete varnames -------------
#  ----------------------------------

data = {"PIP": df_pip,
        "WID": df_wid,
        "LIS": df_lis}

# make an empty list to add all the varnames
all_vars = list()

# A string to use to separate the source name from the var name
sep = ": "

# for each df in 'data'
for k,v in data.items():

  # make a list of varnames
  vars = list(v.columns)

  # remove the id cols
  vars.remove(reference_col)
  vars.remove(group_by_col)

  # append the name of the source (the key of 'data') at the front
  vars = [f'{k}{sep}' + var for var in vars]

  # append to the running list
  all_vars = all_vars + vars

```

We don't run this reactive step in the testing. We just run the prep data function in the plot and table functions below.

```{python}
# def reactive_data():

#   # For testing we define manually, rather than picked up from the app inputs
#   data_to_match = prep_selected_vars(input.metrics())

#   reference_vals = list(input.reference_vals())

#   max_dist_from_refs = input.max_dist_from_refs()
#   min_dist_between = input.min_dist_between()
#   all_metrics_requirement = input.all_metrics_requirement()

#   prepped_data = prep_data(
#       data_to_match = data_to_match,
#       reference_vals = reference_vals,
#       max_dist_from_refs = max_dist_from_refs,
#       min_dist_between = min_dist_between,
#       all_metrics_requirement = all_metrics_requirement,
#       reference_col = reference_col,
#       group_by_col = group_by_col,
#       tolerance = tolerance,
#       outlier_cut_off_upper = outlier_cut_off_upper)

#   return prepped_data

```

```{python}

def interactive_plot(
  selected_vars,
  reference_vals,
  max_dist_from_refs,
  min_dist_between,
  all_metrics_requirement):

  data_to_match = prep_selected_vars(selected_vars)

  prepped_data = prep_data(
      data_to_match = data_to_match,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = min_dist_between,
      all_metrics_requirement = all_metrics_requirement,
      reference_col = reference_col,
      group_by_col = group_by_col,
      outlier_cut_off_upper = outlier_cut_off_upper)

  plot_data = prepped_data[0]

  fig = px.scatter(
    plot_data, 
    x=f'value{reference_vals[0]}', 
    y=f'value{reference_vals[1]}',
    color=region_col, 
    facet_col='source_var', 
    facet_col_wrap=1, 
    size='avg_pop',
    hover_name=group_by_col,
    hover_data={
      f'value{reference_vals[0]}':':.2f',
      f'{reference_col}{reference_vals[0]}':True,
      f'value{reference_vals[1]}':':.2f',
      f'{reference_col}{reference_vals[1]}':True,
      region_col: True,
      'avg_pop': False,
      'source_var': False},
      height=1000
    )

  fig.update_layout(legend=dict(
      orientation="h",
      yanchor="bottom",
      y=1.02,
      xanchor="right",
      x=1
  ))

  fig.for_each_annotation(lambda a: a.update(text=a.text.split("=",1)[1]))

  fig.update_layout(
      legend_title=""
  )

  return fig

```

```{python}
def table_test(selected_vars,
  reference_vals,
  max_dist_from_refs,
  min_dist_between,
  all_metrics_requirement):

  data_to_match = prep_selected_vars(selected_vars)

  prepped_data = prep_data(
    data_to_match = data_to_match,
    reference_vals = reference_vals,
    max_dist_from_refs = max_dist_from_refs,
    min_dist_between = min_dist_between,
    all_metrics_requirement = all_metrics_requirement,
    reference_col = reference_col,
    group_by_col = group_by_col,
    outlier_cut_off_upper = outlier_cut_off_upper)

  summary = prepped_data[1].reset_index()

  return summary

```

```{python}

fig = interactive_plot(
  selected_vars = [
      'WID: Top_10_share_pretax',
      'PIP: Top_10_share'],
  reference_vals = [1990,2015],
  max_dist_from_refs = 5,
  min_dist_between = 10,
  all_metrics_requirement = True
)

fig.show()

```

```{python}

summary_table = table_test(
  selected_vars = [
      'WID: Top_10_share_pretax',
      'PIP: Top_10_share'],
  reference_vals = [1990,2015],
  max_dist_from_refs = 5,
  min_dist_between = 10,
  all_metrics_requirement = True
)

summary_table

```