### 2023-02-10 PhD_pages/Paper_1/Presentation_INET_16_feb_inequality_around_the_world.qmd
Paper 1: Drafting presentation

---
title: "Triangulating a global view on inequality"
author: "Joe Hasell"
format: 
    revealjs:
        theme: [default, revealjs_joe_theme.scss] 
bibliography: references.bib
---

## Concepts

- Three concepts of global inequality (Milanovic):

    1. Average within-country inequality

    2. Average within-country inequality, weighting by population

    3. Inequality across the world population taken as a whole (future work)

- Income(ish) inequality, not wealth

- Relative inequality, not absolute

## Why do we need a global view?

- Cosmopolitan values (e.g. UN's SDGs)
- Opportunities and limitations for national policy to address inequality
- Priorities for international policy (taxation, migration, aid...)
- Priorities for philanthropy 
- Winners and losers of globalization ('Degrowth')

## A clear global view is lacking {.smaller}

- **Limitations in the available data**
    - Coverage
    - Quality and comarabilty
    - Much improved recently!

- **Different conceptual focus in richer and poorer countries**
    - Inequality/relative poverty in high income countries
    - Absolute poverty in poorer countries
    - Much improved recently! (Atkinson commission... Atkinson & Brandolini 2010... Ravallion 2019... 'Societal poverty' – World Bank 2018, 2020, 2022)

- **Different 'worlds' of measurement occupied by global harmonized databases**
    - Different metrics (Gini, top income shares etc.); Different welfare concepts; underlying data sources; methods (e.g. treatment of 'missing top'; coverage or time periods analysed
    - Differences somewhat bundled together
    - Yield very different 'take home' messages

## Different 'worlds' of inequality measurement {.smaller}
**World Inequality Database**

> In recent decades, **income inequality has increased in nearly all countries**, but at different speeds...
[@Alvaredo2018-qq]

> Also find a quote about the within-coutry component of global inequality

## Different 'worlds' of inequality measurement {.smaller}
**World Bank, Poverty and Inequality Platform (formerly 'Povcal')**

> [Since c.1993] the number of countries with rising inequality was **slightly higher** than the number of countries with falling inequality. Yet, the average Gini among these 91 countries actually **declined**.
[@WBpoverty2016]

> Also find a quote about the within-coutry component of global inequality

## The view from the World Inequality Database {.smaller}
Top 1% share; pre-tax 'national' income; combining survey, tax and national accounts data; generalized pareto interpolation; individuals (equal split couples)

(CHART of top 1% WID)

## The view from the World Inequality Database {.smaller}
Top 1% share; pre-tax 'national' income; combining survey, tax and national accounts data; generalized pareto interpolation; individuals (equal split couples)

(TABLE of top 1% WID)

## The view from the World Bank Poverty and Inequality Platform {.smaller}
Gini; mix of consumption and disposable income; survey data (mostly micro, some 'grouped' data); lognormal or XXX interpolation (grouped data only); per capita household

(CHART of Gini PIP)

## The view from the World Bank Poverty and Inequality Platform {.smaller}
Gini; mix of consumption and disposable income; survey data (mostly micro, some 'grouped' data); lognormal or XXX interpolation (grouped data only); per capita household

(Table of Gini PIP)

## Connecting up these 'worlds' is a low hanging fruit {preview-links=true}

- The respective institutions – WIL and World Bank – have built hugely valuable and powerful platforms.
- We are in a better position than ever to get a global view on inequality
- But the two 'worlds' of measurement makes the current data terrain is *very* confusing for policimakers, researchers in adjacent fields, and the public.

Example: a summary of inequality trends [from the UN](https://www.un.org/en/un75/inequality-bridging-divide).

## Connecting up these 'worlds' is a low hanging fruit

- Existing studies tend to be based on only one data source, or only partially bridge between the two.

    - Flagship reports of the two institutions
    - Lakner Milanovic
    - Groeningen...
    - Ravi Kanbur...

## Connecting up these 'worlds'

1. Line up what we can:
    - Coverage and time period
    - Metrics

    - (With lower coverage) Income concept
        - WID pre-tax
        - LIS?
        - Income vs consumption: GCIP

    - Methods: Hacky top income adjustment to PIP data?

2. Triangulate
    - Flag what we *don't* know – Isolate what is causing the differences between 'take home' messages
        - Particular methodological choices?
        - Particular countries?
    - Gather together what we *do* know, irrespecive of may remaining differences

## Bibliography