### 2023-02-10 presentations.qmd

---
title: "Presentations"
format: html
---

### From $1.90 to $2.15: the updated International Poverty Line
20/01/2023

A presentation given to Masters students in Comparative Social Policy at the Department for Social Policy, University of Oxford.

[Slides](other_presentations/DSPI_CSP_masters_Jan2023/DSPI_CSP_masters_Jan2023.qmd)