### 2023-02-22 PhD_pages/Paper_1/Presentation_INET_16_feb_inequality_around_the_world.qmd
Paper 1: preparing presentation

---
title: "Triangulating a global view on inequality"
author: "<br> <br> Joe Hasell – DPhil candidate at DSPI University of Oxford; Our World in Data<br> <br> Pablo Arriagada – Our World in Data"
format: 
    revealjs:
        theme: [default, revealjs_joe_theme.scss] 
bibliography: references.bib
---

```{python}

from plotnine import *
import plotly.express as px
import pandas as pd
import numpy as np
from IPython.display import display, HTML, IFrame, Markdown
import dataframe_image as dfi
import imgkit

from wrangling_and_plotting_functions import *

pd.set_option('display.max_rows', None)

```

```{python}

fp = 'data_appendix/data/clean/pip.csv'

df_pip = pd.read_csv(fp)

fp = 'data_appendix/data/clean/wid.csv'

df_wid = pd.read_csv(fp)

fp = 'data_appendix/data/clean/lis.csv'

df_lis = pd.read_csv(fp)

fp = 'data_appendix/data/clean/region_mapping.csv'

df_regions = pd.read_csv(fp)

fp = 'data_appendix/data/clean/population.csv'

df_pop = pd.read_csv(fp)

fp = 'data_appendix/data/clean/region_population.csv'

df_pop_regions = pd.read_csv(fp)

# Gather data into dictionary
source_dfs = {
    "PIP": df_pip,
    "WID": df_wid,
    "LIS": df_lis
    }

tolerance_lookup =  {
                'var_search_term': [
                    'Gini',
                    'Top 1pc share', 
                    'Top 10pc share',
                    'Bottom 50pc share'],
                'var_tolerance': [
                    1,
                    1,
                    1,
                    1
                ]
            }

```

```{python}
def make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement,selected_vars):
    msg = f'{reference_vals[0]} vs {reference_vals[1]} (+/- {max_dist_from_refs} years)'

    if len(selected_vars)>1:
        if all_metrics_requirement:
            msg = f'{msg} – Matched countries only'
        else:
            msg = f'{msg} – All available countries'

    return(msg)

```

## Concepts
::: {.r-fit-text}
- Three concepts of global inequality [not quite @Milanovic2007; or @AnandSegal2015-nh]

    1*. Average within-country inequality

    2*. Average within-country inequality, weighting by population

    3. Inequality across the world population taken as a whole
        - Future work

- We will be looking at:
    - Income(ish) inequality, not wealth
    - Relative inequality, not absolute
:::

## {.smaller}
### Why do we need a global view?
::: {.r-fit-text}
::: {.incremental}
- Understanding the winners and losers of globalization and economic growth
    - *"We’re fighting for Main Street, not Wall Street. We have rejected globalism and embraced patriotism."* - Donald Trump, [Aug 2020](https://www.youtube.com/watch?v=l83sbXD0nuY&t=11m15s)
- To inform national policy to address inequality – both opportunities and limits
- To inform international policy to address inequality (financial reporting transparency, tax havens, global taxation, migration, aid...)
- A reflection of cosmopolitan values and outlook (e.g. UN's SDGs)
- As a guide to 'Effective Altruism', including charitable giving
:::
:::

## {.smaller}
### A clear global view is lacking
::: {.r-fit-text}
::: {.incremental}

1. **Limitations in the available data**
    - Coverage
    - Quality and comarabilty within secondary databases [@AtkinsonBrandolini2001-qt]
    - Much improved recently!
        - Expanded coverage of harmonized secondary databases: World Bank's *Poverty and Inequality Platform* (PIP), *World Inequality Database* (WID), *Luxembourg Income Study* (LIS)

2. **Different conceptual focus in richer and poorer countries**
    - Income inequality/relative poverty in high income countries
    - Absolute poverty in poorer countries
    - Much improved recently!
        - Atkinson commission [-@Atkinson2016-xp]; 'Societal poverty' – [@World_Bank2018-eo]; UN SDGs; expansion of WID coverage

3. **Different 'worlds' of measurement**
    - Many methodological differences between different global databases, *often bundled together*
        - Yield very different take-home messages about inequality and how it has evolved

:::
:::

## {.smaller}
### Different 'worlds': *World Inequality Database*
::: {.incremental}

> In recent decades, **income inequality has increased in nearly all countries**, but at different speeds...
[@Alvaredo2018-qq]
:::

## {.smaller}
### Different 'worlds': *World Inequality Database*

The vast majority of global inequality (concept 3) is now due to differences **within** countries.

![](images/WID_global_ineq_decomp.png)

## {.smaller}
### Different 'worlds': *World Bank, Poverty and Inequality Platform*
::: {.incremental}

> [Since c.1993] the number of countries with rising inequality was **slightly higher** than the number of countries with falling inequality. Yet, the average Gini among these 91 countries actually **declined**.
[@WBpoverty2016]
:::

## {.smaller}
### Different 'worlds': *World Bank, Poverty and Inequality Platform*

The vast majority of global inequality (concept 3) remains due to differences **between** countries.

![](images/pip_global_ineq_decomp.png){fig-align="center"}

##
### Different 'worlds'
::: {.r-fit-text}
::: {.incremental}
In terms of data availability, quality and comparability, we are in a better position than ever.

Yet some of the *most basic* questions are still open:

- Is rising inequality within countries a near-ubiquitous, global phenomenon?
    - **WID:** Yes
    - **PIP:** No

- Where do global inequalities lie: in differences **between** richer and poorer countries, or in differences between richer and poorer people **within** countries?
    - **WID:** Within countries
    - **PIP:** Between countries
:::
:::

## 
### Different methodologies

::: {.r-fit-text}

::: {.incremental}

The different 'worlds' are separated by many methodological differences, often bundled together:

- Inequality indices (and focal points in the distribution)
    - **WID:** Focus on the top
    - **PIP:** Focus on overall inequality (e.g. Gini)

- Welfare concepts
    - **WID:** Focus on (and greatest coverage for) pre-tax national income
    - **PIP:** mix of disposable income and consumption

:::

:::

## 
### Different methodologies

::: {.r-fit-text}

::: {.incremental}

- Underlying data sources
    - **WID:** Tax + NA + survey 
    - **PIP:** survey only

- Methodologies
    - **WID:** Adjustments for misisng top in survey data, Scaling to national accounts aggregates
    - **PIP:** Some ex-post harmonization of welfare concepts

- Coverage
    - Differences across the two datasets, impacting the time period that can be analysed

:::

:::

## {.smaller}

**Connecting up these 'worlds' is a low-hanging fruit...**

::: {.r-fit-text}
::: {.incremental}

**For policy and public debate**

- The data is better than ever
- But the two 'worlds' of measurement makes the current data terrain *very* confusing
    - Example: a summary of inequality trends [from the UN](https://www.un.org/en/un75/inequality-bridging-divide).

:::
:::

## {.smaller .scrollable}

**Connecting up these 'worlds' is a low-hanging fruit...**

::: {.r-fit-text}
::: {.incremental}

**For academic research**

- Existing studies tend to be based on only one 'world'
    - Flagship reports of the two institutions [@chancel2021; @World_Bank2022-vr]
    - Kanbur et al. [-@Kanbur2022-lk]
    - Niño-Zarazúa et al. [-@Nino-Zarazua2017-tv]
    - van Zanden et al. [-@Van_Zanden2014-zo]
    - Ravallion[-@Ravallion2014-mk]

- Two notable counter-examples:
    - Anand and Segal [-@AnandSegal2015-nh]
        - Adjusts survey data to match WTID top income shares (with fairly small effect)
        - Estimtates up to 2005
    - Lakner and Milanovic [-@laknerMilanovic2016]
        - Adjusts survey data to match WTID top income shares (with fairly small effect)
        - Only considers impact on overall global inequality
        - Estimates up to 2008
    - Neither focus on average within-country trends (Concepts 1* and 2* above)
    - Coverage in the WID data has expanded subsequently
:::
:::

# Connecting up the 'worlds'

## {.smaller}
### Connecting up the 'worlds'
1. Line up what we can:
    - Coverage and time period
        - 1995 vs 2015 (+/-3) – good coverage 
        - 1980 vs 2018 (+/-5) – much worse coverage, more relevant period

    - Inequality indices

    - Income concept 
        - Very cursory glance [unlike @MorelliYonzan2022-zn; @CarranzaNolan2022-ko]

## {.smaller}
### Connecting up the 'worlds'

2. Triangulate
    - Pinpoint what we *don't* know; what is behind the differences:
        - Particular countries?
        - Particular methodological differences?
    - Gather together what we *do* know, irrespecive of the many remaining differences

# Coverage

## {.smaller}
Coverage gets very good in both datasets from ~1990.

```{python}

df_coverage = pd.DataFrame()

data_to_check = {
    'Source': ['PIP', 'WID'],
    'Metric': ['Gini', 'Top 1pc share – pretax'],
    'df': [df_pip, df_wid]
    }

max_dist_from_ref = 3

for yr in range(1960, 2023):

    for i in [0,1]:

        matched_data = closest_to_reference(
            df=data_to_check['df'][i], 
            reference_val = yr, 
            max_dist_from_ref = max_dist_from_ref, 
            reference_col = 'Year', 
            group_by_col = 'Entity', 
            value_col = data_to_check['Metric'][i], 
            tie_break = 'above')

        df_pop_ref_year = df_pop[df_pop['Year']==yr].drop(columns=['Year', 'population'])

        matched_data = matched_data.merge(df_pop_ref_year, how='left')

        coverage_yr = pd.DataFrame({
            'Year': [yr],
            'No. of countries': [matched_data['world_pop_share'].count()],
            'Share of World population': [matched_data['world_pop_share'].sum()]
        })

        source = data_to_check['Source'][i]
        metric = data_to_check['Metric'][i]
        coverage_yr['Data'] = f'{source} – {metric}'

        df_coverage = pd.concat([df_coverage, coverage_yr])

```

```{python}

fig = px.line(df_coverage, x="Year", y="Share of World population", color='Data')

fig.update_layout(
    title = f'Population coverage – within {max_dist_from_ref} years',
    legend_title="",
    legend=dict(
    orientation="h",
    yanchor="bottom",
    y=1.02,
    xanchor="right",
    x=1
))

fig.show()

```

## {.smaller}
Coverage gets very good in both datasets from ~1990.

```{python}

fig = px.line(df_coverage, x="Year", y="No. of countries", color='Data')

fig.update_layout(
    title = f'No. of countries – within {max_dist_from_ref} years',
    legend_title="",
    legend=dict(
    orientation="h",
    yanchor="bottom",
    y=1.02,
    xanchor="right",
    x=1
))

fig.show()

```

## {.smaller}
We can increase 'coverage' in 1980 by widening the net.

```{python}

df_coverage = pd.DataFrame()

data_to_check = {
    'Source': ['PIP', 'WID'],
    'Metric': ['Gini', 'Top 1pc share – pretax'],
    'df': [df_pip, df_wid]
    }

max_dist_from_ref = 5

for yr in range(1960, 2023):

    for i in [0,1]:

        matched_data = closest_to_reference(
            df=data_to_check['df'][i], 
            reference_val = yr, 
            max_dist_from_ref = max_dist_from_ref, 
            reference_col = 'Year', 
            group_by_col = 'Entity', 
            value_col = data_to_check['Metric'][i], 
            tie_break = 'above')

        df_pop_ref_year = df_pop[df_pop['Year']==yr].drop(columns=['Year', 'population'])

        matched_data = matched_data.merge(df_pop_ref_year, how='left')

        coverage_yr = pd.DataFrame({
            'Year': [yr],
            'No. of countries': [matched_data['world_pop_share'].count()],
            'Share of World population': [matched_data['world_pop_share'].sum()]
        })

        source = data_to_check['Source'][i]
        metric = data_to_check['Metric'][i]
        coverage_yr['Data'] = f'{source} – {metric}'

        df_coverage = pd.concat([df_coverage, coverage_yr])

```

```{python}

fig = px.line(df_coverage, x="Year", y="Share of World population", color='Data')

fig.update_layout(
    title = f'Population coverage – within {max_dist_from_ref} years',
    legend_title="",
    legend=dict(
    orientation="h",
    yanchor="bottom",
    y=1.02,
    xanchor="right",
    x=1
))

fig.show()

```

## {.smaller}
We can increase 'coverage' in 1980 by widening the net.

```{python}

fig = px.line(df_coverage, x="Year", y="No. of countries", color='Data')

fig.update_layout(
    title = f'No. of countries – within {max_dist_from_ref} years',
    legend_title="",
    legend=dict(
    orientation="h",
    yanchor="bottom",
    y=1.02,
    xanchor="right",
    x=1
))

fig.show()

```

# 1995 vs 2015

## {.smaller}
```{python}
# Set-up to compare WID and PIP top 10% share, all countries
reference_vals = [1995, 2015]
max_dist_from_refs = 3
all_metrics_requirement = False
selected_vars = [
        'WID: Top 10pc share – pretax',
        'PIP: Top 10pc share'
        ]

print(make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement, selected_vars))

```

```{python}

# Table WID vs PIP top 10% share, all countries

df_summary = change_summary_table(
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_pop_regions = df_pop_regions,
      df_regions = df_regions,
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      all_metrics_requirement = all_metrics_requirement,
      reference_col = "Year",
      group_by_col = "Entity",
      outlier_cut_off_upper = None,
      tolerance_lookup =  tolerance_lookup,
      region_col = 'region'
      )

df_summary.drop(["Other N. America", "Other Oceania"], axis=0, inplace=True)

dfi.export(df_summary,"images/tables/pres_table4.png")

```
![](images/tables/pres_table4.png)

## {.smaller}
```{python}
# Switch to alow only countries with matched data from both PIP and WID for top 10% share

all_metrics_requirement = True

print(make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement, selected_vars))

```

```{python}

# Table WID vs PIP top 10% share, only matched countries

df_summary = change_summary_table(
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_pop_regions = df_pop_regions,
      df_regions = df_regions,
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      all_metrics_requirement = all_metrics_requirement,
      reference_col = "Year",
      group_by_col = "Entity",
      outlier_cut_off_upper = None,
      tolerance_lookup =  tolerance_lookup,
      region_col = 'region'
      )

df_summary.drop(["Other N. America", "Other Oceania"], axis=0, inplace=True)

dfi.export(df_summary,"images/tables/pres_table5.png")

```
![](images/tables/pres_table5.png)

::: {.incremental}

- Country coverage drops (less so population-weighted)
- Drop in coverage not uniform across regions and source (LAC and SS Africa in PIP)

:::

## {.smaller}

```{python}
print(make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement, selected_vars))

```
![](images/tables/pres_table5.png)

::: {.incremental}

Key differences:

- Inequality is far higher in WID than PIP
- A very substantial rise in the pop-weighted average in WID.

:::

## {.smaller}

```{python}
print(make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement, selected_vars))

```
![](images/tables/pres_table5.png)

::: {.incremental}

Key similarities:

- Rising vs falling: approx. equal in both sources
- Unweighted avg approx. stable in both (a fall in PIP)
- Weighted average less benign trend than unweighted average

:::

##
```{python}

# Scatter WID (WID vs PIP top 10% share, only matched countries)

plot_change_scatter(
      plot_var = selected_vars[0],
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_regions = df_regions,
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      all_metrics_requirement = all_metrics_requirement,
      reference_col = "Year",
      group_by_col = "Entity",
      outlier_cut_off_upper = None,
      tolerance_lookup =  tolerance_lookup,
      region_col = 'region'
      )[0]

```

##
```{python}

# Scatter WID (WID vs PIP top 10% share, only matched countries)

plot_change_scatter(
      plot_var = selected_vars[1],
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_regions = df_regions,
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      all_metrics_requirement = all_metrics_requirement,
      reference_col = "Year",
      group_by_col = "Entity",
      outlier_cut_off_upper = None,
      tolerance_lookup =  tolerance_lookup,
      region_col = 'region'
      )[0]

```

## {.smaller .scrollable}
```{python}

# Scatter difference (WID vs PIP top 10% share, only matched countries)

plot_change_scatter(
      plot_var = 'diff',
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_regions = df_regions,
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      all_metrics_requirement = all_metrics_requirement,
      reference_col = "Year",
      group_by_col = "Entity",
      outlier_cut_off_upper = None,
      tolerance_lookup =  tolerance_lookup,
      region_col = 'region'
      )[0]

```

```{python}
df_change = plot_change_scatter(
      plot_var = 'diff',
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_regions = df_regions,
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      all_metrics_requirement = all_metrics_requirement,
      reference_col = "Year",
      group_by_col = "Entity",
      outlier_cut_off_upper = None,
      tolerance_lookup =  tolerance_lookup,
      region_col = 'region'
      )[1]

df_change['change_diff'] = df_change[f'change {selected_vars[0]}'] - df_change[f'change {selected_vars[1]}']

df_change.to_csv('df_change.csv')

```
::: {.incremental}

- More countries below than above the line
- For the *majority* of countries the gap is not huge (28/79 > 2 pt. difference; 20/79 > 4 pt. difference)
- *Very* large differences in a small number of coutries – which happen to be very populous

:::

## Top 1 vs Top 10?

## {.smaller}

```{python}
# Set-up to compare WID top 1% and top 10% share, all countries
reference_vals = [1995, 2015]
max_dist_from_refs = 3
all_metrics_requirement = False
selected_vars = [
        'WID: Top 1pc share – pretax',
        'WID: Top 10pc share – pretax'
        ]

print(make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement, selected_vars))

```

```{python}

# Table WID top 1% and top 10% share, all countries

df_summary = change_summary_table(
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_pop_regions = df_pop_regions,
      df_regions = df_regions,
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      all_metrics_requirement = all_metrics_requirement,
      reference_col = "Year",
      group_by_col = "Entity",
      outlier_cut_off_upper = None,
      tolerance_lookup =  tolerance_lookup,
      region_col = 'region'
      )

df_summary.drop(["Other N. America", "Other Oceania"], axis=0, inplace=True)

dfi.export(df_summary,"images/tables/pres_table_top_1_top10.png")

```
![](images/tables/pres_table_top_1_top10.png)

::: {.incremental}

- Differences for some individual countries – LAC
- But for other regions, and globally, the story over time is very similar.

:::

# 1980 vs 2018

## {.smaller}
```{python}

reference_vals = [1980, 2018]
max_dist_from_refs = 5
all_metrics_requirement = False
selected_vars = [
        'WID: Top 10pc share – pretax',
        'PIP: Top 10pc share'
        ]

print(make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement, selected_vars))

```

```{python}

df_summary = change_summary_table(
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_pop_regions = df_pop_regions,
      df_regions = df_regions,
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      all_metrics_requirement = all_metrics_requirement,
      reference_col = "Year",
      group_by_col = "Entity",
      outlier_cut_off_upper = None,
      tolerance_lookup =  tolerance_lookup,
      region_col = 'region'
      )

df_summary.drop(["Other N. America", "Other Oceania"], axis=0, inplace=True)

dfi.export(df_summary,"images/tables/pres_table8.png")

```
![](images/tables/pres_table8.png)

::: {.incremental}

- Even before aligning coverage, the country coverage drops substantially for both datasets - so too does population coverage
- In WID a substantial rise in the unweighted average; none in PIP
- A huge rise in weighted average in WID; small rise in PIP
- PIP is the binding constraint on coverage: almost no EE & CA coverage; half the number of W. European countries.

:::

## {.smaller}
```{python}

reference_vals = [1980, 2018]
max_dist_from_refs = 5
all_metrics_requirement = True
selected_vars = [
        'WID: Top 10pc share – pretax',
        'PIP: Top 10pc share'
        ]

print(make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement, selected_vars))

```

```{python}

df_summary = change_summary_table(
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_pop_regions = df_pop_regions,
      df_regions = df_regions,
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      all_metrics_requirement = all_metrics_requirement,
      reference_col = "Year",
      group_by_col = "Entity",
      outlier_cut_off_upper = None,
      tolerance_lookup =  tolerance_lookup,
      region_col = 'region'
      )

df_summary.drop(["Other N. America", "Other Oceania"], axis=0, inplace=True)

dfi.export(df_summary,"images/tables/pres_table9.png")

```
![](images/tables/pres_table9.png)

::: {.incremental}

-  Very few countries available in both datasets for this period (though population coverage falls less)
- Aligning coverage does bring the datasets somewhat closer: 
    - smaller rise in global unweighted average in WID
    - more balanced split between rise and fall
- But the stark difference remains: a huge rise in weighted average in WID

:::

## {.smaller .scrollable}

```{python}

plot_change_scatter(
      plot_var = 'diff',
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_regions = df_regions,
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      all_metrics_requirement = all_metrics_requirement,
      reference_col = "Year",
      group_by_col = "Entity",
      outlier_cut_off_upper = None,
      tolerance_lookup =  tolerance_lookup,
      region_col = 'region'
      )[0]

```

- For those countries with data from both PIP and WID, the majority tell a surprisingly similar story about change over time.
- Only for a small number are there *very* large differences – this includes a number of populous countries: India, China, US, Mexico

## Top 1 vs Top 10?

## {.smaller}

```{python}
# Set-up to compare WID top 1% and top 10% share, all countries
all_metrics_requirement = False
selected_vars = [
        'WID: Top 1pc share – pretax',
        'WID: Top 10pc share – pretax'
        ]

print(make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement, selected_vars))

```

```{python}

# Table WID top 1% and top 10% share, all countries

df_summary = change_summary_table(
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_pop_regions = df_pop_regions,
      df_regions = df_regions,
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      all_metrics_requirement = all_metrics_requirement,
      reference_col = "Year",
      group_by_col = "Entity",
      outlier_cut_off_upper = None,
      tolerance_lookup =  tolerance_lookup,
      region_col = 'region'
      )

df_summary.drop(["Other N. America", "Other Oceania"], axis=0, inplace=True)

dfi.export(df_summary,"images/tables/pres_table_top_1_top10_1980.png")

```
![](images/tables/pres_table_top_1_top10_1980.png)

::: {.incremental}

- Much more divergence visible than for 1995-2015 period – especially population-weighted
- Average top 1% doubled, or nearly, in S & SE Asia, East Asia and W. Offshoots. EE & CA quadrupled.

:::

## {.smaller}

```{python}
# Set-up to compare WID top 1% and top 10% share, all countries
all_metrics_requirement = True
selected_vars = [
        'WID: Top 1pc share – pretax',
        'PIP: Top 10pc share'
        ]

print(make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement, selected_vars))

```

```{python}

# Table WID top 1% and PIP top 10% share, all countries

df_summary = change_summary_table(
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_pop_regions = df_pop_regions,
      df_regions = df_regions,
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      all_metrics_requirement = all_metrics_requirement,
      reference_col = "Year",
      group_by_col = "Entity",
      outlier_cut_off_upper = None,
      tolerance_lookup =  tolerance_lookup,
      region_col = 'region'
      )

df_summary.drop(["Other N. America", "Other Oceania"], axis=0, inplace=True)

dfi.export(df_summary,"images/tables/pres_table_WID_top_1_PIP_top10_1980.png")

```
![](images/tables/pres_table_WID_top_1_PIP_top10_1980.png)

::: {.incremental}

- This remains true even after aligning coverage with PIP.

:::

# Welfare concept & methods

## LIS pre- and post-tax data

## {.smaller}

```{python}

reference_vals = [1980, 2018]
max_dist_from_refs = 5
all_metrics_requirement = True
selected_vars = [
        'LIS: Top 10pc share – percap_disposable',
        'LIS: Top 10pc share – percap_market',
        ]

print(make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement, selected_vars))

```

```{python}

df_summary = change_summary_table(
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_pop_regions = df_pop_regions,
      df_regions = df_regions,
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      all_metrics_requirement = all_metrics_requirement,
      reference_col = "Year",
      group_by_col = "Entity",
      outlier_cut_off_upper = None,
      tolerance_lookup =  tolerance_lookup,
      region_col = 'region'
      )

df_summary.drop(["Other N. America", "Other Oceania"], axis=0, inplace=True)

dfi.export(df_summary,"images/tables/LIS_on_LIS.png")

```
![](images/tables/LIS_on_LIS.png)

::: {.incremental}

- Very low coverage
- For W. Europe and W. Offshoots we do see some divergence between pre- and post- tax inequality.

:::

## WID pre- vs post-tax data

## {.smaller}

```{python}

reference_vals = [1980, 2018]
max_dist_from_refs = 5
all_metrics_requirement = True
selected_vars = [
        'WID: Top 10pc share – pretax',
        'WID: Top 10pc share – posttax_nat',
        ]

print(make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement, selected_vars))

```

```{python}

df_summary = change_summary_table(
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_pop_regions = df_pop_regions,
      df_regions = df_regions,
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      all_metrics_requirement = all_metrics_requirement,
      reference_col = "Year",
      group_by_col = "Entity",
      outlier_cut_off_upper = None,
      tolerance_lookup =  tolerance_lookup,
      region_col = 'region'
      )

df_summary.drop(["Other N. America", "Other Oceania"], axis=0, inplace=True)

dfi.export(df_summary,"images/tables/WID_on_WID.png")

```
![](images/tables/WID_on_WID.png)

::: {.incremental}

- Again very low coverage
- For both EE & CA and W. Europe, quite a similar story in pre- and post-tax trends – albeit huge rises within EE & CA.

:::

## Individual countries

##
```{python}

def plot_one_country_comparison(selected_entity, selected_vars):

    selected_source_var_dict = prep_selected_vars(selected_vars)

    df_long = pd.DataFrame()

    for i in range(len(selected_source_var_dict['sources'])):

        source = selected_source_var_dict["sources"][i]
        var = selected_source_var_dict["var_names"][i]

        df = source_dfs[source]

        if source == "PIP":

            df1 = df.loc[df['welfare_type'] == "income", ['Year', 'Entity', var]]\
                .rename(columns = {var: 'value'})

            df1['data'] = f'{source}: {var} (income)'

            df_long = pd.concat([df_long, df1])

            df2 = df.loc[df['welfare_type'] == "consumption", ['Year', 'Entity', var, 'welfare_type']]\
                .rename(columns = {var: 'value'})

            df2['data'] = f'{source}: {var} (consumption)'

            df_long = pd.concat([df_long, df2])

        else:

            df = df[['Year', 'Entity', var]]\
                .rename(columns = {var: 'value'})

            df['data'] = f'{source}: {var}'

            df_long = pd.concat([df_long, df])

    df_plot = df_long[df_long['Entity']==selected_entity]

    fig = px.line(df_plot, x="Year", y="value", color='data',
    title = f'A range of series for {selected_entity}'
    )

    fig.update_layout(showlegend=True, legend=dict(
    title_font_family='Lato',
    font=dict(
        size=8
    )
    ))
    fig.show()

```

## {.smaller}

```{python}
selected_vars = [
        'WID: Top 10pc share – pretax',
        'WID: Top 10pc share – posttax_nat',
        'PIP: Top 10pc share',
        'LIS: Top 10pc share – percap_disposable',
        'LIS: Top 10pc share – percap_market'
        ]

plot_one_country_comparison(selected_entity = "United States", selected_vars = selected_vars)
```

::: {.incremental}

- In the US, similar trends across survey and WID 1975-1995
- Some divergence thereafter between pre- and post-tax series - also in LIS survey data
- Very strong divergence of WID series (both pre- and post-tax) 1995 onwards.

:::

## {.smaller .scrollable}

```{python}

plot_one_country_comparison(selected_entity = "Russia", selected_vars = selected_vars)
```

::: {.incremental}

- In Russia, huge uncertainty concerning the extent and duration of rising inequality over transition.
- Quality of the data subject to a lot of discussion [@Milanovic1998-sy; @Novokmet2018-px; @Atkinson-Micklewright1992-qr]
- More agreement from mid-2000s

:::

## {.smaller}

```{python}
plot_one_country_comparison(selected_entity = "Mexico", selected_vars = selected_vars)
```

::: {.incremental}

- In Mexico (and other LAC countries) WID is using a linear interpolation between decadal estimates prior to ~2000.

- This relates to methods (tax data, top adjustments) rather than income concept.

:::

## {.smaller}

```{python}
plot_one_country_comparison(selected_entity = "Brazil", selected_vars = selected_vars)
```

::: {.incremental}

- Similar in Brazil, except WID goes with the survey data after 2000.

:::

## {.smaller .scrollable}

```{python}
plot_one_country_comparison(selected_entity = "India", selected_vars = selected_vars)
```

::: {.incremental}

- In India, only two survey available in LIS.
- Some divergence between pre- and post- tax inequality in the LIS data, but tiny compared to the divergence of WID pre-tax. 
- Note PIP data relates to *consumption*
- A period of particular divergence between national accounts and survey data [@Deaton2005-oy; @Sandefur2022-lx]

:::

## {.smaller}

```{python}
plot_one_country_comparison(selected_entity = "China", selected_vars = selected_vars)
```

::: {.incremental}

- In China, some divergence of pre- and post-tax nequality apparent in LIS survey data
- Very condensed period of divergence in WID between 2000-2005
- Again substantial divergence between national accounts and survey means over this time [@Deaton2005-oy]

:::

## {.smaller}

```{python}
plot_one_country_comparison(selected_entity = "Nigeria", selected_vars = selected_vars)
```

::: {.incremental}

- Nigeria – WID seems not to account for 1996 survey in which inequality apparently peaked.
- Breaks in comparability noted by the World Bank – see this [Our World in Data chart](https://ourworldindata.org/explorers/poverty-explorer?hideControls=false&Metric=Share+in+poverty&Poverty+line=%242.15+per+day%3A+International+Poverty+Line&Household+survey+data+type=Show+data+from+both+income+and+consumption+surveys&Show+breaks+between+less+comparable+surveys=true&country=~NGA).

:::

# Triangulating the key facts of global inequality
Concept 1* and 2*

## {.smaller}
### Since 1995:

::: {.r-fit-text}
::: {.incremental}

**Key differences:**

- Strong rise in the population-weighted average level of inequality (Concept 2*) in WID that is absent in PIP.

**Key similarities:**

- For a majority of countries, trends look quite similar across PIP and WID.
- Inequality in the average country was roughly flat, with roughly as many coutries with rising inequality as with falling inequality
- Inequality was rising particularly strongly in some very populous countries, bringing up the trend in the weighted average
- Different indices – even a focus on the *very* top – paint very similar pictures overall.

:::

:::

## {.smaller}
### Since 1980
::: {.incremental}
::: {.r-fit-text}

**Key differences:**

- Significant rise in unweighted average inequality (Concept 1*) in WID that is absent in PIP.
- Huge rise in the population-weighted average level of inequality (Concept 2*) in WID that is much more muted in PIP.
- Stronger differences looking at the *very* top

**Key similarities**

- Amongst those (few) countries with data in both PIP and WID, trends look quite similar for a majority of countries.
- Inequality was rising particularly strongly in some very populous countries, bringing up the trend in the weighted average

:::

:::

## {.smaller .scrollable}
### Lessons (Concept 3)

::: {.r-fit-text}

::: {.incremental}

- The global picture – when weighting by population – is greatly affected by a small number of populous countries. Trends diverge substantialy between PIP and WID for a number of these.

- These larger discrepancies in trends relate to methodological choices, adjustment procedures etc. – the income concept plays a minor role at most.

- The fact that these large differences are a (substantial) minority of cases suggests they may reflect idiosynctratic data quality or methodological issues specific to these countries – and not just a general issue of the 'the missing top'.
    - Some examples of possibly idiosyncratic issues demonstrated above

- Room for the two worlds be more proactively connecting up and explaining divergences – particularly in these cases where there are large discrepencies and large populations.

- Analysis of global inequality trends – when weighting by population – should factor in this potential sensitivity to localised data issues
    - Something not currently done in the literature on global inequality
    - Future work!

:::

:::

##

# Appendix

## Countries only found in one source - 1995 vs 2015

```{python}

reference_vals = [1995, 2015]
max_dist_from_refs = 3
all_metrics_requirement = False
selected_vars = [
        'WID: Top 10pc share – pretax',
        'PIP: Top 10pc share'
        ]

```

```{python}
prepped_data = prep_data(
        source_dfs = source_dfs,
        df_pop = df_pop,
        df_regions = df_regions,
        selected_vars = selected_vars,
        reference_vals = reference_vals,
        max_dist_from_refs = max_dist_from_refs,
        min_dist_between = 1,
        all_metrics_requirement = all_metrics_requirement,
        reference_col = "Year",
        group_by_col = "Entity",
        outlier_cut_off_upper = None,
        tolerance_lookup =  tolerance_lookup
        )

prepped_data['source_count'] = prepped_data.groupby('Entity')['source_var'].transform('count')

prepped_data = prepped_data[prepped_data['source_count']==1]
```
## 
PIP only
```{python}
prepped_data.loc[prepped_data['source_var'] == selected_vars[1], ['Entity', 'region', f'value{reference_vals[0]}', f'value{reference_vals[1]}', f'Year{reference_vals[0]}', f'Year{reference_vals[1]}', 'pip_welfare', f'pop{reference_vals[1]}']].sort_values(['region', 'Entity'])

```

## 
WID only
```{python}
prepped_data.loc[prepped_data['source_var'] == selected_vars[0], ['Entity', 'region', f'value{reference_vals[0]}', f'value{reference_vals[1]}', f'Year{reference_vals[0]}', f'Year{reference_vals[1]}', f'pop{reference_vals[1]}']].sort_values(['region', 'Entity'])

```

## Countries only found in one source - 1980 vs 2018

```{python}

reference_vals = [1980, 2018]
max_dist_from_refs = 5
all_metrics_requirement = False
selected_vars = [
        'WID: Top 10pc share – pretax',
        'PIP: Top 10pc share'
        ]

```

```{python}
prepped_data = prep_data(
        source_dfs = source_dfs,
        df_pop = df_pop,
        df_regions = df_regions,
        selected_vars = selected_vars,
        reference_vals = reference_vals,
        max_dist_from_refs = max_dist_from_refs,
        min_dist_between = 1,
        all_metrics_requirement = all_metrics_requirement,
        reference_col = "Year",
        group_by_col = "Entity",
        outlier_cut_off_upper = None,
        tolerance_lookup =  tolerance_lookup
        )

prepped_data['source_count'] = prepped_data.groupby('Entity')['source_var'].transform('count')

prepped_data = prepped_data[prepped_data['source_count']==1]
```

## {.scrollable} 
PIP only
```{python}
prepped_data.loc[prepped_data['source_var'] == selected_vars[1], ['Entity', 'region', f'value{reference_vals[0]}', f'value{reference_vals[1]}', f'Year{reference_vals[0]}', f'Year{reference_vals[1]}', 'pip_welfare', f'pop{reference_vals[1]}']].sort_values(['region', 'Entity'])

```

## {.scrollable} 
WID only
```{python}
prepped_data.loc[prepped_data['source_var'] == selected_vars[0], ['Entity', 'region', f'value{reference_vals[0]}', f'value{reference_vals[1]}', f'Year{reference_vals[0]}', f'Year{reference_vals[1]}', f'pop{reference_vals[1]}']].sort_values(['region', 'Entity'])

```

# Ex-slide graveyard

## {.smaller}
### Different 'worlds': *WID*
```{python}

reference_vals = [1990, 2018]
max_dist_from_refs = 5
all_metrics_requirement = False
selected_vars = [
        'WID: Top 1pc share – pretax',
        'PIP: Gini'
        ]

```

```{python}

plot_change_scatter(
      plot_var = 'WID: Top 1pc share – pretax',
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_regions = df_regions,
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      all_metrics_requirement = all_metrics_requirement,
      reference_col = "Year",
      group_by_col = "Entity",
      outlier_cut_off_upper = None,
      tolerance_lookup =  tolerance_lookup,
      region_col = 'region'
      )[0]

```

## {.smaller}
### Different 'worlds': *World Bank PIP*

```{python}

plot_change_scatter(
      plot_var = 'PIP: Gini',
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_regions = df_regions,
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      all_metrics_requirement = all_metrics_requirement,
      reference_col = "Year",
      group_by_col = "Entity",
      outlier_cut_off_upper = None,
      tolerance_lookup =  tolerance_lookup,
      region_col = 'region'
      )[0]

```

## {.smaller}
### Different worlds: a comparison

```{python}

df_summary = change_summary_table(
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_pop_regions = df_pop_regions,
      df_regions = df_regions,
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      all_metrics_requirement = all_metrics_requirement,
      reference_col = "Year",
      group_by_col = "Entity",
      outlier_cut_off_upper = None,
      tolerance_lookup =  tolerance_lookup,
      region_col = 'region'
      )

df_summary

df_summary.drop(["Other N. America", "Other Oceania"], axis=0, inplace=True)

dfi.export(df_summary,"images/tables/pres_table1.png")

```

![](images/tables/pres_table1.png)

# Aligning inequality indices

, not so sensitive to top [Osberg...]

##
```{python}

reference_vals = [1990, 2015]
max_dist_from_refs = 3
all_metrics_requirement = False
selected_vars = [
        'PIP: Gini',
        'PIP: Top 10pc share'
        ]

print(make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement, selected_vars))

```

```{python}

df_summary = change_summary_table(
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_pop_regions = df_pop_regions,
      df_regions = df_regions,
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      all_metrics_requirement = all_metrics_requirement,
      reference_col = "Year",
      group_by_col = "Entity",
      outlier_cut_off_upper = None,
      tolerance_lookup =  tolerance_lookup,
      region_col = 'region'
      )

df_summary.drop(["Other N. America", "Other Oceania"], axis=0, inplace=True)

dfi.export(df_summary,"images/tables/pres_table2.png")

```
![](images/tables/pres_table2.png)

##
Observations
- All very similar
- MENA is interesting.

##
Find countries where the measures moved in different directions... Plot them (the whole time series.)

##
```{python}

reference_vals = [1990, 2015]
max_dist_from_refs = 3
all_metrics_requirement = False
selected_vars = [
        'WID: Gini – pretax',
        'WID: Top 1pc share – pretax'
        ]

print(make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement, selected_vars))

```

```{python}

df_summary = change_summary_table(
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_pop_regions = df_pop_regions,
      df_regions = df_regions,
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      all_metrics_requirement = all_metrics_requirement,
      reference_col = "Year",
      group_by_col = "Entity",
      outlier_cut_off_upper = None,
      tolerance_lookup =  tolerance_lookup,
      region_col = 'region'
      )

df_summary.drop(["Other N. America", "Other Oceania"], axis=0, inplace=True)

dfi.export(df_summary,"images/tables/pres_table3.png")

```
![](images/tables/pres_table3.png)

##
Observations
- All very similar
- LAC – reverses trend among small number of countries

##
Find countries where the measures moved in different directions... Plot them (the whole time series.)

## 
General relationship – regression

## {.smaller}
###
Take home message isn't all that dependent on the measure
- Concern that Gini misses what's happening at the extremes.
- In individual cases there can be important differences
- But looking globally it's less of an issue.

## Alternative spec - +/-5 years

##
```{python}

reference_vals = [1995, 2018]
max_dist_from_refs = 5
all_metrics_requirement = False
selected_vars = [
        'WID: Top 10pc share – pretax',
        'PIP: Top 10pc share'
        ]

print(make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement, selected_vars))

```

```{python}

df_summary = change_summary_table(
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_pop_regions = df_pop_regions,
      df_regions = df_regions,
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      all_metrics_requirement = all_metrics_requirement,
      reference_col = "Year",
      group_by_col = "Entity",
      outlier_cut_off_upper = None,
      tolerance_lookup =  tolerance_lookup,
      region_col = 'region'
      )

df_summary.drop(["Other N. America", "Other Oceania"], axis=0, inplace=True)

dfi.export(df_summary,"images/tables/pres_table6.png")

```
![](images/tables/pres_table6.png)

##
```{python}

all_metrics_requirement = True

print(make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement, selected_vars))

```

```{python}

df_summary = change_summary_table(
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_pop_regions = df_pop_regions,
      df_regions = df_regions,
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      all_metrics_requirement = all_metrics_requirement,
      reference_col = "Year",
      group_by_col = "Entity",
      outlier_cut_off_upper = None,
      tolerance_lookup =  tolerance_lookup,
      region_col = 'region'
      )

df_summary = df_summary.drop(["Other N. America", "Other Oceania"], axis=0)

dfi.export(df_summary,"images/tables/pres_table7.png")

```
![](images/tables/pres_table7.png)

## Bibliography