### 2023-04-01 PhD_pages/Paper_1/paper1.qmd
Draftind Paper 1

---
title: "Paper 1: Triangulating trends in inequality around the world from secondary databases"
format: html
bibliography: references_07_03_2023.bib
execute:
  echo: false
  warning: false
---

::: {.callout-note collapse="true" appearance="minimal"}
### Load packages and data

```{python}

from plotnine import *
import plotly.express as px
import pandas as pd
import numpy as np
from IPython.display import display, HTML, IFrame, Markdown
import dataframe_image as dfi
import imgkit

from wrangling_and_plotting_functions import *

pd.set_option('display.max_rows', None)

```

```{python}

fp = 'data_appendix/data/clean/pip.csv'

df_pip = pd.read_csv(fp)

fp = 'data_appendix/data/clean/wid.csv'

df_wid = pd.read_csv(fp)

fp = 'data_appendix/data/clean/lis.csv'

df_lis = pd.read_csv(fp)

fp = 'data_appendix/data/clean/region_mapping.csv'

df_regions = pd.read_csv(fp)

fp = 'data_appendix/data/clean/population.csv'

df_pop = pd.read_csv(fp)

fp = 'data_appendix/data/clean/region_population.csv'

df_pop_regions = pd.read_csv(fp)

# Gather data into dictionary
source_dfs = {
    "PIP": df_pip,
    "WID": df_wid,
    "LIS": df_lis
    }

# This specifies the +/- for the 'stable' category in the tables below
tolerance_lookup =  {
                'var_search_term': [
                    'Gini',
                    'Top 1pc share', 
                    'Top 10pc share',
                    'Bottom 50pc share'],
                'var_tolerance': [
                    1,
                    1,
                    1,
                    1
                ]
            }

```

```{python}
def make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement,selected_vars):
    msg = f'{reference_vals[0]} vs {reference_vals[1]} (+/- {max_dist_from_refs} years)'

    if len(selected_vars)>1:
        if all_metrics_requirement:
            msg = f'{msg} – Matched countries only'
        else:
            msg = f'{msg} – All available countries'

    return(msg)

```

:::

READING LIST:
Bourguignon (2015)
The share of wages in total GDP declined in a majority of countries
(91 out of 133 with data) from 1995 to 2014 (ILO, 2016).

 Other authors
question the pre-eminence of globalization as the main driver of global inequality, noting
that inequality trends differ across countries at similar levels of development and that
are equally exposed to trade (Ravallion, 2018; Corlett, 2016). Ravallion (2018) stresses
the vital role that natio

Key point:  the share of income going to both the top 1 per cent and the bottom 50 per
cent has increased in many countries in recent decades.

In 2021, while the very richest individuals buy trips to space, roughly three billion people are unable to afford a healthy diet. Even  rich countries secured COVID vaccines for their entire population, low-income had barely begun. COVID vaccination rates  at all, the richest countries had already administered more than one dose for every citizen.

Such

"Inequality defines our time," states the UN Secretary-General.^[United Nations Secretary-General António Guterres made the comment in the 18th Nelson Mandela Annual Lecture in July 2020. See https://www.nelsonmandela.org/news/entry/annual-lecture-2020-secretary-general-guterress-full-speech.]

Just as the demand for data to monitor inequality globally is greater than ever, so is its supply.

In part, this increased concern is a reflection of events - of the severe and uneven impacts of the 2007-8 Global Financial Crisis, the COVID-19 pandemic and recent sharp rises in the cost of living.

# Introduction

Economic inequalities around the world are enormous. As the very richest individuals buy trips to space, roughly three billion people are unable to afford a healthy diet.

Such stark disparities have gained more attention in recent decades, among policimakers, academics and the public alike. "Inequality defines our time," according to the UN Secretary-General.^[United Nations Secretary-General António Guterres made the comment in the 18th Nelson Mandela Annual Lecture in July 2020. See https://www.nelsonmandela.org/news/entry/annual-lecture-2020-secretary-general-guterress-full-speech.]

In part at least, this increased concern reflects an increased availability of distributional data. New and expanded sources of data have shed new light on extent of inequality around the world and how it is changing. This has enabled concerns about high or rising inequality to be expressed in new ways, formulated more precisely and subsequently monitored. The supply and demand for global data on inequality have been mutually reinforcing.

Two sources of data have been particularly impactful.

One is the collection of national household survey data brought together by the World Bank, published online through its *Poverty and Inequality Platform*. This is produced as part of the World Bank's remit to monitor global poverty and shared prosperity. Most extensive harmonized . Its impact lies in its extensive use also by the UN in setting and monitoring many of the Sustainable Development Goals.

Another especially important source is the World Inequality Database.

 Capital in the 21st C -> WID. "The 1 %".

Both datasets have continued to improve adn expand. In the case of PIP, including high income countries, following recommendations of the Atkinson commission. In the case of WID – transition from earlier, using with more minimal requirements in terms of source data.

In terms of data availability we are thus in a better position than ever to take a global viw on inequlity. Yet amidst apparent wealth of data, limited understanding of very basics.

Contradictory messages emerge concerning the most basic descriptions from the two databases.

The view from *World Inequality Database* and its associated global report is that inequality is rising in 'nearly everywhere' [@alvaredo_world_2018; @chancel_world_2021].[^WID_quotes] The World Bank data and its biannual report offers a much more mixed picture – in which inequality is rising and falling in roughly equal numbers of countries.

[^WID_quotes]:
    > Income and wealth inequalities have been on the rise nearly everywhere since the 1980s, following a series of deregulation and liberalization programs which took different forms in different countries. [@chancel_world_2021, pp. xx]

    > In recent decades, income inequality has increased in nearly all countries, but at different speeds... [@alvaredo_world_2018, pp. xx]

This stark difference is reflected in decompositions of global inequality into between- and within-country components. (Explain)

Thus despite the great data, there remains fundamental ambiguity surrounding which of two possible worlds we live in. In one world, rising inequality within countries has been an almost universal phenomenon, and has now come to dominate income differences between countries. In another, within-country inequality trends in recent decades have been mixed, and the differences in average incomes between countries remain the lion's share of the globally.

Which of these worlds we live in matters hugely – both in our conception, and of where to place efforts to combat inequality in the future.

The first world suggests the winners from globalization and economic growth overwhelmingly and almost universally, have been a global elite. It situates inquality within countries, but emphasises global factors – including fundamental laws of capitalism with regard to capital accumulation (Piketty) – and the need for international coordination over and above national policy.

The second world and the importance of access to both produced and human capital , national policy (Ravallion, 2018; Corlett, 2016).

### Unpicking the two worlds

Many methodolodical choices concerning underlying data sources, the constraints on temporal coverage this gives rise to,  adjust, conception of inequality. These to some extent come bundled together when making the choice between PIP and WID. That is to say, the empirical question of 'world' of inequality we live in depends on which 'world' of inequality *measurement* we adopt.

Whilst recommending particular aspects, for particular countries, the choice put before us requires wholesale adoption of a bundle whose relative strengths and weaknesses when applied globally is not well understood.

What we can do is triangulate

In this paper, I focus on trends in within-country inequality in recent decades.(In a companion paper, I look at the global distribution.) I line up what I can. I triangulate what we know – . And Identify what we don't know – the remaining differences – and try to pinpoint where the differences lie.

I do so with respect to two separate time periods, chosen to tradeoffs with respect to global coverage. The first period – from the fall of the Berlin wall. Driven primarily to the steep drop off in availability before this in household survey data collected in PIP.

This maximises global coverage – 100 countries covering roughly 90% of the world poplation. The downside is that this period misses out – transition, liberalisation and structural adjustment. Question begging, by avoiding specifically the period where inequality thought to rise.

To address this, I analyse a second, longer period extending back to roughly 1980.

(SUMMAR OF FINDINGS)

## Data

### PIP

### WID

(Coverage chart - overlayed 3/5 year)

### LIS

Will be used for income concept. i.e. it represents the PIP world – indeed LIS is the source for PIP for a number of high income countries. But the access to microdata allows us to ask more questions for a subset of countries.

## Methods

### Concepts of global inequality

Adapted three concepts.

### Aligning coverage

## Comparing inequality trends

The focus of this paper is on trends. But worth pointing out levels. Inequality is much higher in WID than in PIP.

Again this relates to many things – pretax vs disposable income/consumption. Later we will try lining up income concepts more closely. But for global coverage we have to pull together these 'worlds' that combine many differences.

What we're trying to do with triangulation is see whether they show a qualitively similar picture of whether, and how rapidly, inequality was rising or falling.

But what would 'similar' be for Gini and top 1. Not a 1 point move.

The main way of operationalising 'similar' is to look at Relative change: a percentage, rather than percentage point. We can put some threshold and see if there were rises or falls.

No reason to expect this given the properties of the two measures. If we take lognormality we see... (MAke a chart in appendix - where we plot Gini against top 10 and top 1% assuming lognormality, and then add actual data for a particular country.) How unlognormal the top is would be another way to judge similarity. Rather than doing this, we bridge by aligning the indices.

- We also do absolute change: a percentage point. This is more intuitive. But somewhat arbitrary – differences in levels affect assessment of differences in trends.

- A third way is residuals from a regression of the two measures
    - This we will use to see countries that are particular outliers.

## Global data on inequality

### Global databases

A global view faces the hard limitation of the underlying data. This has two sides: firstly is there any data? Secondly the comparability of the data, across countries and over time.

Much improved – *harmonized* secondary databases. Three of note.

- PIP

- WID

- LIS

### Two worlds of inequality measurement

#### Connecting up the worlds is valuable

- for Policy
- for academic research

## Triangulating a global view

### Since 1990

### Since 1980

### Comparing metrics

Does it make a difference whether we look at Gini or top 10%?

People worry about whether Gini masks what's happening at the top.

Empirically: 
    - Make a linear model of top 10% on Gini, for each source separately
    - Make predicted top 10%
    - Take a look at the R2
    - Investigate individual countries with very poor predictions

Sub-Saharan Africa: example where the Gini fell in more cases than the top 10%.

### Since 1990

Similarities:

Convergence – Smaller rises in coutries that started out with high inequality. Driven regionally. (We could do a test for this.)

Indeed for some regions very similar picture:

    - MENA - Roughly 2 point fall, both Gini and top 10%. Only one country with a rise: Not total agreeement on coverage (one country either side)

    - South and South East Asia – Very similar not weighting by population. But wildly different when weighting by population, due to the much much stronger rise

    -  Sub-Saharan Afric – Very similar across sources. Top 10: Roughly even mix of rises and falling countries, but with a 3 point fall on average. Gini: More falls, 5 point on average. (Mention data quality issues: e.g. Kenya big fall. Lesoto big fall. But how comparable are the surveys?? Same surveys underlie WID...)

    - Western Europe – Rises in a majority of countries, but relatively moderate of 1-2 points (Whether Gini of top 10).

Differences:

Verry different bottom line at the global level:

- PIP: roughly equal numbers falling and rising. 0 average change – though a modest rise when pop weighted (2 points). Same whether Gini or Top 10. That's without LAtin America (Which WID is missing). Includung LatAm tips nudges it (with about a 1 point fall unweighter and 1 point rise weighted)

- WID: Far more countries with rising inequality. On average a 3 point rise (whether top 10 or Gini) – [Better to express these relative to the level of ineq. i.e. % increase rather than % point increase? A point rise means 'less' in WID terms because inequality levels are much higher.] But the most stark difference comes when we weight by population. Some key populous countries: India, China, US?, Indonesia? Russia

That difference is driven by particular regions:
-

More countries with rises. LArger rises. Especially larger populations. Average view very different:

Countries where the sources paint a radically different picture that affect the aggregate results:

- India: Insane rise in WID 34->56
- China:
- Russia: Massive drop in PIP. Insane rise in WID.
- USA: Much larger rise in WID.

What explains the different trends in these countries? (We can maybe look in more detail, at different welfare concepts etc.)

It sometimes goes the other way: Indonesia – Rise in PIP, fall in WID.

:::

#### Method 2: Use a regression to obtain the time trend

### Results for a range of inequality metrics

::: {.callout-note collapse="true" appearance="minimal"} 
## Calculations for comparisons

## Explaining differences between PIP and WID data

### For a subset of countries we can bridge between the PIP and WID data

Besides the underlying data sources, there are many differences between what is being measured in the PIP and WID data - notably the welfare concept and unit of analysis/equivalization.

In order to get a better understanding of where differences between the datasets arise, for a subset of countries we can bridge between the two somewhat. For example:

-   WID has post-tax income series for some countries
-   We can use LIS to look at how changing the welfare concept, unit of analysis, equivalisation affects things. (For equivalisation, we could also possibly use the newly available PIP harmonized microdata).

### More in-depth discussion of particular outliers

-   Russia is a very notable case where the two datasets tell very different stories.

# Conclusion