### 2023-04-05 _site/PhD_pages/Paper_1/data_appendix/mytable.tex

\begin{tabular}{lllllllllllllllllllllr}
\toprule
{} &           Entity &        region & change PIP: Gini & change WID: Top 1pc share – pretax & value1995 PIP: Gini & value1995 WID: Top 1pc share – pretax & value2015 PIP: Gini & value2015 WID: Top 1pc share – pretax & Year1995 PIP: Gini & Year1995 WID: Top 1pc share – pretax & Year2015 PIP: Gini & Year2015 WID: Top 1pc share – pretax & pop1995 PIP: Gini & pop1995 WID: Top 1pc share – pretax & pop2015 PIP: Gini & pop2015 WID: Top 1pc share – pretax & region PIP: Gini & region WID: Top 1pc share – pretax & pip\_welfare PIP: Gini & pip\_welfare WID: Top 1pc share – pretax &  plot\_pop\_scale \\
\midrule
0  &          Albania &       EE \& CA &        21.263669 &                          18.181818 &           27.010342 &                                  8.14 &           32.753732 &                                  9.62 &               1996 &                                 1996 &               2015 &                                 2015 &         3284370.0 &                           3284370.0 &         2882482.0 &                           2882482.0 &          EE \& CA &                            EE \& CA &           Consumption &                                     NaN &        1.066413 \\
36 &       Kazakhstan &       EE \& CA &       -24.237001 &                           3.501629 &           35.378308 &                                 12.28 &           26.803667 &                                 12.71 &               1996 &                                 1996 &               2015 &                                 2015 &        16434323.0 &                          16434323.0 &        17835908.0 &                          17835908.0 &          EE \& CA &                            EE \& CA &           Consumption &                                     NaN &        1.495755 \\
40 &           Latvia &       EE \& CA &        10.417068 &                           2.434077 &           30.977489 &                                  9.86 &           34.204435 &                                  10.1 &               1995 &                                 1995 &               2015 &                                 2015 &         2570150.0 &                           2570150.0 &         1991951.0 &                           1991951.0 &          EE \& CA &                            EE \& CA &                Income &                                     NaN &        1.040844 \\
42 &        Lithuania &       EE \& CA &        11.099766 &                          24.861265 &           33.641136 &                                  9.01 &           37.375224 &                                 11.25 &               1993 &                                 1995 &               2015 &                                 2015 &         3747269.0 &                           3747269.0 &         2963764.0 &                           2963764.0 &          EE \& CA &                            EE \& CA &                Income &                                     NaN &        1.068747 \\
29 &          Hungary &       EE \& CA &        -8.654903 &                          59.670782 &           33.289395 &                                  7.29 &            30.40823 &                                 11.64 &               1994 &                                 1995 &               2015 &                                 2015 &        10327171.0 &                          10327171.0 &         9844250.0 &                           9844250.0 &          EE \& CA &                            EE \& CA &                Income &                                     NaN &        1.266299 \\
49 &          Moldova &       EE \& CA &       -26.836525 &                          25.581395 &           36.942146 &                                  9.03 &           27.028158 &                                 11.34 &               1997 &                                 1993 &               2015 &                                 2015 &         4461238.0 &                           4461238.0 &         3277390.0 &                           3277390.0 &          EE \& CA &                            EE \& CA &           Consumption &                                     NaN &        1.077751 \\
26 &          Georgia &       EE \& CA &         -1.69401 &                          10.278373 &           37.132707 &                                 18.68 &           36.503676 &                                  20.6 &               1996 &                                 1997 &               2015 &                                 2015 &         4877682.0 &                           4877682.0 &         3771135.0 &                           3771135.0 &          EE \& CA &                            EE \& CA &           Consumption &                                     NaN &        1.091928 \\
55 &  North Macedonia &       EE \& CA &        26.542688 &                         -18.590998 &           28.128578 &                                 10.22 &           35.594659 &                                  8.32 &               1998 &                                 1996 &               2015 &                                 2015 &         1983971.0 &                           1983971.0 &         2107965.0 &                           2107965.0 &          EE \& CA &                            EE \& CA &                 Mixed &                                     NaN &        1.044175 \\
59 &           Poland &       EE \& CA &        -8.351016 &                          35.974499 &           34.644721 &                                 10.98 &           31.751535 &                                 14.93 &               1995 &                                 1995 &               2015 &                                 2015 &        38401980.0 &                          38401980.0 &        38553148.0 &                          38553148.0 &          EE \& CA &                            EE \& CA &                Income &                                     NaN &        2.090589 \\
60 &          Romania &       EE \& CA &        15.263691 &                          30.492572 &           31.158026 &                                 12.79 &           35.913891 &                                 16.69 &               1995 &                                 1995 &               2015 &                                 2015 &        22533138.0 &                          22533138.0 &        19906078.0 &                          19906078.0 &          EE \& CA &                            EE \& CA &                Income &                                     NaN &        1.555194 \\
18 &          Czechia &       EE \& CA &         0.265217 &                           2.520252 &            25.79901 &                                 11.11 &           25.867433 &                                 11.39 &               1996 &                                 1996 &               2015 &                                 2015 &        10311702.0 &                          10311702.0 &        10523801.0 &                          10523801.0 &          EE \& CA &                            EE \& CA &                Income &                                     NaN &        1.285810 \\
17 &          Croatia &       EE \& CA &         8.233354 &                          48.209366 &           28.748361 &                                  7.26 &           31.115315 &                                 10.76 &               1998 &                                 1998 &               2015 &                                 2015 &         4807100.0 &                           4807100.0 &         4254821.0 &                           4254821.0 &          EE \& CA &                            EE \& CA &                 Mixed &                                     NaN &        1.105815 \\
62 &         Slovakia &       EE \& CA &         0.814271 &                          69.863014 &           26.277685 &                                  6.57 &           26.491657 &                                 11.16 &               1996 &                                 1997 &               2015 &                                 2015 &         5349999.0 &                           5349999.0 &         5424449.0 &                           5424449.0 &          EE \& CA &                            EE \& CA &                Income &                                     NaN &        1.139398 \\
61 &           Russia &       EE \& CA &       -18.138815 &                          34.978229 &           46.103463 &                                 13.78 &           37.740842 &                                  18.6 &               1996 &                                 1995 &               2015 &                                 2015 &       148557456.0 &                         148557456.0 &       144668384.0 &                         144668384.0 &          EE \& CA &                            EE \& CA &           Consumption &                                     NaN &        5.137369 \\
38 &       Kyrgyzstan &       EE \& CA &       -37.466734 &                            -6.3543 &           46.371982 &                                 15.58 &           28.997915 &                                 14.59 &               1998 &                                 1998 &               2015 &                                 2015 &         4596457.0 &                           4596457.0 &         5914985.0 &                           5914985.0 &          EE \& CA &                            EE \& CA &                 Mixed &                                     NaN &        1.153482 \\
2  &          Armenia &       EE \& CA &       -27.154942 &                         -20.716921 &           44.418623 &                                 18.97 &           32.356772 &                                 15.04 &               1996 &                                 1996 &               2015 &                                 2015 &         3322788.0 &                           3322788.0 &         2878598.0 &                           2878598.0 &          EE \& CA &                            EE \& CA &                 Mixed &                                     NaN &        1.066301 \\
73 &          Ukraine &       EE \& CA &       -35.201993 &                         -41.273885 &           39.287536 &                                  15.7 &            25.45754 &                                  9.22 &               1995 &                                 1995 &               2015 &                                 2015 &        51061268.0 &                          51061268.0 &        44982568.0 &                          44982568.0 &          EE \& CA &                            EE \& CA &           Consumption &                                     NaN &        2.275190 \\
5  &          Belarus &       EE \& CA &       -19.942076 &                         -27.177419 &            32.01835 &                                  12.4 &           25.633226 &                                  9.03 &               1998 &                                 1995 &               2015 &                                 2015 &        10461694.0 &                          10461694.0 &         9700609.0 &                           9700609.0 &          EE \& CA &                            EE \& CA &           Consumption &                                     NaN &        1.262175 \\
63 &         Slovenia &       EE \& CA &       -12.939966 &                          10.232558 &           29.184993 &                                  6.45 &           25.408465 &                                  7.11 &               1993 &                                 1995 &               2015 &                                 2015 &         1991115.0 &                           1991115.0 &         2080865.0 &                           2080865.0 &          EE \& CA &                            EE \& CA &                Income &                                     NaN &        1.043397 \\
9  &         Bulgaria &       EE \& CA &        25.568016 &                           3.258508 &           30.714308 &                                 13.81 &           38.567347 &                                 14.26 &               1992 &                                 1995 &               2015 &                                 2015 &         8477584.0 &                           8477584.0 &         7309255.0 &                           7309255.0 &          EE \& CA &                            EE \& CA &                Income &                                     NaN &        1.193514 \\
50 &         Mongolia &     East Asia &        -2.559418 &                          12.735849 &            33.19534 &                                 14.84 &           32.345733 &                                 16.73 &               1995 &                                 1995 &               2016 &                                 2015 &         2330339.0 &                           2330339.0 &         2964756.0 &                           2964756.0 &        East Asia &                          East Asia &           Consumption &                                     NaN &        1.068775 \\
14 &            China &     East Asia &         9.559654 &                          49.253731 &            35.22766 &                                  9.38 &           38.595302 &                                  14.0 &               1996 &                                 1995 &               2015 &                                 2015 &      1218144384.0 &                        1218144384.0 &      1393715456.0 &                        1393715456.0 &        East Asia &                          East Asia &           Consumption &                                     NaN &       41.000000 \\
68 &           Taiwan &     East Asia &         0.506233 &                          51.237624 &           31.324143 &                                 12.12 &           31.482716 &                                 18.33 &               1995 &                                 1995 &               2016 &                                 2015 &        21455818.0 &                          21455818.0 &        23512136.0 &                          23512136.0 &        East Asia &                          East Asia &                Income &                                     NaN &        1.658731 \\
8  &           Brazil &           LAC &       -12.823454 &                          -4.170177 &           59.569725 &                                 23.74 &           51.930829 &                                 22.75 &               1995 &                                 1995 &               2015 &                                 2015 &       163515328.0 &                         163515328.0 &       205188208.0 &                         205188208.0 &              LAC &                                LAC &                Income &                                     NaN &        6.875013 \\
1  &        Argentina &           LAC &       -14.051647 &                          11.587983 &           48.904403 &                                 13.98 &           42.032529 &                                  15.6 &               1995 &                                 1995 &               2016 &                                 2015 &        34946116.0 &                          34946116.0 &        43257064.0 &                          43257064.0 &              LAC &                                LAC &                Income &                                     NaN &        2.225647 \\
15 &         Colombia &           LAC &       -10.443994 &                         -17.310587 &           56.936908 &                                 19.93 &           50.990421 &                                 16.48 &               1996 &                                 1995 &               2015 &                                 2015 &        35970104.0 &                          35970104.0 &        47119728.0 &                          47119728.0 &              LAC &                                LAC &                Income &                                     NaN &        2.336552 \\
48 &           Mexico &           LAC &         -9.71656 &                         122.610141 &           52.808923 &                                 12.03 &           47.677713 &                                 26.78 &               1994 &                                 1995 &               2016 &                                 2015 &        89969568.0 &                          89969568.0 &       120149896.0 &                         120149896.0 &              LAC &                                LAC &                Income &                                     NaN &        4.433394 \\
70 &          Tunisia &          MENA &       -21.224167 &                         -35.823529 &           41.657297 &                                  17.0 &           32.815882 &                                 10.91 &               1995 &                                 1995 &               2015 &                                 2015 &         9294106.0 &                           9294106.0 &        11557779.0 &                          11557779.0 &             MENA &                               MENA &           Consumption &                                     NaN &        1.315498 \\
20 &            Egypt &          MENA &          5.63083 &                           22.41066 &            30.12696 &                                 16.51 &           31.823358 &                                 20.21 &               1995 &                                 1995 &               2015 &                                 2015 &        64166908.0 &                          64166908.0 &        97723792.0 &                          97723792.0 &             MENA &                               MENA &           Consumption &                                     NaN &        3.789496 \\
32 &             Iran &          MENA &        -8.203486 &                         -16.334661 &            42.99837 &                                 20.08 &           39.471004 &                                  16.8 &               1994 &                                 1994 &               2015 &                                 2015 &        60794808.0 &                          60794808.0 &        81790848.0 &                          81790848.0 &             MENA &                               MENA &           Consumption &                                     NaN &        3.332029 \\
34 &           Israel &          MENA &         4.258654 &                          -8.049189 &           38.097106 &                                 17.89 &            39.71953 &                                 16.45 &               1997 &                                 1997 &               2015 &                                 2015 &         5478738.0 &                           5478738.0 &         8007777.0 &                           8007777.0 &             MENA &                               MENA &                Income &                                     NaN &        1.213570 \\
77 &            Yemen &          MENA &         4.884046 &                          11.366774 &           34.997786 &                                 21.73 &           36.707094 &                                  24.2 &               1998 &                                 1998 &               2014 &                                 2014 &        16103338.0 &                          16103338.0 &        28516552.0 &                          28516552.0 &             MENA &                               MENA &           Consumption &                                     NaN &        1.802418 \\
71 &           Turkey &          MENA &         3.725639 &                          -2.744186 &           41.315202 &                                  21.5 &           42.854458 &                                 20.91 &               1994 &                                 1995 &               2015 &                                 2015 &        59305492.0 &                          59305492.0 &        79646176.0 &                          79646176.0 &             MENA &                               MENA &           Consumption &                                     NaN &        3.270451 \\
69 &         Thailand &   S \& SE Asia &       -17.213809 &                          -15.49978 &           43.470212 &                                 22.71 &           35.987333 &                                 19.19 &               1994 &                                 1995 &               2015 &                                 2015 &        59424836.0 &                          59424836.0 &        70294408.0 &                          70294408.0 &      S \& SE Asia &                        S \& SE Asia &           Consumption &                                     NaN &        3.001944 \\
65 &        Sri Lanka &   S \& SE Asia &          11.1477 &                          26.670792 &           35.399539 &                                 16.16 &           39.345773 &                                 20.47 &               1995 &                                 1995 &               2016 &                                 2016 &        18177570.0 &                          18177570.0 &        21336692.0 &                          21336692.0 &      S \& SE Asia &                        S \& SE Asia &           Consumption &                                     NaN &        1.596270 \\
46 &         Malaysia &   S \& SE Asia &       -15.366576 &                         -24.364191 &           48.522672 &                                 19.66 &           41.066398 &                                 14.87 &               1995 &                                 1995 &               2015 &                                 2015 &        20136886.0 &                          20136886.0 &        31068834.0 &                          31068834.0 &      S \& SE Asia &                        S \& SE Asia &                Income &                                     NaN &        1.875699 \\
76 &          Vietnam &   S \& SE Asia &        -0.495084 &                          -1.432836 &           35.442207 &                                 16.75 &           35.266739 &                                 16.51 &               1997 &                                 1997 &               2016 &                                 2016 &        73759112.0 &                          73759112.0 &        92191400.0 &                          92191400.0 &      S \& SE Asia &                        S \& SE Asia &           Consumption &                                     NaN &        3.630650 \\
57 &         Pakistan &   S \& SE Asia &         9.161968 &                           3.663664 &           28.664989 &                                 16.65 &           31.291266 &                                 17.26 &               1996 &                                 1996 &               2015 &                                 2015 &       133117480.0 &                         133117480.0 &       210969296.0 &                         210969296.0 &      S \& SE Asia &                        S \& SE Asia &           Consumption &                                     NaN &        7.041000 \\
58 &      Philippines &   S \& SE Asia &        -5.915094 &                         -11.019141 &           42.888387 &                                 19.33 &           40.351499 &                                  17.2 &               1994 &                                 1994 &               2015 &                                 2015 &        69250472.0 &                          69250472.0 &       103031360.0 &                         103031360.0 &      S \& SE Asia &                        S \& SE Asia &           Consumption &                                     NaN &        3.941887 \\
39 &             Laos &   S \& SE Asia &        11.196288 &                           1.965856 &            34.89544 &                                 19.33 &           38.802434 &                                 19.71 &               1997 &                                 1997 &               2018 &                                 2018 &         4927432.0 &                           4927432.0 &         6787420.0 &                           6787420.0 &      S \& SE Asia &                        S \& SE Asia &           Consumption &                                     NaN &        1.178531 \\
4  &       Bangladesh &   S \& SE Asia &        -1.692956 &                          -3.351287 &           32.942983 &                                 16.71 &           32.385273 &                                 16.15 &               1995 &                                 1995 &               2016 &                                 2016 &       117793336.0 &                         117793336.0 &       157830000.0 &                         157830000.0 &      S \& SE Asia &                        S \& SE Asia &           Consumption &                                     NaN &        5.515265 \\
31 &        Indonesia &   S \& SE Asia &        13.395688 &                          -4.942166 &           35.625137 &                                 19.02 &           40.397369 &                                 18.08 &               1996 &                                 1995 &               2015 &                                 2015 &       198140160.0 &                         198140160.0 &       259091968.0 &                         259091968.0 &      S \& SE Asia &                        S \& SE Asia &           Consumption &                                     NaN &        8.422698 \\
30 &            India &   S \& SE Asia &         9.497208 &                          64.794007 &           31.696184 &                                 13.35 &           34.706436 &                                  22.0 &               1993 &                                 1995 &               2015 &                                 2013 &       964279104.0 &                         964279104.0 &      1322866560.0 &                        1322866560.0 &      S \& SE Asia &                        S \& SE Asia &           Consumption &                                     NaN &       38.965787 \\
22 &         Ethiopia &     SS Africa &       -21.477605 &                         -53.452341 &           44.564513 &                                 29.69 &           34.993123 &                                 13.82 &               1995 &                                 1995 &               2015 &                                 2015 &        57476540.0 &                          57476540.0 &       102471896.0 &                         102471896.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        3.925823 \\
72 &           Uganda &     SS Africa &         9.502102 &                          11.074919 &           39.043719 &                                 15.35 &           42.753693 &                                 17.05 &               1996 &                                 1996 &               2016 &                                 2016 &        20680834.0 &                          20680834.0 &        37477356.0 &                          37477356.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        2.059701 \\
7  &         Botswana &     SS Africa &       -12.280875 &                         -36.973392 &            60.79065 &                                 36.08 &           53.325027 &                                 22.74 &               1993 &                                 1993 &               2015 &                                 2015 &         1543641.0 &                           1543641.0 &         2305177.0 &                           2305177.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        1.049837 \\
10 &     Burkina Faso &     SS Africa &       -26.566681 &                          -31.54678 &           48.072493 &                                 21.27 &           35.301228 &                                 14.56 &               1994 &                                 1994 &               2014 &                                 2014 &        10353268.0 &                          10353268.0 &        18718022.0 &                          18718022.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        1.521083 \\
11 &          Burundi &     SS Africa &        15.872792 &                           17.63767 &           33.333812 &                                 12.53 &           38.624819 &                                 14.74 &               1992 &                                 1992 &               2013 &                                 2013 &         5932789.0 &                           5932789.0 &        10727153.0 &                          10727153.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        1.291649 \\
12 &         Cameroon &     SS Africa &         4.936503 &                         -14.277988 &           44.446758 &                                 18.49 &           46.640874 &                                 15.85 &               1996 &                                 1996 &               2014 &                                 2014 &        13211643.0 &                          13211643.0 &        23012650.0 &                          23012650.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        1.644390 \\
16 &    Cote d'Ivoire &     SS Africa &         2.245793 &                           5.195599 &            40.56382 &                                 16.36 &             41.4748 &                                 17.21 &               1995 &                                 1995 &               2015 &                                 2015 &        14299727.0 &                          14299727.0 &        23596738.0 &                          23596738.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        1.661160 \\
21 &         Eswatini &     SS Africa &        -9.717904 &                         -38.848466 &           60.454726 &                                 31.61 &           54.579794 &                                 19.33 &               1994 &                                 1994 &               2016 &                                 2016 &          953586.0 &                            953586.0 &         1133941.0 &                           1133941.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        1.016209 \\
37 &            Kenya &     SS Africa &         -5.41145 &                         -19.926199 &           43.108572 &                                 18.97 &           40.775773 &                                 15.19 &               1994 &                                 1994 &               2015 &                                 2015 &        26878350.0 &                          26878350.0 &        46851496.0 &                          46851496.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        2.328851 \\
25 &           Gambia &     SS Africa &       -25.970265 &                         -26.670259 &           48.519465 &                                 18.56 &           35.918831 &                                 13.61 &               1998 &                                 1992 &               2015 &                                 2015 &         1242160.0 &                           1242160.0 &         2253137.0 &                           2253137.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        1.048343 \\
54 &          Nigeria &     SS Africa &       -30.897099 &                           2.930728 &           51.922246 &                                 11.26 &           35.879779 &                                 11.59 &               1996 &                                 1992 &               2015 &                                 2018 &       108187616.0 &                         108187616.0 &       183995776.0 &                         183995776.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        6.266536 \\
78 &           Zambia &     SS Africa &        18.211681 &                          21.947368 &           48.333683 &                                  19.0 &           57.136058 &                                 23.17 &               1996 &                                 1996 &               2015 &                                 2015 &         8684139.0 &                           8684139.0 &        16248231.0 &                          16248231.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        1.450170 \\
41 &          Lesotho &     SS Africa &        -29.03314 &                          -35.38874 &           63.240365 &                                 22.38 &           44.879701 &                                 14.46 &               1994 &                                 1994 &               2017 &                                 2017 &         1934085.0 &                           1934085.0 &         2118532.0 &                           2118532.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        1.044478 \\
51 &       Mozambique &     SS Africa &         0.815567 &                          29.194352 &           53.559865 &                                 24.08 &           53.996682 &                                 31.11 &               1996 &                                 1996 &               2014 &                                 2014 &        15594829.0 &                          15594829.0 &        26843252.0 &                          26843252.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        1.754374 \\
44 &       Madagascar &     SS Africa &        -5.761739 &                            2.29265 &           45.255698 &                                 14.83 &           42.648182 &                                 15.17 &               1993 &                                 1997 &               2012 &                                 2012 &        13869136.0 &                          13869136.0 &        24850912.0 &                          24850912.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        1.697170 \\
45 &           Malawi &     SS Africa &       -32.029625 &                           64.15883 &           65.755571 &                                 19.14 &           44.694308 &                                 31.42 &               1997 &                                 1997 &               2016 &                                 2016 &        10112813.0 &                          10112813.0 &        16938944.0 &                          16938944.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        1.470002 \\
47 &       Mauritania &     SS Africa &       -13.585089 &                         -14.205906 &           37.747363 &                                 12.53 &            32.61935 &                                 10.75 &               1995 &                                 1995 &               2014 &                                 2014 &         2380345.0 &                           2380345.0 &         3946225.0 &                           3946225.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        1.096955 \\
53 &            Niger &     SS Africa &       -17.457484 &                         -30.911257 &           41.525965 &                                 16.79 &           34.276576 &                                  11.6 &               1994 &                                 1994 &               2014 &                                 2014 &         9813920.0 &                           9813920.0 &        20128132.0 &                          20128132.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        1.561570 \\
52 &      Netherlands &     W. Europe &        -7.206846 &                           16.41286 &           30.434982 &                                  5.91 &            28.24158 &                                  6.88 &               1993 &                                 1995 &               2015 &                                 2015 &        15446670.0 &                          15446670.0 &        17041110.0 &                          17041110.0 &        W. Europe &                          W. Europe &                Income &                                     NaN &        1.472935 \\
74 &   United Kingdom &     W. Europe &        -6.423917 &                          21.541502 &            35.51664 &                                 10.12 &           33.235081 &                                  12.3 &               1995 &                                 1995 &               2015 &                                 2015 &        57930364.0 &                          57930364.0 &        65224368.0 &                          65224368.0 &        W. Europe &                          W. Europe &                Income &                                     NaN &        2.856373 \\
33 &          Ireland &     W. Europe &       -14.100525 &                          16.699605 &           37.045092 &                                 10.12 &            31.82154 &                                 11.81 &               1995 &                                 1995 &               2015 &                                 2015 &         3579486.0 &                           3579486.0 &         4665764.0 &                           4665764.0 &        W. Europe &                          W. Europe &                Income &                                     NaN &        1.117614 \\
6  &          Belgium &     W. Europe &        -2.403209 &                           5.341615 &           28.412269 &                                  8.05 &           27.729463 &                                  8.48 &               1995 &                                 1995 &               2015 &                                 2015 &        10095204.0 &                          10095204.0 &        11248306.0 &                          11248306.0 &        W. Europe &                          W. Europe &                Income &                                     NaN &        1.306612 \\
43 &       Luxembourg &     W. Europe &        22.185633 &                         -21.158129 &           26.944439 &                                 13.47 &           32.922233 &                                 10.62 &               1994 &                                 1994 &               2015 &                                 2015 &          407639.0 &                            407639.0 &          569415.0 &                            569415.0 &        W. Europe &                          W. Europe &                Income &                                     NaN &        1.000000 \\
67 &      Switzerland &     W. Europe &        -4.576537 &                          16.213494 &           33.856408 &                                  9.93 &           32.306957 &                                 11.54 &               1992 &                                 1995 &               2015 &                                 2015 &         7038302.0 &                           7038302.0 &         8281735.0 &                           8281735.0 &        W. Europe &                          W. Europe &                Income &                                     NaN &        1.221436 \\
66 &           Sweden &     W. Europe &        15.745782 &                          -9.957447 &           25.236506 &                                 11.75 &           29.210191 &                                 10.58 &               1995 &                                 1995 &               2015 &                                 2015 &         8758496.0 &                           8758496.0 &         9849344.0 &                           9849344.0 &        W. Europe &                          W. Europe &                Income &                                     NaN &        1.266445 \\
28 &           Greece &     W. Europe &        -2.637722 &                          26.701031 &           36.975572 &                                   9.7 &            36.00026 &                                 12.29 &               1995 &                                 1995 &               2015 &                                 2015 &        10727928.0 &                          10727928.0 &        10806644.0 &                          10806644.0 &        W. Europe &                          W. Europe &                Income &                                     NaN &        1.293931 \\
19 &          Denmark &     W. Europe &        22.746008 &                          36.029412 &           22.955896 &                                  9.52 &           28.177446 &                                 12.95 &               1995 &                                 1995 &               2015 &                                 2015 &         5235205.0 &                           5235205.0 &         5677795.0 &                           5677795.0 &        W. Europe &                          W. Europe &                Income &                                     NaN &        1.146672 \\
27 &          Germany &     W. Europe &         8.954429 &                          51.025057 &           28.987663 &                                  8.78 &           31.583342 &                                 13.26 &               1995 &                                 1995 &               2015 &                                 2015 &        81123264.0 &                          81123264.0 &        82073232.0 &                          82073232.0 &        W. Europe &                          W. Europe &                Income &                                     NaN &        3.340137 \\
23 &          Finland &     W. Europe &        15.528727 &                           6.696429 &           23.459137 &                                  8.96 &           27.102042 &                                  9.56 &               1995 &                                 1995 &               2015 &                                 2015 &         5107942.0 &                           5107942.0 &         5479457.0 &                           5479457.0 &        W. Europe &                          W. Europe &                Income &                                     NaN &        1.140977 \\
24 &           France &     W. Europe &         1.156214 &                          13.687783 &           32.330525 &                                  8.84 &           32.704335 &                                 10.05 &               1994 &                                 1995 &               2015 &                                 2015 &        57420356.0 &                          57420356.0 &        63809768.0 &                          63809768.0 &        W. Europe &                          W. Europe &                Income &                                     NaN &        2.815757 \\
56 &           Norway &     W. Europe &          5.67384 &                          16.471963 &           26.020532 &                                  8.56 &           27.496896 &                                  9.97 &               1995 &                                 1995 &               2015 &                                 2015 &         4359327.0 &                           4359327.0 &         5190357.0 &                           5190357.0 &        W. Europe &                          W. Europe &                Income &                                     NaN &        1.132676 \\
64 &            Spain &     W. Europe &        -1.289625 &                           8.311911 &            36.65412 &                                 11.67 &           36.181419 &                                 12.64 &               1995 &                                 1995 &               2015 &                                 2015 &        39814568.0 &                          39814568.0 &        46431344.0 &                          46431344.0 &        W. Europe &                          W. Europe &                Income &                                     NaN &        2.316787 \\
35 &            Italy &     W. Europe &         0.452275 &                          27.468785 &           35.238064 &                                  8.81 &           35.397436 &                                 11.23 &               1995 &                                 1995 &               2015 &                                 2014 &        56885124.0 &                          56885124.0 &        60232908.0 &                          60232908.0 &        W. Europe &                          W. Europe &                Income &                                     NaN &        2.713058 \\
13 &           Canada &  W. Offshoots &         7.968926 &                          49.239544 &           31.255736 &                                 10.52 &           33.746482 &                                  15.7 &               1994 &                                 1995 &               2015 &                                 2015 &        29289430.0 &                          29289430.0 &        35732120.0 &                          35732120.0 &     W. Offshoots &                       W. Offshoots &                Income &                                     NaN &        2.009591 \\
75 &    United States &  W. Offshoots &         3.472042 &                          30.096419 &           39.854995 &                                 14.52 &           41.238778 &                                 18.89 &               1995 &                                 1995 &               2015 &                                 2015 &       265660560.0 &                         265660560.0 &       324607776.0 &                         324607776.0 &     W. Offshoots &                       W. Offshoots &                Income &                                     NaN &       10.303787 \\
3  &        Australia &  W. Offshoots &         3.387823 &                          40.714286 &            32.58199 &                                   8.4 &           33.685811 &                                 11.82 &               1995 &                                 1995 &               2016 &                                 2015 &        18003002.0 &                          18003002.0 &        23820240.0 &                          23820240.0 &     W. Offshoots &                       W. Offshoots &                Income &                                     NaN &        1.667578 \\
\bottomrule
\end{tabular}