### 2023-04-19 PhD_pages/Paper_1/data_appendix/1_H_figures_and_tables.qmd
Paper 1: Further drafting and adding figures

---
title: "Appendix 1.H: Preparing figures and tables used in the paper"
format: html
execute:
  echo: false
  warning: false
---

## Set-up

Load packages and functions.
```{python}

from plotnine import *
import plotly.express as px
import pandas as pd
import numpy as np

import dataframe_image as dfi

from G_functions_for_figures_and_tables import *

pd.set_option('display.max_rows', None)

```

Read in data.
```{python}

fp = 'data/clean/pip.csv'

df_pip = pd.read_csv(fp)

fp = 'data/clean/wid.csv'

df_wid = pd.read_csv(fp)

fp = 'data/clean/lis.csv'

df_lis = pd.read_csv(fp)

fp = 'data/clean/region_mapping.csv'

df_regions = pd.read_csv(fp)

fp = 'data/clean/population.csv'

df_pop = pd.read_csv(fp)

fp = 'data/clean/region_population.csv'

df_pop_regions = pd.read_csv(fp)

```

Gather the inequality data into a dictionary – this is the format assumed by the functions that will later be applied.
```{python}
# Gather data into dictionary
source_dfs = {
    "PIP": df_pip,
    "WID": df_wid,
    "LIS": df_lis
    }

```

Specify a mapping between inequality variables and the 'tolerance' that will be used to catagorise whether the metric has risen or fallen substantially (e.g. +/- 1point change). The mapping is achieved by a string match - i.e. 'Gini' within 'var_search_term' will map the correspondending tolerance to all columns whose name includes the string 'Gini'.

```{python}
tolerance_lookup =  {
                'var_search_term': [
                    'Gini',
                    'Top 1pc share', 
                    'Top 10pc share',
                    'Bottom 50pc share'],
                'var_tolerance': [
                    1,
                    1,
                    1,
                    1
                ], 
                'relative_tolerance': [
                    0.05
                ]
            }

```

This builds astring that gets added to the tables identifying what paremeters were set when the table was built.
```{python}
def make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement,selected_vars):
    msg = f'{reference_vals[0]} vs {reference_vals[1]} (+/- {max_dist_from_refs} years)'

    if len(selected_vars)>1:
        if all_metrics_requirement:
            msg = f'{msg} – Matched countries only'
        else:
            msg = f'{msg} – All available countries'

    return(msg)

```

## SECTION 2:

### WID top 1% vs PIP Gini

#### Since 1993

This figure compares relative change in the 'typical' metrics for WID and PIP: Top 1% share (pretax) and Gini
```{python}
# Set-up to compare WID and PIP top 10% share, all countries
reference_vals = [1993, 2015]
max_dist_from_refs = 3
all_metrics_requirement = False
selected_vars = [
        'WID: Top 1pc share – pretax',
        'PIP: Gini'
        ]

all_metrics_requirement = True

```

```{python}

df_prepped = prep_data(
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_regions = df_regions,
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      all_metrics_requirement = all_metrics_requirement,
      reference_col = "Year",
      group_by_col = "Entity",
      outlier_cut_off_upper = None,
      tolerance_lookup =  tolerance_lookup
    )

```

```{python}

df_summary = change_summary_table(
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_pop_regions = df_pop_regions,
      df_regions = df_regions,
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      all_metrics_requirement = all_metrics_requirement,
      reference_col = "Year",
      group_by_col = "Entity",
      outlier_cut_off_upper = None,
      tolerance_lookup =  tolerance_lookup,
      absolute_or_relative_tolerance = "absolute",
      region_col = 'region'
      )

# df_summary = df_summary['WID: Top 1pc share – pretax']
df_summary = df_summary['PIP: Gini']

df_summary
```

#### Table 2

#### Table 3

## Figures

#### Figure 1

This figure compares relative change in the 'typical' metrics for WID and PIP: Top 1% share (pretax) and Gini
```{python}
# Set-up to compare WID and PIP top 10% share, all countries
reference_vals = [1995, 2015]
max_dist_from_refs = 3
all_metrics_requirement = False
selected_vars = [
        'WID: Top 1pc share – pretax',
        'PIP: Gini'
        ]

all_metrics_requirement = True

```

```{python}

# Scatter WID (WID vs PIP top 10% share, only matched countries)

plot_change_scatter(
      plot_var = 'relative_diff',
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_regions = df_regions,
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      all_metrics_requirement = all_metrics_requirement,
      reference_col = "Year",
      group_by_col = "Entity",
      outlier_cut_off_upper = None,
      tolerance_lookup =  tolerance_lookup,
      region_col = 'region'
      )[0]

```

#### Figure 2

#### Figure 3