### 2023-05-03 PhD_pages/Paper_1/data_appendix/prepare_data_for_figures_and_tables.qmd

---
title: "Appendix 1.H: Preparing figures and tables used in the paper"
format: html
execute:
  echo: false
  warning: false
---

```{python}

from plotnine import *
import plotly.express as px
import pandas as pd
import numpy as np
# from IPython.display import display, HTML, IFrame, Markdown
# import dataframe_image as dfi
# import imgkit

from G_functions_for_figures_and_tables import *

# pd.set_option('display.max_rows', None)

```

```{python}

fp = 'data/clean/pip.csv'

df_pip = pd.read_csv(fp)

fp = 'data/clean/wid.csv'

df_wid = pd.read_csv(fp)

fp = 'data/clean/lis.csv'

df_lis = pd.read_csv(fp)

fp = 'data/clean/region_mapping.csv'

df_regions = pd.read_csv(fp)

fp = 'data/clean/population.csv'

df_pop = pd.read_csv(fp)

fp = 'data/clean/region_population.csv'

df_pop_regions = pd.read_csv(fp)

# Gather data into dictionary
source_dfs = {
    "PIP": df_pip,
    "WID": df_wid,
    "LIS": df_lis
    }

# This specifies the +/- for the 'stable' category in the tables below
tolerance_lookup =  {
                'var_search_term': [
                    'Gini',
                    'Top 1pc share', 
                    'Top 10pc share',
                    'Bottom 50pc share',
                    "Ratio Top 10 Bottom 50 share"],
                'var_tolerance': [
                    1,
                    1,
                    1,
                    1,
                    1
                ], 
                'relative_tolerance': [
                    0.05
                ]
            }

```

```{python}

selected_vars = [
        'PIP: Gini',
        "PIP: Top 10pc share",
        'PIP: Bottom 50pc share',
        'PIP: Ratio Top 10 Bottom 50 share',
        'WID: Gini – pretax',
        "WID: Top 1pc share – pretax",
        "WID: Top 10pc share – pretax",
        'WID: Bottom 50pc share – pretax',
        'WID: Ratio Top 10 Bottom 50 share – pretax'
        ]

```

Prep data for 1993-2015 period

```{python}
# Set-up to compare PIP Gini and WID top 1% share, all observations, for shorter time period
reference_vals = [1993, 2015]
max_dist_from_refs = 3

prepped_data_dict = prep_data(
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_regions = df_regions,
      region_col = 'region',
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      reference_col = "Year",
      group_by_col = "Entity",
      tolerance_lookup =  tolerance_lookup,
      outlier_cut_off_upper = None
      )

owid_format = prepped_data_to_owid_format(
      prepped_data_all_obs = prepped_data_dict['all_obs'],
      prepped_data_matching_obs = prepped_data_dict['matching_obs'],
      reference_vals = reference_vals
      )
owid_format

# # 
# Save data to be uploaded to OWID to make charts
fp = 'figures_and_tables/Compare WID and PIP inequality data 1993 vs 2015 (Joe temp).csv'

owid_format.to_csv(fp, index=False)

```

```{python}
summary_tables_dict = prep_summary_tables(
    prepped_data_dict = prepped_data_dict,
    df_pop_regions = df_pop_regions,
    selected_vars = [ 'PIP: Gini',
                    'WID: Top 10pc share – pretax'],
    reference_vals = reference_vals)

summary_tables_dict['all_obs']['absolute_counts']

```

Prep data for 1980-2018 period

```{python}
# Set-up to compare PIP Gini and WID top 1% share, all observations, for shorter time period
reference_vals = [1980, 2018]
max_dist_from_refs = 5

prepped_data_dict = prep_data(
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_regions = df_regions,
      region_col = 'region',
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      reference_col = "Year",
      group_by_col = "Entity",
      tolerance_lookup =  tolerance_lookup,
      outlier_cut_off_upper = None
      )

owid_format = prepped_data_to_owid_format(
      prepped_data_all_obs = prepped_data_dict['all_obs'],
      prepped_data_matching_obs = prepped_data_dict['matching_obs'],
      reference_vals = reference_vals
      )

# # 
# Save data to be uploaded to OWID to make charts
fp = 'figures_and_tables/Compare WID and PIP inequality data 1980 vs 2018 (Joe temp).csv'

owid_format.to_csv(fp, index=False)

# prepped_data_dict['all_obs']
```